---
title: "Cross-group Within-time and Cross-time Within-group Statistics"
description: |
  Compute relative values across countries at each time, and relative values within country across time. 
core:
  - package: dplyr
    code: |
      arrange(v1, v2) %>% group_by(v1) %>% mutate(stats := v3/first(v3))
date: 2023-03-21
date_start: 2023-03-21
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Within Panel Cross-time and Cross-group Statistics

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Comparing Three Countries over Time

Given three time series for three "countries", we compute percentage change from initial year for each country, and compare relative values within each timer period versus one country. 

First, we generate the core data inputs. We assume that the output here would be the data structure we face prior to generating the figures we are interested in. We use data from the attitude dataset, but re-interpret what the columns are. We work with data from three "countries" at the same time, which generalizes also to the two countries case. 

```{r}
# Load data, and treat index as "year"
# pretend data to be country-data
df_attitude <- as_tibble(attitude) %>%
  rowid_to_column(var = "year") %>% 
  select(year, rating, complaints, learning) %>% 
  rename(stats_usa = rating, 
         stats_canada = complaints, 
         stats_uk = learning)

# Wide to Long
df_attitude <- df_attitude %>%
  pivot_longer(cols = starts_with('stats_'),
               names_to = c('country'),
               names_pattern = paste0("stats_(.*)"),
               values_to = "rating")

# Print 
kable(df_attitude[1:10,]) %>% kable_styling_fc()
```

Second, we generate additional data inputs. Specifically, we also generate ratios of values with respect to he "first" country, within each year. 

```{r}
# Sort and get list of countries
ar_countries_sorted <- df_attitude %>% 
  ungroup() %>% distinct(country) %>% arrange(country) %>% 
  pull(country)
st_ratio_var <- paste0('ratings_ratio_vs_country', ar_countries_sorted[1])

# Generate ratio over the first location
df_attitude <- df_attitude %>% 
  arrange(year, country) %>% group_by(year) %>% 
  mutate(!!sym(st_ratio_var) := rating/first(rating))
  
# Print 
kable(df_attitude[1:10,]) %>% kable_styling_fc()
```

Third, we now generate ratios of values with respect to the first year, within each country. 

```{r}
# Sort and get list of countries
ar_years_sorted <- df_attitude %>% 
  ungroup() %>% distinct(year) %>% arrange(year) %>% 
  pull(year)
st_ratio_var <- paste0('ratings_ratio_vs_year', ar_years_sorted[1])

# Generate ratio over the first location
df_attitude <- df_attitude %>% 
  arrange(country, year) %>% group_by(country) %>% 
  mutate(!!sym(st_ratio_var) := rating/first(rating))
  
# Print
# Within each country, we show the first 3 years 
kable(df_attitude %>% 
  group_by(country) %>% 
  slice_min(order_by = year, n = 3)
  ) %>% kable_styling_fc()
```