---
title: "Convert R Markdown File to R, PDF and HTML"
description: |
  Find all files in a folder with a particula suffix, with exclusion.
  Convert R Markdow File to R, PDF and HTML.
  Modify markdown pounds hierarchy.
core:
  - package: r
    code: |
      file()
      writeLines()
      readLines()
      close()
      gsub()
date: 2020-05-16
date_start: 2020-05-16
output:
  pdf_document:
    pandoc_args: '../../../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../../../preamble.tex'
  html_document:
    pandoc_args: '../../../../_output_kniti_html.yaml'
    includes:
      in_header: "../../../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Rmd to HTML

```{r global_options, include = FALSE}
try(source("../../../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Search and Find all Files in Repository

Search inside directories, for all files in a repository that have a particular suffix and that don't contain skip pattern list string items.

```{r}
# Serch Folder and skip list
spt_roots <- c('C:/Users/fan/R4Econ/amto', 'C:/Users/fan/R4Econ/summarize')
spn_skip <- c('summarize', 'panel', 'support')

# Search and get all Path
ls_sfls <- list.files(path=spt_roots, recursive=T, pattern=".Rmd", full.names=T)

# Skip path if contains words in skip list
if(!missing(spn_skip)) {
  ls_sfls <- ls_sfls[!grepl(paste(spn_skip, collapse = "|"), ls_sfls)]
}

# Loop and print
for (spt_file in ls_sfls) {
    st_fullpath_nosufx <- tail(strsplit(spt_file, "/")[[1]],n=1)
    print(paste0(spt_file, '---', st_fullpath_nosufx))
}
```

#### Search and Find all Git Modified or New Rmd

Search inside directories, for all files in a git repo folder that are new or have been modified. Ignore possible subset of file based on string search.

```{r}
# Serch Folder and skip list
spt_roots <- c('C:/Users/fan/R4Econ/amto', 'C:/Users/fan/R4Econ/development')
spn_skip <- c('summarize', 'panel', 'support')
ls_sfls <- list.files(path=spt_roots, recursive=T, pattern=".Rmd", full.names=T)
if(!missing(spn_skip)) {
  ls_sfls <- ls_sfls[!grepl(paste(spn_skip, collapse = "|"), ls_sfls)]
}

# Loop and print
for (spt_file in ls_sfls) {
  spg_check_git_status <- paste0('git status -s ', spt_file)
  st_git_status <- toString(system(spg_check_git_status, intern=TRUE))
  bl_modified <- grepl(' M ', st_git_status, fixed=TRUE)
  bl_anewfile <- grepl('?? ', st_git_status, fixed=TRUE)
  bl_nochange <- (st_git_status == "")

  if (bl_modified == 1) {
    print(paste0('MODIFIED: ', spt_file))
  } else if (bl_anewfile == 1) {
    print(paste0('A NEW FL: ', spt_file))
  } else {
    print(paste0('NO CHNGE: ', spt_file))
  }
}
```

#### Resave an Existing File with Different Name Different Folder

Given an existing Rmd File, Resave it with a different name (add to name suffix), and then save in a different folder:

- old file: */R4Econ/development/fs_rmd_pdf_html.Rmd*
- new file: *R4Econ/development/inout/_file/rmd/fs_rmd_pdf_html_mod.Rmd*

```{r}
# Serch Folder and skip list
spt_roots <- c('C:/Users/fan/R4Econ/development/inout/')
spn_skip <- c('_main', '_file')
ls_sfls <- list.files(path=spt_roots, recursive=T, pattern=".Rmd", full.names=T)
if(!missing(spn_skip)) {
  ls_sfls <- ls_sfls[!grepl(paste(spn_skip, collapse = "|"), ls_sfls)]
}

# Loop and print
for (spt_file in ls_sfls) {
  spt_new <- paste0('_file/rmd/')
  spn_new <- paste0(spt_new, sub('\\.Rmd$', '', basename(spt_file)), '_mod.Rmd')
  print(spt_new)
  print(spn_new)

  fileConn_rd <- file(spt_file, "r")
  st_file_read <- readLines(fileConn_rd)

  fileConn_sr <- file(spn_new)
  writeLines(st_file_read, fileConn_sr)

  close(fileConn_rd)
  close(fileConn_sr)
}
```

##### Replacment Function Change Markdown Hierarchy and Add to YAML

Given an existing Rmd File, Resave it with a different name, and replace (add in) additional yaml contents.

```{r}
spn_file = '_file/rmd/fs_rmd_pdf_html_mod.Rmd'
fileConn_sr <- file(spn_file)
st_file <- readLines(fileConn_sr)
# print(st_file)

st_search <- "html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      toc_depth: 3"
st_replace <- paste0("html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      toc_depth: 3\n",
                     "    toc: true\n",
                     "    number_sections: true\n",
                     "    toc_float:\n",
                     "      collapsed: false\n",
                     "      smooth_scroll: false\n",
                     "      toc_depth: 3")
st_file_updated <- gsub(x = st_file,
                        pattern = st_search,
                        replacement = st_replace)

st_search <- "../../../../"
st_replace <- paste0("../../../../../../../../")
st_file_updated <- gsub(x = st_file_updated,
                        pattern = st_search,
                        replacement = st_replace)

st_file_updated <- gsub(x = st_file_updated, pattern = '# ', replacement = '# ')
st_file_updated <- gsub(x = st_file_updated, pattern = '## ', replacement = '## ')
st_file_updated <- gsub(x = st_file_updated, pattern = '### ', replacement = '# ')

spn_file = '_file/rmd/fs_rmd_pdf_html_mod.Rmd'
fileConn_sr <- file(spn_file)
st_file <- writeLines(st_file_updated, fileConn_sr)

```

#### Search and Render Rmd File and Save HTML, PDF or R

1. Search files satisfying conditions in a folder
2. knit files to HTML (and re-run the contents of the file)
3. Save output to a different folder

```{r}
# Specify Parameters
ar_spt_root = c('C:/Users/fan/R4Econ/amto/array/', 'C:/Users/fan/R4Econ/math/integration')
bl_recursive = TRUE
st_rmd_suffix_pattern = "*.Rmd"
ar_spn_skip <- c('basics', 'integrate', 'main', 'mesh')
ls_bool_convert <- list(bl_pdf=TRUE, bl_html=TRUE, bl_R=TRUE)
spt_out_directory <- 'C:/Users/fan/Downloads/_data'
bl_verbose <- TRUE

# Get Path
ls_sfls  <- list.files(path=ar_spt_root,
                       recursive=bl_recursive,
                       pattern=st_rmd_suffix_pattern,
                       full.names=T)

# Exclude Some Files given ar_spn_skip
if(!missing(ar_spn_skip)) {
  ls_sfls <- ls_sfls[!grepl(paste(ar_spn_skip, collapse = "|"), ls_sfls)]
}

# Loop over files
for (spn_file in ls_sfls) {

  # Parse File Name
  spt_file <- dirname(spn_file)
  sna_file <- tools::file_path_sans_ext(basename(spn_file))

  # Output FIles
  spn_file_pdf <- paste0(spt_file, sna_file, '.pdf')
  spn_file_html <- paste0(spt_file, sna_file, '.html')
  spn_file_R <- paste0(spt_file, sna_file, '.R')

  # render to PDF
  if (ls_bool_convert$bl_pdf) {
    if (bl_verbose) message(paste0('spn_file_pdf:',spn_file_pdf, ', PDF started'))
    rmarkdown::render(spn_file, output_format='pdf_document',
                      output_dir = spt_out_directory, output_file = sna_file)
    if (bl_verbose) message(paste0('spn_file_pdf:',spn_file_pdf, ', PDF finished'))
    spn_pdf_generated <- paste0(spt_out_directory, '/', spn_file_pdf)
  }

  # render to HTML
  if (ls_bool_convert$bl_html) {
    if (bl_verbose) message(paste0('spth_html:',spn_file_html, ', HTML started.'))
    rmarkdown::render(spn_file, output_format='html_document',
                      output_dir = spt_out_directory, output_file = sna_file)
    if (bl_verbose) message(paste0('spth_html:',spn_file_html, ', HTML finished.'))
    spn_html_generated <- paste0(spt_out_directory, '/', spn_file_html)
  }

  # purl to R
  if (ls_bool_convert$bl_R) {
    if (bl_verbose) message(paste0('purl_to:', paste0(spn_file_R, ".R")))
    knitr::purl(spn_file, output=paste0(spt_out_directory, '/', sna_file, '.R'), documentation = 1)
    spn_R_generated <- paste0(spt_out_directory, '/', sna_file, '.R')
  }

  # return(list(ls_spt_pdf_generated=ls_spt_pdf_generated,
  #             ls_spt_html_generated=ls_spt_html_generated,
  #             ls_spt_R_generated=ls_spt_R_generated))


}
