---
title: "Save Text to File, Read Text from File, Replace Text in File"
description: |
  Save data to file, read text from file, replace text in file.
core:
  - package: r
    code: |
      kable()
      file()
      writeLines()
      readLines()
      close()
      gsub()
date: 2020-05-16
date_start: 2020-05-16
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Text to File

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Latex Table to File

Tabular outputs, text outputs, etc are saved as variables, which could be printed in console. They can also be saved to file for future re-used. For example, latex outputs need to be saved to file.

```{r}
# Load Data
dt <- mtcars[1:5, 1:6]
# Generate latex string variable
st_out_tex <- kable(dt, "latex")
print(st_out_tex)
# File out
# fileConn <- file("./../../_file/tex/tex_sample_a_tab.tex")
fileConn <- file("_file/tex/tex_sample_a_tab.tex")
writeLines(st_out_tex, fileConn)
close(fileConn)
```

#### Create a Text File from Strings

```{r}

st_file <- "\\documentclass[12pt,english]{article}
\\usepackage[bottom]{footmisc}
\\usepackage[urlcolor=blue]{hyperref}
\\begin{document}
\\title{A Latex Testing File}
\\author{\\href{http://fanwangecon.github.io/}{Fan Wang} \\thanks{See information \\href{https://fanwangecon.github.io/Tex4Econ/}{Tex4Econ} for more.}}
\\maketitle
Ipsum information dolor sit amet, consectetur adipiscing elit. Integer Latex placerat nunc orci.
\\paragraph{\\href{https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3140132}{Data}}
Village closure information is taken from a village head survey.\\footnote{Generally students went to schools.}
output:
  pdf_document:
    pandoc_args: '..//..//_output_kniti_pdf.yaml'
    includes:
      in_header: '..//..//preamble.tex'
  html_document:
    pandoc_args: '..//..//_output_kniti_html.yaml'
    includes:
      in_header: '..//..//hdga.html'
\\end{document}"

print(st_file)

fl_test_tex <- "_file/tex/test_fan.tex"
fileConn <- file(fl_test_tex)
writeLines(st_file, fileConn)
close(fileConn)
```

#### Open A File and Read Lines

Open and Replace Text in File:

```{r}
fileConn <- file(fl_test_tex, "r")
st_file_read <- readLines(fileConn)
print(st_file_read)
close(fileConn)
```

#### Open a File and Replace Some Lines

Append additional strings into the file after *html_document* with proper spacings:

```{r}
# Read in Lines from existing file
fileConn <- file(fl_test_tex, "r")
st_file_read <- readLines(fileConn)
close(fileConn)

# Search and Replace String
st_search <- "html_document:"
st_replace <- paste0("html_document:\r\n",
                     "    toc: true\r\n",
                     "    number_sections: true\r\n",
                     "    toc_float:\r\n",
                     "      collapsed: false\r\n",
                     "      smooth_scroll: false\r\n",
                     "      toc_depth: 3")

# Search and Replace
st_file_updated <- gsub(x = st_file_read,
                        pattern = st_search,
                        replacement = st_replace)

# Print
print(st_file_updated)

# Save Updated File
fl_srcrep_tex <- "_file/tex/test_fan_search_replace.tex"
fileConn_sr <- file(fl_srcrep_tex)
writeLines(st_file_updated, fileConn_sr)
close(fileConn_sr)
```
