---
title: "Evaluate a Many One Unknown Nonlinear Equations Jointly, Brute Force Find the Roots of All Nonlinear Equations"
output:
  html_document: default
  word_document: default
  pdf_document: default
  html_notebook: default
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/CodeDynaAsset/)'s [R4Econ](https://fanwangecon.github.io/R4Econ/) Repository or [Intro Stats with R](https://fanwangecon.github.io/Stat4Econ/) Repository.

## Issue and Goal

We want evaluate linear function $0=f(z_{ij}, x_i, y_i, \textbf{X}, \textbf{Y}, c, d)$. There are $i$ functions that have $i$ specific $x$ and $y$. For each $i$ function, we evaluate along a grid of feasible values for $z$, over $j\in J$ grid points, looking for the $j$ that is closest to the root. $\textbf{X}$ and $\textbf{Y}$ are arrays common across the $i$ equations, and $c$ and $d$ are constants.

The evaluation strategy is the following, given min and max for $z$ that are specific for each $j$, and given common number of grid points, generate a matrix of $z_{ij}$. Suppose there the number of $i$ is $I$, and the number of grid points for $j$ is $J$.

1. Generate a $J \cdot I$ by $3$ matrix where the columns are $z,x,y$ as tibble
2. Follow [this](https://fanwangecon.github.io/R4Econ/support/function/fs_funceval.html) Mutate to evaluate the $f(\cdot)$ function.
3. Add two categorical columns for grid levels and wich $i$, $i$ and $j$ index. Plot Mutate output evaluated column categorized by $i$ as color and $j$ as x-axis.

## Set Up

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(tidyverse)
library(tidyr)
library(knitr)
library(kableExtra)
# file name
st_file_name = 'fs_func_graph_eval'
# Generate R File
purl(paste0(st_file_name, ".Rmd"), output=paste0(st_file_name, ".R"), documentation = 2)
# Generate PDF and HTML
# rmarkdown::render("C:/Users/fan/R4Econ/support/function/fs_funceval.Rmd", "pdf_document")
# rmarkdown::render("C:/Users/fan/R4Econ/support/function/fs_funceval.Rmd", "html_document")
```

## Set up Input Arrays

There is a function that takes $M=Q+P$ inputs, we want to evaluate this function $N$ times. Each time, there are $M$ inputs, where all but $Q$ of the $M$ inputs, meaning $P$ of the $M$ inputs, are the same. In particular, $P=Q*N$.

$$M = Q+P = Q + Q*N$$

Now we need to expand this by the number of choice grid. Each row, representing one equation, is expanded by the number of choice grids. We are graphically searching, or rather brute force searching, which means if we have 100 individuals, we want to plot out the nonlinear equation for each of these lines, and show graphically where each line crosses zero. We achieve this, by evaluating the equation for each of the 100 individuals along a grid of feasible choices. 

In this problem here, the feasible choices are shared across individuals.

```{r setup}
# it_child_count = N, the number of children
it_N_child_cnt = 5
# it_heter_param = Q, number of parameters that are heterogeneous across children
it_Q_hetpa_cnt = 2

# P fixed parameters, nN is N dimensional, nP is P dimensional
ar_nN_A = seq(-2, 2, length.out = it_N_child_cnt)
ar_nN_alpha = seq(0.1, 0.9, length.out = it_N_child_cnt)
ar_nP_A_alpha = c(ar_nN_A, ar_nN_alpha)

# N by Q varying parameters
mt_nN_by_nQ_A_alpha = cbind(ar_nN_A, ar_nN_alpha)

# Choice Grid for nutritional feasible choices for each
it_N_choice_cnt = 3
fl_N_agg = 100
fl_N_min = 0
ar_N_choices = seq(fl_N_min, fl_N_agg, length.out = it_N_choice_cnt)

# Mesh Expand
tb_states_choices <- as_tibble(mt_nN_by_nQ_A_alpha) %>% expand_grid(choices = ar_N_choices)

# display
kable(tb_states_choices) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

## Evaluate nonlinear function over entire tibble dataframe


```{r linear_apply}

# Convert Matrix to Tibble
ar_st_col_names = c('fl_A', 'fl_alpha', 'fl_N')
tb_states_choices <- tb_states_choices %>% rename_all(~c(ar_st_col_names))

# Define Implicit Function
ffi_nonlin_dplyrdo <- function(fl_A, fl_alpha, fl_N, ar_A, ar_alpha, fl_N_agg, fl_rho){

  # Test Parameters
  # ar_A = ar_nN_A
  # ar_alpha = ar_nN_alpha
  # fl_N = 100
  # fl_rho = -1
  # fl_N_q = 10

  # Apply Function
  ar_p1_s1 = exp((fl_A - ar_A)*fl_rho)
  ar_p1_s2 = (fl_alpha/ar_alpha)
  ar_p1_s3 = (1/(ar_alpha*fl_rho - 1))
  ar_p1 = (ar_p1_s1*ar_p1_s2)^ar_p1_s3
  ar_p2 = fl_N^((fl_alpha*fl_rho-1)/(ar_alpha*fl_rho-1))
  ar_overall = ar_p1*ar_p2
  fl_overall = fl_N_agg - sum(ar_overall)

  return(fl_overall)
}

# Parameters
fl_rho = 0.20

# fl_A, fl_alpha are from columns of tb_nN_by_nQ_A_alpha
tb_states_choices_eval = tb_states_choices %>% rowwise() %>%
                        mutate(dplyr_eval = ffi_nonlin_dplyrdo(fl_A, fl_alpha, fl_N,
                                                               ar_nN_A, ar_nN_alpha,
                                                               fl_N_agg, fl_rho))
# Show
kable(tb_states_choices_eval) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```
