---
title: "Multinomial-Logit and CRRA"
titleshort: "Multinomial-Logit and CRRA"
description: |
  Multinomial logit and CRRA
core:
  - package: r
    code: |
date: 2021-01-08
date_start: 2020-01-08
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Multinomial Logit and CRRA

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Multinomial Logit Expected Value

The log-sum term. $U_{tj}$ denotes pecuniary related utility at time $t$ for choice alternative $j$.

$$
\log\left(
  \exp\left(U_{11}\right)
  +\exp\left(U_{12}\right) 
\right)
+ 
\log\left(
  \exp\left(U_{21}\right)
  +\exp\left(U_{22}\right) 
\right)
$$

#### Two Regimes with Utility (Does not Work)

With numerical values, utility under regime A. 

```{r}
UA11 = -0.5
UA12 = -1
UA21 = -0.6
UA22 = -1.1
fl_EV_A = log(exp(UA11) + exp(UA12)) + log(exp(UA21) + exp(UA22))
print(fl_EV_A)
```

With numerical values, utility under regime B. 

```{r}
UB11 = -0.1
UB12 = -0.65
UB21 = -0.21
UB22 = -0.2
fl_EV_B = log(exp(UB11) + exp(UB12)) + log(exp(UB21) + exp(UB22))
print(fl_EV_B)
```

How much must U under A increase by to match overall U under B? Note that the structure below does not make sense, but U is already negative, it can not be about how chaning Shares of U impact things.

```{r}
ar_increase = seq(0, 0.5, length.out=10)
fl_EV_A = log(exp(UA11*(1+ar_increase)) + exp(UA12*(1+ar_increase))) + log(exp(UA21*(1+ar_increase)) + exp(UA22*(1+ar_increase)))
print(fl_EV_A)
```

#### Two Regimes with Consumption

Preference CRRA

```{r}
gamma = 1.3
# Define utility Function
ffi_crra <- function(fl_c){
  fl_U = (fl_c^(1-gamma))/(1-gamma)
  return(fl_U)
}
```

```{r}
fl_increment = 1.5
CA11 = -0.5+fl_increment
CA12 = -1+fl_increment
CA21 = -0.6+fl_increment
CA22 = -1.1+fl_increment
fl_EV_A = log(exp(ffi_crra(CA11)) + exp(ffi_crra(CA12))) + log(exp(ffi_crra(CA21)) + exp(ffi_crra(CA22)))
print(fl_EV_A)
```

With numerical values, utility under regime B. 

```{r}
fl_increment = 3
CB11 = -0.1+fl_increment
CB12 = -0.65+fl_increment
CB21 = -0.21+fl_increment
CB22 = -0.2+fl_increment
fl_EV_B = log(exp(ffi_crra(CB11)) + exp(ffi_crra(CB12))) + log(exp(ffi_crra(CB21)) + exp(ffi_crra(CB22)))
print(fl_EV_B)
```

How much must C under A increase by to match overall U under B? This structure works, negative or positive values for overall EV does not matter, work in either case. 

```{r}
ar_increase = seq(0, 3, length.out=10)
# Method 1
fl_EV_A_with_c_increments = 
  log(exp(ffi_crra(CA11*(1+ar_increase))) + 
        exp(ffi_crra(CA12*(1+ar_increase)))) + 
  log(exp(ffi_crra(CA21*(1+ar_increase))) + 
        exp(ffi_crra(CA22*(1+ar_increase))))
print(fl_EV_A_with_c_increments)
# Method 2
fl_EV_A_with_c_increments_m2 = 
  log(exp((1+ar_increase)^(1-gamma)*ffi_crra(CA11)) + 
        exp((1+ar_increase)^(1-gamma)*ffi_crra(CA12))) + 
  log(exp((1+ar_increase)^(1-gamma)*ffi_crra(CA21)) + 
        exp((1+ar_increase)^(1-gamma)*ffi_crra(CA22)))
print(fl_EV_A_with_c_increments_m2)
# Method 3
fl_EV_A_with_c_increments_m3 = 
  log(exp((1+ar_increase)^(1-gamma))*exp(ffi_crra(CA11)) + 
        exp((1+ar_increase)^(1-gamma))*exp(ffi_crra(CA12))) + 
  log(exp((1+ar_increase)^(1-gamma))*exp(ffi_crra(CA21)) + 
        exp((1+ar_increase)^(1-gamma))*exp(ffi_crra(CA22)))
print(fl_EV_A_with_c_increments_m3)
# Method 3
fl_EV_A_with_c_increments_m3 = 
  (1+ar_increase)^(1-gamma) + 
  log(exp(ffi_crra(CA11)) + exp(ffi_crra(CA12))) + 
  (1+ar_increase)^(1-gamma) + 
  log(exp(ffi_crra(CA21)) + exp(ffi_crra(CA22)))
print(fl_EV_A_with_c_increments_m3)
```

#### Analytically Show 

What is analytically the CEV with CRRA + Mlogit?

$$
\log\left(
  \exp\left(U\left(C_{11}\right)\right)
  +\exp\left(U\left(C_{12}\right)\right)
\right)
+ 
\log\left(
  \exp\left(U\left(C_{21}\right)\right)
  +\exp\left(U\left(C_{22}\right)\right)
\right)
$$

Given CRRA Utility

$$
U(C;\psi) = \frac{\left(c\cdot\left(1+\psi\right)\right)^{1-\gamma}}{1-\gamma}\\
U(C;\psi) = \frac{\left(1+\psi\right)^{1-\gamma}\left(c\right)^{1-\gamma}}{1-\gamma} = 
\left(1+\psi\right)^{1-\gamma}\cdot U(C)
$$

Plugging $\psi$ into the equation 

$$
\log\left(
  \exp\left(\left(1+\psi\right)^{1-\gamma}\cdot U\left(C_{11}\right)\right)
  +\exp\left(\left(1+\psi\right)^{1-\gamma}\cdot U\left(C_{12}\right)\right)
\right)
+ 
\log\left(
  \exp\left(\left(1+\psi\right)^{1-\gamma}\cdot U\left(C_{21}\right)\right)
  +\exp\left(\left(1+\psi\right)^{1-\gamma}\cdot U\left(C_{22}\right)\right)
\right)
$$
This is as far as we can easily go, note:

$$
\exp(a\cdot b) \neq \exp(a)\cdot\exp(b)
$$


