---
title: "R String Arrays"
titleshort: "String Operations"
description: |
  Split, concatenate, subset, replace, and substring strings.
  Convert number to string without decimal and negative sign.
  Concatenate numeric and string arrays as a single string.
core:
  - package: r
    code: |
      paste0()
      paste0(round(runif(3),3), collapse=',')
      sub()
      gsub()
      grepl()
      sprintf()
date: 2022-07-13
date_start: 2020-04-11
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### String Arrays

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Positive or Negative Floating Number to String

There is a number, that contains decimal and possibly negative sign and has some decimals, convert this to a string that is more easily used as a file or folder name. 

```{r}
ls_fl_rho <- c(1, -1, -1.5 -100, 0.5, 0.11111111, -199.22123)
for (fl_rho in ls_fl_rho) {
  st_rho <- paste0(round(fl_rho, 4))
  st_rho <- gsub(x = st_rho,  pattern = "-", replacement = "n")
  st_rho <- gsub(x = st_rho,  pattern = "\\.", replacement = "p")
  print(paste0('st_rho=', st_rho))
}
```

#### String Replace

- r string wildcard replace between regex
- [R - replace part of a string using wildcards](https://stackoverflow.com/a/27246520/8280804)

```{r, amto.array.fs_ary_string.replace, eval=FALSE}
# String replacement
gsub(x = paste0(unique(df.slds.stats.perc$it.inner.counter), ':',
                unique(df.slds.stats.perc$z_n_a_n), collapse = ';'),
     pattern = "\n",
     replacement = "")
gsub(x = var,  pattern = "\n", replacement = "")
gsub(x = var.input,  pattern = "\\.", replacement = "_")
```

String replaces a segment, search by wildcard. Given the string below, delete all text between carriage return and pound sign:

```{r}
st_tex_text <- "\n% Lat2ex Comments\n\\newcommand{\\exa}{\\text{from external file: } \\alpha + \\beta}\n% More LaLatex Comments\n"
st_clean_a1 <- gsub("\\%.*?\\\n", "", st_tex_text)
st_clean_a2 <- gsub("L.*?x", "[LATEX]", st_tex_text)
print(paste0('st_tex_text:', st_tex_text))
print(paste0('st_clean_a1:', st_clean_a1))
print(paste0('st_clean_a2:', st_clean_a2))
```

String delete after a particular string:

```{r}
st_tex_text <- "\\end{equation}\n}\n% Even more comments from Latex preamble"
st_clean_a1 <- gsub("\\\n%.*","", st_tex_text)
print(paste0('st_tex_text:', st_tex_text))
print(paste0('st_clean_a1:', st_clean_a1))
```


#### Search If and Which String Contains

- r if string contains
- r if string contains either or grepl
- [Use grepl to search either of multiple substrings in a text](https://stackoverflow.com/a/26319765/8280804)

Search for a single substring in a single string:

```{r support dtype string if contains}
st_example_a <- 'C:/Users/fan/R4Econ/amto/tibble/fs_tib_basics.Rmd'
st_example_b <- 'C:/Users/fan/R4Econ/amto/tibble/_main.html'
grepl('_main', st_example_a)
grepl('_main', st_example_b)
```

Search for if one of a set of substring exists in a set of strings. In particular which one of the elements of *ls_spn* contains at least one of the elements of *ls_str_if_contains*. In the example below, only the first path does not contain either the word *aggregate* or *index* in the path. This can be used after all paths have been found recursively in some folder to select only desired paths from the full set of possibilities:

```{r}
ls_spn <- c("C:/Users/fan/R4Econ//panel/basic/fs_genpanel.Rmd",
            "C:/Users/fan/R4Econ//summarize/aggregate/main.Rmd",
            "C:/Users/fan/R4Econ//summarize/index/fs_index_populate.Rmd")
ls_str_if_contains <- c("aggregate", "index")
str_if_contains <- paste(ls_str_if_contains, collapse = "|")
grepl(str_if_contains, ls_spn)
```

#### String Split

Given some string, generated for example by cut, get the lower cut starting points, and also the higher end point

```{r}
# Extract 0.216 and 0.500 as lower and upper bounds
st_cut_cate <- '(0.216,0.500]'
# Extract Lower Part
substring(strsplit(st_cut_cate, ",")[[1]][1], 2)
# Extract second part except final bracket Option 1
intToUtf8(rev(utf8ToInt(substring(intToUtf8(rev(utf8ToInt(strsplit(st_cut_cate, ",")[[1]][2]))), 2))))
# Extract second part except final bracket Option 2
gsub(strsplit(st_cut_cate, ",")[[1]][2],  pattern = "]", replacement = "")
```

Make a part of a string bold. Go from "ABC EFG, OPQ, RST" to "**ABC EFG**, OPQ, RST". This could be for making the name of an author bold, and preserve affiliation information. 

```{r}
st_paper_author_ori <- "ABC EFG, OPQ, RST"
ar_st_ori <- strsplit(st_paper_author_ori, ", ")[[1]]
st_after_1stcomma <- paste0(ar_st_ori[2:length(ar_st_ori)], collapse= ", ")
st_paper_author <- paste0('<b>', ar_st_ori[1], "</b>, ", st_after_1stcomma )
print(st_paper_author)
```            

#### String Concatenate

Concatenate string array into a single string. 

```{r amto.array.fs_ary_string.concatenate}
# Simple Collapse
vars.group.by <- c('abc', 'efg')
paste0(vars.group.by, collapse='|')
```

Concatenate a numeric array into a single string. 

```{r amto.array.fs_ary_string.concatenate}
# Simple Collapse
set.seed(123)
ar_fl_numbers <- runif(5)
paste0('ar_fl_numbers = ', 
       paste(round(ar_fl_numbers,3), collapse=', ')
)
```

#### String Add Leading Zero

```{r amto.array.fs_ary_string.leadingzero}
# Add Leading zero for integer values to allow for sorting when
# integers are combined into strings
it_z_n <- 1
it_a_n <- 192
print(sprintf("%02d", it_z_n))
print(sprintf("%04d", it_a_n))
```

#### Substring Components

Given a string, with certain structure, get components.

- r time string get month and year and day

```{r}
snm_full <- "20100701"
snm_year <-substr(snm_full,0,4)
snm_month <-substr(snm_full,5,6)
snm_day <-substr(snm_full,7,8)
print(paste0('full:', snm_full,
             ', year:', snm_year,
             ', month:', snm_month,
             ', day:', snm_day))
```
