---
title: "Arrays Operations in R"
description: |
  Basic array operations in R, rep, head, tail, na, etc.
  E notation.
core:
  - package: r
    code: |
      rep()
      head()
      tail()
      na_if()
date: 2020-05-27
date_start: 2020-04-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Array Basics

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Multidimesional Arrays

##### Repeat one Number by the Size of an Array

```{r}
ar_a <- c(1,2,3)
ar_b <- c(1,2,3/1,2,3)
rep(0, length(ar_a))
```

##### Generate 2 Dimensional Array

```{r amto.array.fs_array_basics.multi.2d}
# Multidimensional Array
# 1 is r1c1t1, 1.5 in r2c1t1, 0 in r1c2t1, etc.
# Three dimensions, row first, column second, and tensor third
x <- array(c(1, 1.5, 0, 2, 0, 4, 0, 3), dim=c(2, 2, 2))
dim(x)
print(x)
```

#### Array Slicing

##### Remove Elements of Array

Select elements with direct indexing, or with head and tail functions. Get the first two elements of three elements array.

```{r amto.array.fs_array_basics.slice.lastelement}
# Remove last element of array
vars.group.bydf <- c('23','dfa', 'wer')
vars.group.bydf[-length(vars.group.bydf)]
# Use the head function to remove last element
head(vars.group.bydf, -1)
head(vars.group.bydf, 2)
```

Get last two elements of array.

```{r}
# Remove first element of array
vars.group.bydf <- c('23','dfa', 'wer')
vars.group.bydf[2:length(vars.group.bydf)]
# Use Tail function
tail(vars.group.bydf, -1)
tail(vars.group.bydf, 2)
```

#### NA in Array

##### Check if NA is in Array

```{r amto.array.fs_array_basics.NA.check}
# Convert Inf and -Inf to NA
x <- c(1, -1, Inf, 10, -Inf)
na_if(na_if(x, -Inf), Inf)
```

#### Notations

##### e notation

1. Case one: *1.149946e+00*
    + this is approximately: 1.14995
2. Case two: *9.048038e-01*
    + this is approximately: 0.90480
3. Case three: *9.048038e-01*
    + this is approximately: 0.90480
