---
title: "R String Arrays"
titleshort: "String Operations"
description: |
  Split, concatenate, subset, replace, substring strings
core:
  - package: r
    code: |
      paste0()
      sub()
      gsub()
      grepl()
      sprintf()
      tail()
      strsplit()
      basename()
      dirname()
      substring()
date: 2020-08-21
date_start: 2020-04-11
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### String Arrays

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### String Replace

- r string wildcard replace between regex
- [R - replace part of a string using wildcards](https://stackoverflow.com/a/27246520/8280804)

```{r, amto.array.fs_ary_string.replace, eval=FALSE}
# String replacement
gsub(x = paste0(unique(df.slds.stats.perc$it.inner.counter), ':',
                unique(df.slds.stats.perc$z_n_a_n), collapse = ';'),
     pattern = "\n",
     replacement = "")
gsub(x = var,  pattern = "\n", replacement = "")
gsub(x = var.input,  pattern = "\\.", replacement = "_")
```

String replaces a segment, search by wildcard. Given the string below, delete all text between carriage return and pound sign:

```{r}
st_tex_text <- "\n% Lat2ex Comments\n\\newcommand{\\exa}{\\text{from external file: } \\alpha + \\beta}\n% More LaLatex Comments\n"
st_clean_a1 <- gsub("\\%.*?\\\n", "", st_tex_text)
st_clean_a2 <- gsub("L.*?x", "[LATEX]", st_tex_text)
print(paste0('st_tex_text:', st_tex_text))
print(paste0('st_clean_a1:', st_clean_a1))
print(paste0('st_clean_a2:', st_clean_a2))
```

String delete after a particular string:

```{r}
st_tex_text <- "\\end{equation}\n}\n% Even more comments from Latex preamble"
st_clean_a1 <- gsub("\\\n%.*","", st_tex_text)
print(paste0('st_tex_text:', st_tex_text))
print(paste0('st_clean_a1:', st_clean_a1))
```


##### Search If and Which String Contains

- r if string contains
- r if string contains either or grepl
- [Use grepl to search either of multiple substrings in a text](https://stackoverflow.com/a/26319765/8280804)

Search for a single substring in a single string:

```{r support dtype string if contains}
st_example_a <- 'C:/Users/fan/R4Econ/amto/tibble/fs_tib_basics.Rmd'
st_example_b <- 'C:/Users/fan/R4Econ/amto/tibble/_main.html'
grepl('_main', st_example_a)
grepl('_main', st_example_b)
```

Search for if one of a set of substring exists in a set of strings. In particular which one of the elements of *ls_spn* contains at least one of the elements of *ls_str_if_contains*. In the example below, only the first path does not contain either the word *aggregate* or *index* in the path. This can be used after all paths have been found recursively in some folder to select only desired paths from the full set of possibilities:

```{r}
ls_spn <- c("C:/Users/fan/R4Econ//panel/basic/fs_genpanel.Rmd",
            "C:/Users/fan/R4Econ//summarize/aggregate/main.Rmd",
            "C:/Users/fan/R4Econ//summarize/index/fs_index_populate.Rmd")
ls_str_if_contains <- c("aggregate", "index")
str_if_contains <- paste(ls_str_if_contains, collapse = "|")
grepl(str_if_contains, ls_spn)
```

#### String Split

Given some string, generated for example by cut, get the lower cut starting points, and also the higher end point

```{r}
# Extract 0.216 and 0.500 as lower and upper bounds
st_cut_cate <- '(0.216,0.500]'
# Extract Lower Part
substring(strsplit(st_cut_cate, ",")[[1]][1], 2)
# Extract second part except final bracket Option 1
intToUtf8(rev(utf8ToInt(substring(intToUtf8(rev(utf8ToInt(strsplit(st_cut_cate, ",")[[1]][2]))), 2))))
# Extract second part except final bracket Option 2
gsub(strsplit(st_cut_cate, ",")[[1]][2],  pattern = "]", replacement = "")
```

#### String Concatenate

```{r amto.array.fs_ary_string.concatenate}
# Simple Collapse
vars.group.by <- c('abc', 'efg')
paste0(vars.group.by, collapse='|')
```

#### String Add Leading Zero

```{r amto.array.fs_ary_string.leadingzero}
# Add Leading zero for integer values to allow for sorting when
# integers are combined into strings
it_z_n <- 1
it_a_n <- 192
print(sprintf("%02d", it_z_n))
print(sprintf("%04d", it_a_n))
```

#### Substring Components

Given a string, with certain structure, get components.

- r time string get month and year and day

```{r}
snm_full <- "20100701"
snm_year <-substr(snm_full,0,4)
snm_month <-substr(snm_full,5,6)
snm_day <-substr(snm_full,7,8)
print(paste0('full:', snm_full, 
             ', year:', snm_year, 
             ', month:', snm_month,
             ', day:', snm_day))
```

#### Substring and File Name

From path, get file name without suffix.

- r string split
- [r list last element](https://stackoverflow.com/a/83222/8280804)
- [r get file name from path](https://stackoverflow.com/a/29114007/8280804)
- [r get file path no name](https://stackoverflow.com/a/47189541/8280804)

```{r}
st_example <- 'C:/Users/fan/R4Econ/amto/tibble/fs_tib_basics.Rmd'
st_file_wth_suffix_s <- tail(strsplit(st_example, "/")[[1]],n=1)
st_file_wno_suffix_s <- tools::file_path_sans_ext(basename(st_example))
st_fullpath_nosufx_s <- sub('\\.Rmd$', '', st_example)
st_fullpath_noname_s <- dirname(st_example)

print(strsplit(st_example, "/"))
print(paste0('st_file_wth_suffix_s:', st_file_wth_suffix_s))
print(paste0('st_file_wno_suffix_s:', st_file_wno_suffix_s))
print(paste0('st_fullpath_nosufx_s:', st_fullpath_nosufx_s))
print(paste0('st_fullpath_noname_s:', st_fullpath_noname_s))
```
