---
title: "R DPLYR Group and sort, and Slice and Summarize"
titleshort: "Group and sort, and Slice and Summarize"
description: |
  Group a dataframe by a variable, sort within group by another variable, keep only highest rows.
core:
  - package: dplyr
    code: |
      arrange()
      group_by()
      slice_head(n=1)
date: 2020-12-17
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Group, Sort and Slice

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Get Highest Values from Groups

There is a dataframe with a grouping variable. Get N rows that have the highest sorted value for another numeric variable. In the example below, group by *cyl* and get the cars with the lowest *mpg* in each *cyl* group.

Show all values.

```{r}
kable(mtcars %>% arrange(cyl, mpg)) %>% kable_styling_fc()
```

Three groups min mpg each group: 

```{r}
# use mtcars: slice_head gets the lowest sorted value
df_groupby_top_mpg <- mtcars %>%
  arrange(cyl, mpg) %>%
  group_by(cyl) %>%
  slice_head(n=1) %>%
  select(cyl, mpg)

# display
kable(df_groupby_top_mpg) %>% kable_styling_fc()
```
