---
title: "R Within-group Ascending and Descending Sort, Selection, and Differencing"
titleshort: "R Within-group Ascending and Descending Sort, Selection, and Differencing"
description: |
  Sort a dataframe by multiple variables, some in descending order.
  Select observations with the highest M values from within N groups (top scoring students from each class).
core:
  - package: dplyr
    code: |
      arrange(a, b, desc(c))
      group_by()
      lag()
      lead()
      slice_head(n=1)
date: 2022-07-15
date_start: 2020-12-17
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Group, Sort and Slice

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Sort in Ascending and Descending Orders

We sort the mtcars dataset, sorting in *ascending* order by cyl, and in *descending* order by mpg. Using *arrange*, *desc(disp)* means sorting the disp variable in descending order. In the table shown below, *cyc* is increasing, and *disp* id decreasing within each *cyc* group.

```{r}
kable(mtcars %>%
  arrange(cyl, desc(disp)) %>%
    # Select and filter to reduce display clutter
    select(cyl, disp, mpg)) %>%
  kable_styling_fc()
```

#### Get Highest Values from Groups

There is a dataframe with a grouping variable with N unique values, for example N classes. Find the top three highest scoring students from each class. In the example below, group by *cyl* and get the cars with the highest and second highest *mpg* cars in each *cyl* group.

```{r}
# use mtcars: slice_head gets the lowest sorted value
df_groupby_top_mpg <- mtcars %>%
  rownames_to_column(var = "car") %>%
  arrange(cyl, desc(mpg)) %>%
  group_by(cyl) %>%
  slice_head(n=3) %>%
  select(car, cyl, mpg, disp, hp)

# display
kable(df_groupby_top_mpg) %>% kable_styling_fc()
```

#### Differences in Within-group Sorted Value

We first take the largest N values in M groups, then we difference between the ranked top values in each group.

We have N classes, and M students in each class. We first select the 3 students with the highest scores from each class, then we take the difference between 1st and 2nd, and the difference between the 2nd and the 3rd students.

Note that when are using descending sort, so *lead* means the next value in descending sequencing, and *lag* means the last value which was higher in descending order.

```{r}
# We use what we just created in the last block.
df_groupby_top_mpg_diff <- df_groupby_top_mpg %>%
  group_by(cyl) %>%
  mutate(mpg_diff_higher_minus_lower = mpg - lead(mpg)) %>%
  mutate(mpg_diff_lower_minus_higher = mpg - lag(mpg))

# display
kable(df_groupby_top_mpg_diff)  %>% kable_styling_fc()
```
