---
title: "R Example DPLYR Generated Sorted Index and Expand Value from an Index to All Rows"
output:
  pdf_document: default
  word_document: default
  html_document: default
  html_notebook: default
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/CodeDynaAsset/)'s [R4Econ](https://fanwangecon.github.io/R4Econ/) Repository or [Intro Stats with R](https://fanwangecon.github.io/Stat4Econ/) Repository.

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```

```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(tidyverse)
library(knitr)
library(kableExtra)
library(R4Econ)
# file name
st_file_name = 'fs_index_populate'
# Generate R File
purl(paste0(st_file_name, ".Rmd"), output=paste0(st_file_name, ".R"), documentation = 2)
# Generate PDF and HTML
# rmarkdown::render("C:/Users/fan/R4Econ/summarize/index/fs_index_populate.Rmd", "pdf_document")
# rmarkdown::render("C:/Users/fan/R4Econ/summarize/index/fs_index_populate.Rmd", "html_document")
```

# Generate Sorted Index within Group and Spread

## Generate Sorted Index within Group with Repeating Values

There is a variable, sort by this variable, then generate index from 1 to N representing sorted values of this index. If there are repeating values, still assign index, different index each value. 

- r generate index sort
- dplyr mutate equals index

```{r}
# dataset subsetting
df_iris <- iris %>% arrange(Sepal.Length) %>% 
              mutate(Sepal.Length.Index = row_number()) %>%
              select(Sepal.Length, Sepal.Length.Index, everything())

# Show results Head 10
df_iris %>% head(10) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

## Populate Value from Lowest Index to All other Rows

We would like to calculate for example the ratio of each individual's highest to the the person with the lowest height in a dataset. We first need to generated sorted index from lowest to highest, and then populate the lowest height to all rows, and then divide. 

- r spread value to all rows from one row
- see also: [Aggregating and analyzing data with dplyr](https://datacarpentry.org/dc_zurich/R-ecology/04-dplyr.html)

```{r}
# In the df_hgt_wgt from R4Econ, there is a country id, village id,
# and individual id, and various other statistics
vars.group <- c('S.country', 'vil.id', 'indi.id')
vars.values <- c('hgt', 'momEdu')

# dataset subsetting
df_use <- df_hgt_wgt %>% select(!!!syms(c(vars.group, vars.values)))

# Group, count and generate means for each numeric variables
df.group <- df_use %>% group_by(!!!syms(vars.group)) %>%
            arrange(!!!syms(vars.group)) %>%
            summarise_if(is.numeric,
                         funs(mean = mean(., na.rm = TRUE),
                              sd = sd(., na.rm = TRUE),
                              n = sum(is.na(.)==0)))

# Show results Head 10
df.group %>% head(10) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
# Show results Head 10
df.group %>% tail(10) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
