---
title: "DPLYR Evaluate Function with N Arrays of Inputs J times"
output:
  html_document: default
  word_document: default
  pdf_document: default
  html_notebook: default
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/CodeDynaAsset/)'s [R4Econ](https://fanwangecon.github.io/R4Econ/) Repository or [Intro Stats with R](https://fanwangecon.github.io/Stat4Econ/) Repository.

## Issue and Goal

We want evaluate nonlinear function f(Q_i, y_i, ar_x, ar_y, c_j, d), where c_j is an element of some array, d is a constant, and ar_x and ar_y are arrays, both fixed. x_i and y_i vary over each row of matrix. We would like to evaluate this nonlinear function concurrently across $N$ individuals. The eventual goal is to find the $i$ specific $Q$ that solves the nonlinear equations.

In [Evaluate Nonlinear Function with N arrays of Inputs](https://fanwangecon.github.io/R4Econ/support/function/fs_funceval.html), we evaluated the function fixing $c$, the difference here is that we want to evaluate $J$ times the overall matrix of inputs.

The results should be stored in a tibble where each of the $j \in J$ evaluations are stacked together with a variable $c$ that records which $c_j$ value was used for this evaluation.

This is achieved here by using [expand_grid](https://fanwangecon.github.io/R4Econ/support/array/fs_meshr.html).

## Set Up

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(tidyverse)
library(knitr)
library(kableExtra)
# file name
st_file_name = 'fs_funceval'
# Generate R File
purl(paste0(st_file_name, ".Rmd"), output=paste0(st_file_name, ".R"), documentation = 2)
# Generate PDF and HTML
# rmarkdown::render("C:/Users/fan/R4Econ/support/function/fs_funceval.Rmd", "pdf_document")
# rmarkdown::render("C:/Users/fan/R4Econ/support/function/fs_funceval.Rmd", "html_document")
```

## Set up Input Arrays

There is a function that takes $M=Q+P$ inputs, we want to evaluate this function $N$ times. Each time, there are $M$ inputs, where all but $Q$ of the $M$ inputs, meaning $P$ of the $M$ inputs, are the same. In particular, $P=Q*N$.

$$M = Q+P = Q + Q*N$$

```{r setup_data}
# it_child_count = N, the number of children
it_N_child_cnt = 5
# it_heter_param = Q, number of parameters that are heterogeneous across children
it_Q_hetpa_cnt = 2

# P fixed parameters, nN is N dimensional, nP is P dimensional
ar_nN_A = seq(-2, 2, length.out = it_N_child_cnt)
ar_nN_alpha = seq(0.1, 0.9, length.out = it_N_child_cnt)
ar_nP_A_alpha = c(ar_nN_A, ar_nN_alpha)
ar_nN_N_choice = seq(1,it_N_child_cnt)/sum(seq(1,it_N_child_cnt))

# N by Q varying parameters
mt_nN_by_nQ_A_alpha = cbind(ar_nN_A, ar_nN_alpha, ar_nN_N_choice)
```

Now generate a vector of $\rho$, which represents varying planner preference, and mesh it together with the matrix *mt_nN_by_nQ_A_alpha*.

```{r setup_data planner preference and mesh}
# Vector of Planner Preference
ar_rho = c(-0.25, -0.15, 0.15, 0.25)

# N by Q varying parameters but now Mesh with RHO, the J elements we want to vary over
mt_nN_by_nQ_A_alpha_mesh_rho <- as_tibble(mt_nN_by_nQ_A_alpha) %>% expand_grid(rho = ar_rho) %>%
                                  arrange(rho, ar_nN_A, ar_nN_alpha, ar_nN_N_choice)

# Show
kable(mt_nN_by_nQ_A_alpha_mesh_rho) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

And the aggregate resources available for allocations, along with observed allocations. 

```{r setup_data total resource}
# Total Resources available
fl_N_agg = 100

```

## Define Function

```{r define optimal choice function}

# Define Implicit Function
ffi_nonlin_dplyrdo <- function(fl_A, fl_alpha, fl_N, fl_rho, ar_A, ar_alpha, fl_N_agg){

  # Test Parameters
  # ar_A = ar_nN_A
  # ar_alpha = ar_nN_alpha
  # fl_N = 100
  # fl_rho = -1
  # fl_N_q = 10

  # Apply Function
  ar_p1_s1 = exp((fl_A - ar_A)*fl_rho)
  ar_p1_s2 = (fl_alpha/ar_alpha)
  ar_p1_s3 = (1/(ar_alpha*fl_rho - 1))
  ar_p1 = (ar_p1_s1*ar_p1_s2)^ar_p1_s3
  ar_p2 = (fl_N*fl_N_agg)^((fl_alpha*fl_rho-1)/(ar_alpha*fl_rho-1))
  ar_overall = ar_p1*ar_p2
  fl_overall = fl_N_agg - sum(ar_overall)

  return(fl_overall)
}

```

## Evaluate Nonlinear Function using dplyr mutate Once

Below, repeat what we did in [Evaluate Nonlinear Function with N arrays of Inputs](https://fanwangecon.github.io/R4Econ/support/function/fs_funceval.html).

```{r nonlinear_dplyr_mutate}
# Convert Matrix to Tibble
ar_st_col_names = c('fl_A', 'fl_alpha', 'fl_N')
tb_nN_by_nQ_A_alpha <- as_tibble(mt_nN_by_nQ_A_alpha) %>% rename_all(~c(ar_st_col_names))

# fl_A, fl_alpha are from columns of tb_nN_by_nQ_A_alpha
fl_rho_here = ar_rho[1]
tb_nN_by_nQ_A_alpha = tb_nN_by_nQ_A_alpha %>% rowwise() %>%
                        mutate(dplyr_eval = ffi_nonlin_dplyrdo(fl_A, fl_alpha, fl_N, fl_rho_here,
                                                               ar_nN_A, ar_nN_alpha,
                                                               fl_N_agg))
# Show
kable(tb_nN_by_nQ_A_alpha) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


## Evaluate Nonlinear Function using dplyr multiple times and stack results

Evaluate multiple times and stack results. 

```{r nonlinear_dplyr_mutate multiple rhos}
# Convert Matrix to Tibble
ar_st_col_names = c('fl_A', 'fl_alpha', 'fl_N', 'fl_rho')
tb_nN_by_nQ_A_alpha_mesh_rho <- as_tibble(mt_nN_by_nQ_A_alpha_mesh_rho) %>% rename_all(~c(ar_st_col_names))

# fl_A, fl_alpha are from columns of tb_nN_by_nQ_A_alpha
tb_nN_by_nQ_A_alpha_mesh_rho = tb_nN_by_nQ_A_alpha_mesh_rho %>% rowwise() %>%
                        mutate(dplyr_eval = ffi_nonlin_dplyrdo(fl_A, fl_alpha, fl_N, fl_rho,
                                                               ar_nN_A, ar_nN_alpha,
                                                               fl_N_agg))
# Show
kable(tb_nN_by_nQ_A_alpha_mesh_rho) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

