---
title: "R Wide Data to Long Data Example (TIDYR Pivot Longer)"
description: |
  A matrix of ev given states, rows are states and cols are shocks. Convert to Long table with shock and state values and ev.
core:
  - package: dplyr
    code: |
      left_join()
      pivot_longer(cols = starts_with('zi'), names_to = c('zi'), names_pattern = paste0("zi(.)"), values_to = "ev")
date: 2020-05-14
date_start: 2020-05-14
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Wide to Long

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

Using the [pivot_wider](https://tidyr.tidyverse.org/reference/pivot_wider.html) function in tidyr to reshape panel or other data structures

#### Generated Matrix by States to Long Table

A matrix of ev given states, rows are states and cols are shocks. Convert to Long table with shock and state values and ev.

Generated Matrix by States to Long Table where state values are stored as variables, with correct value labels for states:

1. Generate a matrix
2. Convert matrix to tibble
3. Tibble make longer, and store column and row id var names

```{r}
# Generate A Matrix
set.seed(123)
ar_a <- c(1.2, 22.2, 33.33)
ar_z <- c(-2.5,  -1, 0, 1, 2.53)
mt_ev = matrix(rnorm(ar_a*ar_z), nrow=length(ar_a), ncol=length(ar_z))

# Name Matrix
rownames(mt_ev) <- paste0('ai', seq(1:length(ar_a)))
colnames(mt_ev) <- paste0('zi', seq(1:length(ar_z)))

# to tibble
tb_ev <- as_tibble(mt_ev) %>% rowid_to_column(var = "ai")

# longer 
tb_ev_long <- tb_ev %>% 
  pivot_longer(cols = starts_with('zi'), 
               names_to = c('zi'),
               names_pattern = paste0("zi(.)"),
               values_to = "ev") %>%
  mutate(zi = as.numeric(zi))

# Merge with a and z values
tb_ev_long <- tb_ev_long %>% 
  left_join(as_tibble(ar_a) %>% 
              rowid_to_column(var = "ai") %>%
              rename(a = value)
              , by = 'ai') %>% 
  left_join(as_tibble(ar_z) %>% 
              rowid_to_column(var = "zi") %>%
              rename(z = value), 
            by = 'zi') %>%
  select(a,ai,z,zi,ev)

# Display
kable(tb_ev) %>% kable_styling_fc()  
kable(tb_ev_long) %>% kable_styling_fc()  
```
