---
title: "Convert Table from Wide to Long with dplyr"
titleshort: "Convert Table from Wide to Long with dplyr"
description: |
  Given a matrix of values with row and column labels, create a table where the unit of observation are the row and column categories, and the values in the matrix is stored in a single variable.
  Reshape wide to long two sets of variables, two categorical variables added to wide table.
core:
  - package: tidyr
    code: |
      pivot_longer(cols = starts_with('zi'), names_to = c('zi'), names_pattern = paste0("zi(.)"), values_to = "ev")
      pivot_longer(cols = matches('a line b'), names_to = c('va', 'vb'), names_pattern = paste0("(.)_(.)"), values_to = "ev")
  - package: dplyr
    code: |
      left_join()
date: 2023-07-18
date_start: 2020-05-14
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Wide to Long

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

Using the [pivot_wider](https://tidyr.tidyverse.org/reference/pivot_wider.html) function in tidyr to reshape panel or other data structures

#### Wide to long panel, single variable

We have a matrix of values, the values are *ev*. Each row corresponds to a different value of the *a* variable, each column represents a different value of the *z* variable. 

Based on this matrix, we create a table where each unit of observation is for a specific *a* and *z* variable combination. So the matrix is turned from wide to long. 

The resulting long table has 5 variables 

1. *a*: values of the *a* variable, the original matrix row labels 
2. *ai*: an index from 1, indicating the original matrix row index
3. *z*: values of the *z* variable, the original matrix column lables
4. *zi*: an index from 1, indicating hte original matrix column index

First, we create the matrix. 

```{r}
# Generate A Matrix
set.seed(123)
ar_a <- c(1.1,5.1)
ar_z <- seq(-2.5, 2.53, length.out=11)
mt_ev = matrix(rnorm(length(ar_a)*length(ar_z)), 
  nrow=length(ar_a), ncol=length(ar_z))

# Name Matrix
rownames(mt_ev) <- paste0('ai', seq(1:length(ar_a)))
colnames(mt_ev) <- paste0('zi', seq(1:length(ar_z)))

# to tibble
tb_ev <- as_tibble(mt_ev) %>% rowid_to_column(var = "ai")

# Print
print(mt_ev)
# Display
kable(tb_ev, caption = "Wide table") %>% kable_styling_fc()
```

Second, we convert the table wide to long. 

```{r}

# longer
tb_ev_long <- tb_ev %>%
  pivot_longer(cols = starts_with('zi'),
               names_to = c('zi'),
               names_pattern = paste0("zi(.*)"),
               values_to = "ev") %>%
  mutate(zi = as.numeric(zi))

# Merge with a and z values
tb_ev_long <- tb_ev_long %>%
  left_join(as_tibble(ar_a) %>%
              rowid_to_column(var = "ai") %>%
              rename(a = value)
              , by = 'ai') %>%
  left_join(as_tibble(ar_z) %>%
              rowid_to_column(var = "zi") %>%
              rename(z = value),
            by = 'zi') %>%
  select(a,ai,z,zi,ev)

# Display
kable(tb_ev_long, caption = "Long table") %>% kable_styling_fc()
```

#### Wide to long panel, multiple variables

We have a dataset where each row contains data from a different year. We have four variables, observed wage, simulated wage, observed labor quantities, and simulated labor quantities. 

We generate reshape this file to have four variables: 

1. year
2. categorical for wage or quantity
3. categorical for observed or simulated
4. a numerical column with wage and quantity values

This is different then the situation prior, because we are need to convert to long two different numerical variables that will be in the same long variable, but differentiated by two categorical variables (rather than one).

First, we create the matrix. 

```{r}
# Generate A Matrix
set.seed(123)
ar_year <- c(1995, 1997, 1999)
ar_vars <- c("wage_model", "quant_model", "wage_simu", "quant_simu")
mt_equi = matrix(rnorm(length(ar_year)*length(ar_vars)), 
  nrow=length(ar_year), ncol=length(ar_vars))

# Name Matrix
rownames(mt_equi) <- ar_year
colnames(mt_equi) <- ar_vars

# to tibble
tb_equi <- as_tibble(mt_equi, rownames = "year")

# Print
print(mt_equi)
# Display
kable(tb_equi, caption = "Wide table") %>% kable_styling_fc()
```

Second, we convert the table wide to long. We select columns that includes either wage or quant, see tidyselect [Select variables that match a pattern](https://tidyselect.r-lib.org/reference/starts_with.html) for additional verbs for how to select variables. 

```{r}
# longer
tb_equi_long <- tb_equi %>%
  pivot_longer(cols = matches('wage|quant'),
               names_to = c('variable', 'source'),
               names_pattern = paste0("(.*)_(.*)"),
               values_to = "value") 

# Display
kable(tb_equi_long, caption = "Long table, Two Variables") %>% kable_styling_fc()
```

