---
title: "R Summary By Groups, One Variable All Statistics"
titleshort: "By Groups, One Variable All Statistics"
description: |
  Pick stats, overall, and by multiple groups, stats as matrix or wide row with name=(ctsvar + catevar + catelabel).
core:
  - package: tidyr
    code: |
      group_by()
      summarize_at(, funs())
      rename(!!var := !!sym(var))
      mutate(!!var := paste0(var,'str',!!!syms(vars)))
      gather()
      unite()
      spread(varcates, value)
date: 2020-04-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### One Variable Group Summary

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

There is a categorical variable (based on one or the interaction of multiple variables), there is a continuous variable, obtain statistics for the continuous variable conditional on the categorical variable, but also unconditionally.

Store results in a matrix, but also flatten results wide to row with appropriate keys/variable-names for all group statistics.

Pick which statistics to be included in final wide row

#### Build Program

```{r}
# Single Variable Group Statistics (also generate overall statistics)
ff_summ_by_group_summ_one <- function(
  df, vars.group, var.numeric, str.stats.group = 'main',
  str.stats.specify = NULL, boo.overall.stats = TRUE){

    # List of statistics
    # https://rdrr.io/cran/dplyr/man/summarise.html
    strs.center <- c('mean', 'median')
    strs.spread <- c('sd', 'IQR', 'mad')
    strs.range <- c('min', 'max')
    strs.pos <- c('first', 'last')
    strs.count <- c('n_distinct')

    # Grouping of Statistics
    if (missing(str.stats.specify)) {
        if (str.stats.group == 'main') {
            strs.all <- c('mean', 'min', 'max', 'sd')
        }
        if (str.stats.group == 'all') {
            strs.all <- c(strs.center, strs.spread, strs.range, strs.pos, strs.count)
        }
    } else {
        strs.all <- str.stats.specify
    }

    # Start Transform
    df <- df %>% drop_na() %>% mutate(!!(var.numeric) := as.numeric(!!sym(var.numeric)))

    # Overall Statistics
    if (boo.overall.stats) {
        df.overall.stats <- df %>% summarize_at(vars(var.numeric), funs(!!!strs.all))
        if (length(strs.all) == 1) {
            # give it a name, otherwise if only one stat, name of stat not saved
            df.overall.stats <- df.overall.stats %>% rename(!!strs.all := !!sym(var.numeric))
        }
        names(df.overall.stats) <- paste0(var.numeric, '.', names(df.overall.stats))
    }

    # Group Sort
    df.select <- df %>%
                  group_by(!!!syms(vars.group)) %>%
                  arrange(!!!syms(c(vars.group, var.numeric)))

    # Table of Statistics
    df.table.grp.stats <- df.select %>% summarize_at(vars(var.numeric), funs(!!!strs.all))

    # Add Stat Name
    if (length(strs.all) == 1) {
        # give it a name, otherwise if only one stat, name of stat not saved
        df.table.grp.stats <- df.table.grp.stats %>% rename(!!strs.all := !!sym(var.numeric))
    }


    # Row of Statistics
    str.vars.group.combine <- paste0(vars.group, collapse='_')
    if (length(vars.group) == 1) {
        df.row.grp.stats <- df.table.grp.stats %>%
                mutate(!!(str.vars.group.combine) := paste0(var.numeric, '.',
                                               vars.group, '.g',
                                               (!!!syms(vars.group)))) %>%
                gather(variable, value, -one_of(vars.group)) %>%
                unite(str.vars.group.combine, c(str.vars.group.combine, 'variable')) %>%
                spread(str.vars.group.combine, value)
    } else {
        df.row.grp.stats <- df.table.grp.stats %>% 
          mutate(vars.groups.combine := paste0(paste0(vars.group, collapse='.')),
                 !!(str.vars.group.combine) := paste0(interaction(!!!(syms(vars.group))))) %>%
          mutate(!!(str.vars.group.combine) := paste0(var.numeric, '.', vars.groups.combine, '.',
                                                      (!!sym(str.vars.group.combine)))) %>%
          ungroup() %>%
          select(-vars.groups.combine, -one_of(vars.group)) %>%
          gather(variable, value, -one_of(str.vars.group.combine)) %>%
          unite(str.vars.group.combine, c(str.vars.group.combine, 'variable')) %>%
          spread(str.vars.group.combine, value)
    }

    # Clean up name strings
    names(df.table.grp.stats) <- 
      gsub(x = names(df.table.grp.stats),pattern = "_", replacement = "\\.")
    names(df.row.grp.stats) <- 
      gsub(x = names(df.row.grp.stats),pattern = "_", replacement = "\\.")

    # Return
    list.return <- 
      list(df_table_grp_stats = df.table.grp.stats, df_row_grp_stats = df.row.grp.stats)

    # Overall Statistics, without grouping
    if (boo.overall.stats) {
        df.row.stats.all <- c(df.row.grp.stats, df.overall.stats)
        list.return <- append(list.return, list(df_overall_stats = df.overall.stats,
                                                df_row_stats_all = df.row.stats.all))
    }

    # Return
    return(list.return)

}
```

#### Test

Load data and test

```{r}
# Library
library(tidyverse)

# Load Sample Data
setwd('C:/Users/fan/R4Econ/_data/')
df <- read_csv('height_weight.csv')
```

##### Function Testing By Gender Groups

Need two variables, a group variable that is a factor, and a numeric

```{r}
vars.group <- 'sex'
var.numeric <- 'hgt'
```

```{r}
df.select <- df %>% select(one_of(vars.group, var.numeric)) %>% drop_na()
```

Main Statistics:

```{r}
# Single Variable Group Statistics
ff_summ_by_group_summ_one(
  df.select, vars.group = vars.group, var.numeric = var.numeric, 
  str.stats.group = 'main')
```

Specify Two Specific Statistics:

```{r}
ff_summ_by_group_summ_one(
  df.select, vars.group = vars.group, var.numeric = var.numeric, 
  str.stats.specify = c('mean', 'sd'))
```

Specify One Specific Statistics:

```{r}
ff_summ_by_group_summ_one(
  df.select, vars.group = vars.group, var.numeric = var.numeric, 
  str.stats.specify = c('mean'))
```

##### Function Testing By Country and Gender Groups

Need two variables, a group variable that is a factor, and a numeric. Now joint grouping variables.

```{r}
vars.group <- c('S.country', 'sex')
var.numeric <- 'hgt'
```

```{r}
df.select <- df %>% select(one_of(vars.group, var.numeric)) %>% drop_na()
```

Main Statistics:

```{r}
ff_summ_by_group_summ_one(
  df.select, vars.group = vars.group, var.numeric = var.numeric, 
  str.stats.group = 'main')
```

Specify Two Specific Statistics:

```{r}
ff_summ_by_group_summ_one(
  df.select, vars.group = vars.group, var.numeric = var.numeric, 
  str.stats.specify = c('mean', 'sd'))
```

Specify One Specific Statistics:

```{r}
ff_summ_by_group_summ_one(
  df.select, vars.group = vars.group, var.numeric = var.numeric, str.stats.specify = c('mean'))
```
