---
title: "Handling R Packages"
titleshort: "Handling R Packages"
description: |
  Resolve conflicts between two packages with identically named function.
core:
  - package: tidyverse
    code: |
      tidyverse_conflicts
  - package: dplyr 
    code: |
      filter
  - package: stats 
    code: |
      filter
  - package: conflicted
    code: |
      conflict_prefer()
date: 2024-01-01
date_start: 2024-01-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### R Package Installation

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Duplicate function names across packages

[dplyr::filter](https://dplyr.tidyverse.org/reference/filter.html) and [stats::filter](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/filter) are two functions from two popular packages that have the same name. And this leads to erros, when `dplyr::filter` is confused for `stats::filter`. We use the [conflicted::conflict_prefer](https://conflicted.r-lib.org/reference/conflict_prefer.html) to resolve this issue. 

This is an issue related to [namespaces](https://rlang.r-lib.org/reference/ns_env.html).

Below return the environment on the search path via [rlang::search_envs()](https://rlang.r-lib.org/reference/search_envs.html):

```{r}
print(rlang::search_envs())
```

We can use [tidyverse_conflicts()](https://tidyverse.tidyverse.org/reference/tidyverse_conflicts.html) to "lists all the conflicts between packages in the tidyverse and other packages that you have loaded". We can see that we have problems due to `filter`, `lag`, and `group_rows`.

```{r}
tidyverse_conflicts()
```

For example, the code below fails: 

```{r, eval=FALSE}
library(stats)
library(dplyr)
as_tibble(mtcars, rownames = "car") %>% filter(car == "Valiant")

# Error message
# > as_tibble(mtcars, rownames = "car") %>% filter(car == "Valiant")
# Error: object 'car' not found
```

The code below works, because we explicitly write `dplyr::filter`:

```{r}
library(stats)
library(dplyr)
print(as_tibble(mtcars, rownames = "car") %>% dplyr::filter(car == "Valiant"))
```

To deal with this, we use the [conflicted::conflict_prefer](https://conflicted.r-lib.org/reference/conflict_prefer.html) to resolve this issue. Now we can run the `filter` function safely, knowing that it is the `dplyr::filter` function will be used.

```{r, eval = FALSE}
library(conflicted)
conflict_prefer("filter", "dplyr", "stats")
library(stats)
library(dplyr)
print(as_tibble(mtcars, rownames = "car") %>% filter(car == "Valiant"))

# > conflict_prefer("filter", "dplyr", "stats")
# [conflicted] Will prefer dplyr::filter over stats::filter.
# > library(stats)
# > library(dplyr)
# > print(as_tibble(mtcars, rownames = "car") %>% filter(car == "Valiant"))
# # A tibble: 1 × 12
#   car       mpg   cyl  disp    hp  drat    wt  qsec    vs    am  gear  carb
#   <chr>   <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
# 1 Valiant  18.1     6   225   105  2.76  3.46  20.2     1     0     3     1
```
