---
title: "Summarize statistics across countries grouped by regions using kableExtra"
description: |
  Summarize statistics across countries grouped by regions using kableExtra
core:
  - package: kableExtra
    code: |
      ggplot()
      geom_line()
date: 2024-04-07
date_start: 2024-04-07
output:
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
always_allow_html: true
urlcolor: blue
---

### Summarize statistics from across countries

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Generating preparing data input

In this example file, we have the average exposure to air pollution among children from Latin America and the Caribbean in 2010, as well as average temperature, measured in share of hours in the year over UTCI 32 degrees. We will create a table that groups Latin American and Caribbean countries into groupings, show various statistics by each country row by row, including ranking for each country. The output generated is in HTML and tex formats.

Files of this nature generates all the tables in the appendix of [Hannum, Kim, and Wang (2024)](https://jeonghyeok-kim.github.io/assets/GlobalChildrenTeachersSchools_HannumKimWang.pdf). It is implemented also [here](https://github.com/ClimateInequality/PrjCEC/blob/main/R-script/tabfig_1_mean_child/ffs_pirecec_tf_mean_child_season_tab_b.R) for appendix tables in [Feng et al (2024)](https://szkaifeng.github.io/pdf/FengLBHWChinaChildrenHeat1990t2020.pdf).

First, we load the pollution file in. Additionally, we load in a file containing the full names of countries. Clean up the input files so that the country files have the same variable names. Ignore the prior variable names, with the prefix pm10, they are not pm10 measurements.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(tibble)
library(kableExtra)
```

```{r}
# Parameters
verbose <- TRUE
# Generate paths
spt_path_root <- file.path('C:', 'users', 'fan', 'R4Econ', fsep = .Platform$file.sep)
spt_path_in <- file.path(spt_path_root, 'table', 'bigtable', '_file', fsep = .Platform$file.sep)
spn_aod <- file.path(spt_path_in, 'lac_aod_children.csv', fsep = .Platform$file.sep)
spn_utci <- file.path(spt_path_in, 'lac_utci_children.csv', fsep = .Platform$file.sep)
spn_keys <- file.path(spt_path_in, 'country_code.csv', fsep = .Platform$file.sep)
# Load files
df_lac_aod_pollution <- read_csv(spn_aod) %>% 
  select(-order)
df_lac_utci_temperature <- read_csv(spn_utci) %>% 
  select(
    one_of(colnames(df_lac_aod_pollution)),
    pm10_grp_mean, pm10_overall_mean, pm10_grp_exc_burden
  ) %>% 
  rename(
    pm10_grp_mean_x = pm10_grp_mean, 
    pm10_grp_mean_y = pm10_overall_mean,
    relative_excess_burden = pm10_grp_exc_burden
  )
df_country_code <- read_csv(spn_keys)
# output tex file path
spn_tex_out <- file.path(
  spt_path_in, 'lac_climate_children_rank.tex', fsep = .Platform$file.sep)
```

Second, we merge the country names into the data file with three letter ISO code, select the subset of variables to be used by the table, and generate some summary stats on the file. We also merge the UTCI temperature information into the pollution file. 

```{r}
# Merge with key
df_lac_aod_pollution <- df_lac_aod_pollution %>% 
  left_join(df_country_code, by = "ISOCODE") 
# Merge with temperature information
df_lac_aod_pollution <- df_lac_aod_pollution %>% 
  left_join(df_lac_utci_temperature %>% 
    rename(utcige32c_age0t5 = pm10_grp_mean_x) %>%
    select(ISOCODE, utcige32c_age0t5), by = "ISOCODE") 
# Filter region to consider 
df_lac_aod_pollution_lac <- df_lac_aod_pollution %>%
  dplyr::filter(region_name == "Americas") %>%
  dplyr::filter(!(ISOCODE %in% c(
    "CAN", "USA", "CUB", "GRL"
    )))
# Select variables
df_lac_aod_pollution_lac <- df_lac_aod_pollution_lac %>% 
  rename(
    group_sorter_desc = 'Income group',
    aodavg_age0t5 = pm10_grp_mean_x
    ) %>%
  select(
    ISOCODE, country_name, group_sorter_desc, 
    aodavg_age0t5 , utcige32c_age0t5
  )
# Display 
str(df_lac_aod_pollution_lac)
```

Third, for a number of locations that are dependencies and constituent entities, no full country names were included, we include them now. 

```{r}
# Identify countries without country_name
ar_st_isocode_dependencies <- df_lac_aod_pollution_lac %>% 
  dplyr::filter(is.na(country_name)) %>% 
  pull(ISOCODE)
print(ar_st_isocode_dependencies)
# Add in names
# https://www.britannica.com/topic/list-of-countries-in-Latin-America-2061416
df_lac_aod_pollution_lac <- df_lac_aod_pollution_lac %>% 
  mutate(country_name = case_when(
    ISOCODE == "MTQ" ~ 'Martinique',
    ISOCODE == "MSR" ~ 'Montserrat',
    ISOCODE == "AIA" ~ 'Anguilla',
    ISOCODE == "BES" ~ 'St. Eustatius',
    ISOCODE == "GUF" ~ 'French Guiana',
    ISOCODE == "FLK" ~ 'Falkland Islands',
    TRUE ~ country_name
  ))
# Country name and ISO combine
df_lac_aod_pollution_lac <- df_lac_aod_pollution_lac %>% 
  mutate(country_name = paste0(
    country_name, " (", ISOCODE , ")"
    ))
```

Fourth, we will use the world bank income groups to categorize countries, and include an additional category for the dependency and constituent entities just renamed above. 

```{r}
# Add in categorical name for locations without world bank income group designation
df_lac_aod_pollution_lac <- df_lac_aod_pollution_lac %>%
  mutate(group_sorter_desc= case_when(
    is.na(group_sorter_desc) ~ 'Without World Bank income group designation',
    TRUE ~ group_sorter_desc
  ))
```

Fifth, we generate a ranking variable, for which country has the highest level of AOD, and rescale AOD value to be between 0 and 1.

```{r}
# basic chart with two lines
df_lac_aod_pollution_lac <- df_lac_aod_pollution_lac %>% 
  mutate(aodavg_age0t5 = aodavg_age0t5/1000) %>% 
  arrange(aodavg_age0t5) %>%
  mutate(aod_rank = row_number()) %>%
  arrange(utcige32c_age0t5) %>%
  mutate(utcige32c_rank = row_number())
```

Sixth, to faciliate group sorting in the way we want groups to be displayed in sequence, we now generate a sorting variable so that richer group comes first. 

```{r}
df_lac_aod_pollution_lac <- df_lac_aod_pollution_lac %>%
  mutate(group_sorter = case_when(
    group_sorter_desc == 'High income' ~ 1,
    group_sorter_desc == 'Upper middle income' ~ 2,
    group_sorter_desc == 'Lower middle income' ~ 3,
    group_sorter_desc == 'Without World Bank income group designation' ~ 4
  ))
# display
kable(df_lac_aod_pollution_lac, caption="Input data") %>% kable_styling_fc()
```

#### Generate automatic sorter and grouping frame

Before generate the table using [KableExtra](https://www.rdocumentation.org/packages/kableExtra/versions/1.4.0), we need to do some final preparation to generating a auxillary sorting file to help us group observations by groups and generate proper headings. We generate `df_lac_aod_pollution_lac`, which counts the number of observations within each group, and the starting and ending row for each group.

```{r}
# Sorted file
df_data_sorted <- df_lac_aod_pollution_lac %>%
  arrange(
    group_sorter, country_name,
  ) %>%
  group_by(group_sorter) %>%
  ungroup()
# Count by group
df_group_counts <- df_data_sorted %>%
  group_by(group_sorter_desc, group_sorter) %>%
  summarize(group_count = n()) %>%
  arrange(group_sorter) %>% ungroup() %>%
  mutate(group_count_start = cumsum(group_count) - group_count + 1) %>%
  mutate(group_count_end = cumsum(group_count)) %>%
  select(group_count_start, group_count_end, everything())
# display
kable(df_group_counts, caption="Group counter") %>% kable_styling_fc()
```

#### Formatting inputs with decimal controls and percentage signs

We make a final adjustment to the file so that data values appear with proper formating. 

```{r}
# 4. Format columns, decimals, percentage signs, etc
df_data_formated <- df_data_sorted %>%
  arrange(group_sorter, country_name) %>%
  select(-group_sorter_desc, -group_sorter, -ISOCODE) %>%
  select(
    country_name, 
    aodavg_age0t5, aod_rank, 
    utcige32c_age0t5, utcige32c_rank) %>%
  mutate_at(
    vars(contains("utcige32c_age")),
    list(~ paste0(
      format(round(., 3) * 100,
        nsmall = 1,
        big.mark = ","
      ),
      "%"
    ))
  ) %>%
  mutate_at(
    vars(contains("aodavg_age")),
    list(~ ifelse(abs(.) <= 0.0001,
      paste0(
        formatC(. ,
          digits = 1,
          format = "fg",
          drop0trailing = T,
          flag = "#"
        )
      ),
      paste0(
        format(round(., 3),
          nsmall = 1,
          big.mark = ","
        )
      )
    ))
  )
```

#### We now generate the table

The output tex file is stored here: [lac_climate_children_rank.tex](https://github.com/FanWangEcon/R4Econ/tree/master/table/bigtable/_file/lac_climate_children_rank.tex).

```{r}
bl_main_save <- TRUE
ar_st_kableformat <- c("html", "latex")
for (st_kableformat in ar_st_kableformat) {
  # Column names
  ar_st_col_names <- c(
    "Country name",
    "Child (0-5) mean AOD exposure",
    "LAC child AOD rank",
    "Child (0-5) mean annual share of time over 32 UTCI degrees",
    "LAC child temperature rank"
  )
  # Define column groups, grouping the names above
  # =1/3/2 are number of columns group title covers
  ar_st_col_groups1 <- c(
    " " = 1,
    "Air pollution by aerosols (AOD between 0 and 1)" = 2,
    "At least strong heat stress exposure" = 2
  )
  # Second, we construct main table, and add styling.
  st_title <- paste(
    "LAC country child air pollution and heat exposure ranking in 2010"
    )
  bk_tab_a <- kbl(
    df_data_formated,
    format = st_kableformat,
    label = "tab:lac:aod:temp:rank",
    # escape = F,
    linesep = "",
    booktabs = T,
    longtable = T,
    align = "c",
    caption = st_title,
    col.names = ar_st_col_names
  ) %>%
    # see https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Bootstrap_table_classes
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed", "responsive"),
      full_width = F, position = "left"
    )
  # Third, we add in column groups.
  bk_tab_a <- bk_tab_a %>%
    add_header_above(ar_st_col_groups1)
  # Fourth, we add in row groups.
  for (it_group in seq(1, dim(df_group_counts)[1])) {
    # Reion full name info
    st_loc <- as.character(df_group_counts[[it_group, "group_sorter_desc"]])
    # groups start and end
    it_group_count_start <- df_group_counts[[it_group, "group_count_start"]]
    it_group_count_end <- df_group_counts[[it_group, "group_count_end"]]
    # display text
    st_panel_letter <- base::LETTERS[it_group]
    # Heading group row, year
    st_panel_text <- paste0(
      "Panel ", st_panel_letter, ": ", st_loc 
    )
    # Add to table
    bk_tab_a <- bk_tab_a %>%
      pack_rows(
        st_panel_text, it_group_count_start, it_group_count_end,
        latex_gap_space = "0.25em",
        latex_align = "c",
        hline_after = TRUE
      )
  }
  # Fifth, column formatting.
  fl_width_country <- 6
  st_width_country <- paste0(fl_width_country, "cm")
  bk_tab_a <- bk_tab_a %>%
    column_spec(1, width = st_width_country) %>%
    column_spec(2:dim(df_data_formated)[2], width = "3cm")
  # Final adjustments
  # Headings on all pages, note use `sub` to replace first midrule
  st_headend <- paste0(
    "\\midrule\\endhead\n",
    "\\addlinespace[0.2em]\\hline\\addlinespace[0.2em]\n",
    "\\multicolumn{", dim(df_data_formated)[2], "}{r}{\\emph{Continued on next page}}\\\\\n",
    "\\endfoot\\endlastfoot"
  )
  bk_tab_a <- sub(bk_tab_a,
    pattern = "\\midrule", replacement = st_headend,
    fixed = TRUE
  )

  # country-names left-align
  bk_tab_a <- gsub(bk_tab_a,
    pattern = paste0("\\centering\\arraybackslash}p{", st_width_country, "}"),
    replacement = paste0("\\raggedright\\arraybackslash}p{", st_width_country, "}"),
    fixed = TRUE
  )
  bk_tab_a <- gsub(bk_tab_a,
    pattern = paste0("\\$\\textasciicircum{}\\{\\textbackslash{}circ\\}C\\$"),
    replacement = paste0("$^{\\circ}C$"),
    fixed = TRUE
  )
  bk_tab_a <- gsub(bk_tab_a,
    pattern = paste0("\\$\\textbackslash{}ge\\$"),
    replacement = paste0("$\\ge$"),
    fixed = TRUE
  )
  st_text <- ""
  bk_tab_a <- gsub(bk_tab_a,
    pattern = paste0("\\textbackslash{}", st_text, "\\"),
    replacement = paste0("\\", st_text),
    fixed = TRUE
  )
  # midrule replacing hline
  bk_tab_a <- gsub(bk_tab_a,
    pattern = "hline",
    replacement = "midrule", fixed = TRUE
  )

  # 6. Finally, save table content to file
  if (st_kableformat == "latex") {
    if (bl_main_save) {
      fileConn <- file(spn_tex_out)
      writeLines(bk_tab_a, fileConn)
      close(fileConn)
      if (verbose) {
        print(glue::glue("F-815346, S3"))
        print(glue::glue("Latex saved: {spn_tex_out}"))
      }
    }
  } else if (st_kableformat == "html") {
    bk_tab_a_html <- bk_tab_a
  }
}
bk_tab_a_html
```