---
title: "R Example DPLYR Counting"
author: Fan Wang
output:
  pdf_document: default
  word_document: default
  html_document: default
  html_notebook: default
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [REconTools](https://fanwangecon.github.io/REconTools/) Package, [R4Econ](https://fanwangecon.github.io/R4Econ/) Repository, or [Intro Stats with R](https://fanwangecon.github.io/Stat4Econ/) Repository.

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```

```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(tidyverse)
library(knitr)
library(kableExtra)
library(REconTools)
# file name
st_file_name = 'fs_count_basics'
# Generate R File
try(purl(paste0(st_file_name, ".Rmd"), output=paste0(st_file_name, ".R"), documentation = 2))
# Generate PDF and HTML
# rmarkdown::render("C:/Users/fan/R4Econ/summarize/count/fs_count_basics.Rmd", "pdf_document")
# rmarkdown::render("C:/Users/fan/R4Econ/summarize/count/fs_count_basics.Rmd", "html_document")
```

## Uncount

In some panel, there are $N$ individuals, each observed for $Y_i$ years. Given a dataset with two variables, the individual index, and the $Y_i$ variable, expand the dataframe so that there is a row for each individual index's each unique year in the survey.

*Search*:

- r duplicate row by variable

*Links*:

- see: [Create duplicate rows based on a variable](https://stackoverflow.com/questions/52498169/r-create-duplicate-rows-based-on-a-variable-dplyr-preferred)

*Algorithm*:

1. generate testing frame, the individual attribute dataset with invariant information over panel
2. uncount, duplicate rows by years in survey
3. group and generate sorted index
4. add indiviual specific stat year to index

```{r}
# 1. Array of Years in the Survey
ar_years_in_survey <- c(2,3,1,10,2,5)
ar_start_yaer <- c(1,2,3,1,1,1)
ar_end_year <- c(2,4,3,10,2,5)
mt_combine <- cbind(ar_years_in_survey, ar_start_yaer, ar_end_year)

# This is the individual attribute dataset, attributes that are invariant acrosss years
tb_indi_attributes <- as_tibble(mt_combine) %>% rowid_to_column(var = "ID")

# 2. Sort and generate variable equal to sorted index
tb_indi_panel <- tb_indi_attributes %>% uncount(ar_years_in_survey)

# 3. Panel now construct exactly which year in survey, note that all needed is sort index
# Note sorting not needed, all rows identical now
tb_indi_panel <- tb_indi_panel %>%
                    group_by(ID) %>%
                    mutate(yr_in_survey = row_number())

tb_indi_panel <- tb_indi_panel %>%
                    mutate(calendar_year = yr_in_survey + ar_start_yaer - 1)

# Show results Head 10
tb_indi_panel %>% head(10) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
