---
title: "Solving A Constrained Maximization Problem  \nwith Fixed Cost and Minimum Choice Bound"
author: Fan Wang
date: 2020-04-06
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
always_allow_html: true
---

### Discrete and Continuous

> Go back to [fan](http://fanwangecon.github.io/)'s [REconTools](https://fanwangecon.github.io/REconTools/) Package, [R4Econ](https://fanwangecon.github.io/R4Econ/) Repository, or [Intro Stats with R](https://fanwangecon.github.io/Stat4Econ/) Repository.

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(tidyverse)
library(tidyr)
library(knitr)
library(kableExtra)
```

Study the relationship between fixed cost $\phi$ and minimum choice bounds $a^{min}$. This is a general problem that appears in a lot of context.

There is a cost to choosing $U$ over $B$. And when choosing $B$, there is a minimum choice associated with choosing $B$ category. So the question is, should we choose $U$ or $B$, and if we choose $B$, how much of $a$ should we choose.

Individuals are defined by a single state variable $z$, which determines current wealth, and also helps to determine future wealth. We are interested in solving for the proportion of people choosing $U$ and $B$, and then within those choosing $B$, the proportion of individuals choosing the $a^{min}$

There are three problems to solve:

1. The discrete choice problem over $U$ and $B$
2. The bounding threshold problem
3. The unbounded continuous choice problem

The problem presented here is a simple version of Udupa and Wang (2020)'s savings friction paper. 

#### Unconstrained Savings and Borrowing

There are endowments today and tomorrow. The endowment tomorrow is a function of the endowment today, higher endowment today increases endowment tomorrow. Households are defined by a single state variable $z$. In particular, there are three possible future, same endowment as today, endowment increases by $\epsilon$, or endowment decreases by $\epsilon$

$$
V^{unc}
\left(
  z
\right)
=
\max_{a}
\left(
  \log
    \Big(
      z - \phi - a'
    \Big)
  +
  \beta
  \left(
    \begin{array}{l}
      P_h
        \cdot
      \log
        \Big(
          \exp(z + \epsilon) + a' \cdot R
        \Big) +
        \\
      P_m
        \cdot
      \log
        \Big(
          \exp(z) + a' \cdot R
        \Big) +
        \\
      P_l
        \cdot
      \log
        \Big(
          \exp(z - \epsilon) + a' \cdot R
        \Big) \\
    \end{array}
  \right)
\right)
$$

It is straight forward to solve the above problem numerically. we can solve the problem given a dense grid of $a$ and $z$ points, and find approximately exactly at each $z$ point.

We will solve the problem twice, first using grid method, then using the R bisection method from Fan's [REconTools](https://github.com/FanWangEcon/REconTools/blob/master/R/ff_opti_bisect.R). We should see that the discretized solution is almost identical to the bisection solution which should be more exact. 

Clearly some of the optimal savings or borrowing choices $a^{\prime,\ast}$ will be negative, when the household wants to borrow at some $z$, and it will be positive at other $z$ points where households want to save. 


#### The Asset Choice Constraint 

Now we add in a constraint, the constraint could be a borrowing or savings constraint.

#### Compare Utility 

What would be the optimal choice if households 








