---
title: "R Examples Generate Tibble Dataframes"
titleshort: "Tibble Basics"
description: |
  generate tibbles, rename tibble variables, tibble row and column names
  rename numeric sequential columns with string prefix and suffix
core:
  - package: dplyr
    code: |
      as_tibble(mt)
      rename_all(~c(ar_names))
      rename_at(vars(starts_with("xx")), funs(str_replace(., "yy", "yyyy"))
      rename_at(vars(num_range('',ar_it)), funs(paste0(st,.)))
      rowid_to_column()
      colnames
      rownames
date: 2020-05-27
date_start: 2020-04-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

<!-- includes:
  in_header: ../../preamble.tex -->

### Generate Dataframe

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Simple Meshed Dataframe Name Columns

```{r}
# 5 by 3 matrix
mt_rnorm_a <- matrix(rnorm(4,mean=0,sd=1), nrow=5, ncol=3)

# Column Names
ar_st_varnames <- c('id','var1','varb','vartheta')

# Combine to tibble, add name col1, col2, etc.
tb_combine <- as_tibble(mt_rnorm_a) %>%
  rowid_to_column(var = "id") %>%
  rename_all(~c(ar_st_varnames))

# Display
kable(tb_combine) %>% kable_styling_fc()
```

#### Generate Tibble given Matrixes and Arrays

Given Arrays and Matrixes, Generate Tibble and Name Variables/Columns

- naming tibble columns
- tibble variable names
- dplyr rename tibble
- dplyr rename tibble all variables
- dplyr rename all columns by index
- dplyr tibble add index column
- see also: [SO-51205520](https://stackoverflow.com/questions/45535157/difference-between-dplyrrename-and-dplyrrename-all)

```{r}
# Base Inputs
ar_col <- c(-1,+1)
mt_rnorm_a <- matrix(rnorm(4,mean=0,sd=1), nrow=2, ncol=2)
mt_rnorm_b <- matrix(rnorm(4,mean=0,sd=1), nrow=2, ncol=4)

# Combine Matrix
mt_combine <- cbind(ar_col, mt_rnorm_a, mt_rnorm_b)
colnames(mt_combine) <- c('ar_col',
                          paste0('matcolvar_grpa_', seq(1,dim(mt_rnorm_a)[2])),
                          paste0('matcolvar_grpb_', seq(1,dim(mt_rnorm_b)[2])))

# Variable Names
ar_st_varnames <- c('var_one',
                    paste0('tibcolvar_ga_', c(1,2)),
                    paste0('tibcolvar_gb_', c(1,2,3,4)))

# Combine to tibble, add name col1, col2, etc.
tb_combine <- as_tibble(mt_combine) %>% rename_all(~c(ar_st_varnames))

# Add an index column to the dataframe, ID column
tb_combine <- tb_combine %>% rowid_to_column(var = "ID")

# Change all gb variable names
tb_combine <- tb_combine %>%
                  rename_at(vars(starts_with("tibcolvar_gb_")),
                            funs(str_replace(., "_gb_", "_gbrenamed_")))

# Tibble back to matrix
mt_tb_combine_back <- data.matrix(tb_combine)

# Display
kable(mt_combine) %>% kable_styling_fc_wide()
kable(tb_combine) %>% kable_styling_fc_wide()
kable(mt_tb_combine_back) %>% kable_styling_fc_wide()
```

#### Rename Tibble with Numeric Column Names

After reshaping, often could end up with variable names that are all numeric, intgers for example, how to rename these variables to add a common prefix for example.

```{r}
# Base Inputs
ar_col <- c(-1,+1)
mt_rnorm_c <- matrix(rnorm(4,mean=0,sd=1), nrow=5, ncol=10)
mt_combine <- cbind(ar_col, mt_rnorm_c)

# Variable Names
ar_it_cols_ctr <- seq(1, dim(mt_rnorm_c)[2])
ar_st_varnames <- c('var_one', ar_it_cols_ctr)

# Combine to tibble, add name col1, col2, etc.
tb_combine <- as_tibble(mt_combine) %>% rename_all(~c(ar_st_varnames))

# Add an index column to the dataframe, ID column
tb_combine_ori <- tb_combine %>% rowid_to_column(var = "ID")

# Change all gb variable names
tb_combine <- tb_combine_ori %>%
                  rename_at(
                    vars(num_range('',ar_it_cols_ctr)),
                    funs(paste0("rho", . , 'var'))
                    )

# Display
kable(tb_combine_ori) %>% kable_styling_fc_wide()
kable(tb_combine) %>% kable_styling_fc_wide()
```

#### Tibble Row and Column and Summarize
Show what is in the table: 1, column and row names; 2, contents inside table.

```{r}
tb_iris <- as_tibble(iris)
print(rownames(tb_iris))
colnames(tb_iris)
colnames(tb_iris)
summary(tb_iris)
```

#### Tibble Sorting

- dplyr arrange desc reverse
- dplyr sort

```{r}
# Sort in Ascending Order
tb_iris %>% select(Species, Sepal.Length, everything()) %>%
  arrange(Species, Sepal.Length) %>% head(10) %>%
  kable() %>% kable_styling_fc()

# Sort in Descending Order
tb_iris %>% select(Species, Sepal.Length, everything()) %>%
  arrange(desc(Species), desc(Sepal.Length)) %>% head(10) %>%
  kable() %>% kable_styling_fc()
```

#### REconTools Summarize over Tible

Use R4Econ's summary tool.

```{r}
df_summ_stats <- ff_summ_percentiles(tb_iris)
kable(t(df_summ_stats)) %>% kable_styling_fc_wide()
```
