---
title: "R Analyze Constant Elasticity of Substitution (Atkinson Family Utility)"
titleshort: "CES abd Atkinson Utility"
description: |
  Analyze how changing individual outcomes shift utility given inequality preference parameters.
  Draw Cobb-Douglas, Utilitarian and Leontief indifference curve
core:
  - package: r
    code: |
      apply(mt, 1, funct(x){})
      do.call(rbind, ls_mt)
  - package: tidyr
    code: |
      expand_grid()
  - package: ggplot2
    code: |
      geom_line()
      facet_wrap()
date: 2020-04-14
date_start: 2020-04-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Atkinson Family Utility

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Individual Outcomes and Preference

How does the Aktinson Family utility function work? THe Atkinson Family Utility has the following functional form.

$$
V^{\text{social}}
=
\left(
\alpha
\cdot
A^{\lambda}
+
\beta
\cdot
B^{\lambda}
\right)^{\frac{1}{\lambda}}
$$

Several key issues here:

1. $V^{\text{social}}$ is the utility of some social planner
2. $A$ and $B$ are allocations for Alex and Ben.
3. $\alpha$ and $\beta$ are biases that a social planner has for Alex and Ben: $\alpha+\beta=1$, $\alpha>0$, and $\beta>0$
4. $-\infty < \lambda \le 1$ is a measure of inequality aversion
    * $\lambda=1$ is when the planner cares about weighted total allocations (efficient, Utilitarian)
    * $\lambda=-\infty$ is when the planner cares about only the minimum between $A$ and $B$ allocations (equality, Rawlsian)

What if only care about Alex? Clearly, if the planner only cares about Ben, $\beta=1$, then:

$$
V^{\text{social}}
=
\left(
B^{\lambda}
\right)^{\frac{1}{\lambda}}
= B
$$

Clearly, regardless of the value of $\lambda$, as $B$ increases $V$ increases. What Happens to V when A or B increases? What is the derivative of $V$ with respect to $A$ or $B$?

$$
    \frac{\partial V}{\partial A}
    =
    \frac{1}{\lambda}
    \left(
    \alpha
    A^{\lambda}
    +
    \beta
    B^{\lambda}
    \right)^{\frac{1}{\lambda}-1}
    \cdot
    \lambda
    \alpha
    A^{\lambda -1}
$$

$$
    \frac{\partial V}{\partial A}
    =
    \left(
    \alpha
    A^{\lambda}
    +
    \beta
    B^{\lambda}
    \right)^{\frac{1-\lambda}{\lambda}}
    \cdot
    \alpha
    A^{\lambda -1}
    >0
$$


Note that $\frac{\partial V}{\partial A}>0$. When $\lambda <0$, $Z^{\lambda}>0$. For example $10^{-2}=\frac{1}{100}$. And For example $0.1^{\frac{3}{-2}}=\frac{1}{0.1^{1.5}}$. Still Positive.

While the overall $V$ increases with increasing $A$, but if we did not have the outter power term, the situation is different. In particular, when $\lambda < 0$:
$$
\text{ if } \lambda <0
\thinspace\thinspace
\text{ then }
\thinspace\thinspace
\frac{d \left(\alpha A^{\lambda} + \beta B^{\lambda}\right)}{dA}=\alpha\lambda A^{\lambda -1}<0
$$
Without the outter $\frac{1}{\lambda}$ power, negative $\lambda$ would lead to decreasing weighted sum. But:
$$
\text{ if } \lambda <0
\thinspace\thinspace
\text{ then }
\thinspace\thinspace
\frac{dG^{\frac{1}{\lambda}}}{dG}=\frac{1}{\lambda}\cdot G^{\frac{1-\lambda}{\lambda}}<0
$$
so when $G$ is increasing and $\lambda <0$, $V$ would decrease. But when $G\left(A,B\right)$ is decreasing, as is the case with increasing $A$ when $\lambda <0$, $V$ will actually increase. This confirms that $\frac{\partial V}{\partial A}>0$ for $\lambda <0$. The result is symmetric for $\lambda >0$.


#### Indifference Curve Graph 

Given $V^{\ast}$, we can show the combinations of $A$ and $B$ points that provide the same utility. We want to be able to potentially draw multiple indifference curves at the same time. Note that indifference curves are defined by $\alpha$, $\lambda$ only. Each indifference curve is a set of $A$ and $B$ coordinates. So to generate multiple indifference curves means to generate many sets of $A$, $B$ associated with different planner preferences, and then these could be graphed out. 


```{r}
# A as x-axis, need bounds on A
fl_A_min = 0.01
fl_A_max = 3
it_A_grid = 10000

# Define parameters
# ar_lambda <- 1 - (10^(c(seq(-2,2, length.out=3))))
ar_lambda <- c(1, 0.6, 0.06, -6)
ar_beta <- seq(0.25, 0.75, length.out = 3)
ar_beta <- c(0.3, 0.5, 0.7)
ar_v_star <- seq(1, 2, length.out = 1)
tb_pref <- as_tibble(cbind(ar_lambda)) %>% 
  expand_grid(ar_beta) %>% expand_grid(ar_v_star) %>% 
  rename_all(~c('lambda', 'beta', 'vstar')) %>% 
  rowid_to_column(var = "indiff_id")

# Generate indifference points with apply and anonymous function
# tb_pref, whatever is selected from it, must be all numeric
# if there are strings, would cause conversion error.
ls_df_indiff <- apply(tb_pref, 1, function(x){
  indiff_id <- x[1]
  lambda <- x[2]
  beta <- x[3]
  vstar <- x[4]
  ar_fl_A_indiff <- seq(fl_A_min, fl_A_max, length.out=it_A_grid)
  ar_fl_B_indiff <- (((vstar^lambda) - 
                        (beta*ar_fl_A_indiff^(lambda)))/(1-beta))^(1/lambda)
  mt_A_B_indiff <- cbind(indiff_id, lambda, beta, vstar,
                         ar_fl_A_indiff, ar_fl_B_indiff)
  colnames(mt_A_B_indiff) <- c('indiff_id', 'lambda', 'beta', 'vstar',
                               'indiff_A', 'indiff_B')
  tb_A_B_indiff <- as_tibble(mt_A_B_indiff) %>% 
    rowid_to_column(var = "A_grid_id") %>% 
    filter(indiff_B >= 0 & indiff_B <= max(ar_fl_A_indiff))
  return(tb_A_B_indiff)
})
df_indiff <- do.call(rbind, ls_df_indiff) %>% drop_na()
```

Note that many more A grid points are needed to fully plot out the leontief line. 

```{r}
# Labeling
st_title <- paste0('Indifference Curves Aktinson Atkinson Utility (CES)')
st_subtitle <- paste0('Each Panel Different beta=A\'s Weight lambda=inequality aversion\n',
                      'https://fanwangecon.github.io/',
                      'R4Econ/math/func_ineq/htmlpdfr/fs_atkinson_ces.html')
st_caption <- paste0('Indifference Curve 2 Individuals, ',
                     'https://fanwangecon.github.io/R4Econ/')
st_x_label <- 'A'
st_y_label <- 'B'

# Graphing
plt_indiff <- 
  df_indiff %>% mutate(lambda = as_factor(lambda),
                       beta = as_factor(beta),
                       vstar = as_factor(vstar)) %>%
  ggplot(aes(x=indiff_A, y=indiff_B,
             colour=lambda)) +
  facet_wrap( ~ beta) +
  geom_line(size=1) +
  labs(title = st_title, subtitle = st_subtitle,
       x = st_x_label, y = st_y_label, caption = st_caption) +
  theme_bw()

# show
print(plt_indiff)
```
