---
title: "Share of Environmental Exposure Burden Across Population Groups"
titleshort: "Share of Environmental Exposure Burden Across Population Groups"
description: |
  Simulate pollution exposures by location.
  Compute share of pollution burden for a population group relative to the share of overall population accounted for by this population group.
  core:
    matrix()
    runif()
    sum()
    arrange()
    group_by()
    left_join()
    filter()
    slice()
  - package: r
    code: |
    - package: stats
    code: |
    - package: dplyr
    code: |
    date: 2021-12-27
date_start: 2021-12-27
output:
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
always_allow_html: true
urlcolor: blue
---

### Location, Population, and Pollution

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Location, Population, and Environmental Exposures (Pollution)

##### Local Public Exposures $Z_{l,y}$

Environmental exposure, specifically PM 2.5, is a local public good (bad). Prior research has shown that due to the nature of particular matter formation in the air, while there is particular matter variation at the city level, there is not significant variation in particular matter pollution at the neighborhood level [He et al. (2019)](https://www.aeaweb.org/articles?id=10.1257/app.20170286) and [Liu et al. (2022)](https://fanwangecon.github.io/assets/FanWang_PollutionTempBirthweight.pdf). Additionally, there is a difference between ambient environmental exposures and the amount of particular matter inhaled by individual residents: the former is a common level shared by all residents, but the latter can differ depending on individuals' physical and socio-economic attributes.

There are $M$ different locations (cities/counties/townships), indexed from $l=1$ to $l=M$. In a particular time-period, we assume that the potential pollution exposure for all residents in the same location. Suppose additionally that we compute statistics at the interval of years $y$, and each year includes daily pollution measure for each day of the year $t$, indexed from $t=1$ to $t=T_y$, where $T_y$ is the total number of days in a year.

Let $\text{Z}_{l,y}$ be the total pollution exposure by a resident in location $l$. This is equal to the sum of pollution exposure during the course of a year:

$$
Z_{l,y} = \sum_{t=1}^{T_y} Z_{l,y,t}
$$

##### Group Exposures $\mathcal{Z}_{i,y}$

Let there be $N$ population groups, indexed from $i=1$ to $i=N$. The $N$ population groups reside in the $M$ locations. the share, let $P_{l,i}$ denote share of population belong to group $i$ that resides in location $l$:

$$
1 =
\sum_{l=1}^{M}
\sum_{i=1}^{N}
P_{l,i}
$$

The share of population belong to population group $i$ is:
$$
P_i =
\sum_{l=1}^{M}
P_{l,i}
$$

$\mathcal{Z}_{i,y}$, which is the average pollution exposure facing an individual belonging to group $i$ in year $y$, is determined by how individuals from population group $i$ are distributed across the $M$ locations:

$$
\mathcal{Z}_{i,y} =
\sum_{l=1}^{M}
\left(
\frac{P_{l,i}}{P_i}
\times
Z_{l,y}
\right)
$$

Additionally, $\mathcal{Z}_{y}$ is the average pollution exposure facing an individual, regardless of population group, in year $y$:

$$
\mathcal{Z}_{y} =
\sum_{i=1}^N
\left(
\sum_{l=1}^{M}
\left(
\frac{P_{l,i}}{P_i}
\times
Z_{l,y}
\right)
\times
P_i
\right)
= \sum_{i=1}^N
\sum_{l=1}^{M}
\left(
P_{l,i}
\times
Z_{l,y}
\right)
$$

##### Excess Environmental Exposure Burden to Population

We have $\mathcal{Z}_{i,y}$, the average pollution burden facing an individual in a particular population group. We also have $P_i$, the share of population belong to population group $i$.

How does the share of pollution burden facing a population group relate to the share of population this group has in the overall population?

We define $\mathcal{E}_{i,y}$ as the share of pollution burden for population group $i$ that is in excess of its population share as:
$$
\mathcal{E}_{i,y} =
\left(
\left(
\frac{\mathcal{Z}_{i,y} \times P_i
}{
\sum_{\widehat{i}}^{N}
\left(
\mathcal{Z}_{\widehat{i},y}
\times P_{\widehat{i}}
\right)
}
\right)
\times
\frac{1}{P_i}
\right)
-1
= \frac{
\mathcal{Z}_{i,y}
}{
\mathcal{Z}_{y}
}
-1
$$

Note that $\mathcal{E}_{i,y}$ is simply the ratio between the average pollution exposure for an individual in group $i$ in year $y$ and the overall average pollution exposure for an individual in year $y$, minus 1.

##### Scenarios for Within Group Exposure Variations

The prior sections discussed how to measure environmental exposure (pollution) variations across groups. For example, we might learn that $\mathcal{E}_{i=\text{white},y}=-0.60$ and $\mathcal{E}_{i=\text{black},y}=+0.60$: black population is a lot more exposed to pollution relative to their population than white people are, due to where they live. 

The next question we want to ask is how much variation there is in environmental exposure within group. The interest in this is natural. Limited difference in across group variation could mask great inequality in environmental exposures within group, something relevant for the concept of environmental justice. Additionally, if variations within group are large, it is not clear that mean difference across group would properly capture the magnitude of environmental inequality that exists within a population. 

Suppose there are two locations and two population groups, we have the following scenarios:

1. Zero within group variation for both population groups:
   + In the most extreme case, all people of one population group live in one location, and all people of another population group live in another location. 
   + In this case, there would be zero within group variation in environmental exposure. 
2. Positive within group variation, but identical within group variation for both population groups:
   + Suppose the distribution of each of the population groups across locations could be identical (say 40 percent in location 1, 60 percent in location 2).
   + As long as environmental exposure differs across the locations, there will be within group exposure variation.
   + However, the within group variations will be the same for both population groups. 
3. Different within group variations across population groups:
   + The population distribution across location varies for the population groups. Group 1 might have 40 percent in location 1, group 2 might have 70 percent in location 1. 
   + Same as case (2), as long as environmental exposure differs across the locations, there will be within group exposure variation.
   + But now, additionally, the within group variations will be different across the population groups. In the most extreme case, one group---if they all reside in one location---might have no within group variation, however, another group---if they co-resides in multiple locations---will have positive within group variations. 

These three scenarios above present contrasting worlds of environmental inequality which can only be captured by summarizing the magnitude of within-group exposure inequalities. 

##### Within Group Exposure Distribution Percentiles 

In our setting here, it is important to note that the within group environmental exposure is driven only by environmental exposure differences across locations. 

<!-- First, we sort over the $M$ locations, and create an rank index, $R_{l,y}$, that ranks locations from the least to the most exposed (polluting, note that higher $Z_{l,y}$ is more polluting):

$$
R_{l,y} = \sum_{\tilde{l}=1}^M 
\mathbf{1}\left\{ 
  Z_{l,y} \ge \text{Z}_{\tilde{l},y}
  \right\}
$$ -->

<!-- The indicator 1 denotes the indicator function. Note that if $R_{l,y}=1$, that means location $l$ is the least polluting location. If $R_{\tilde{l},y}=M$, that means location $\tilde{l}$ is the most pollution location among the $M$ locations.  -->

$Q_{i,y}(\tau)$ is the quantile function for population group $i$ in year $y$ where $\tau\in[0,1)$ and $\tau=0.01$ is 1 percent. The population-group-specific quantile function $Q_{i,y}(\tau)$ is defined as: 

$$
Q_{i,y}(\tau) = 
\max_{l\in\left\{1, \dots, M \right\}}
\left(   
    \text{Z}_{l,y}
    \times
    \mathbf{1}\left\{
    \tau \ge
    % \left(
      \sum_{\tilde{l}=1}^{M}  
      \frac{P_{i,\tilde{l}}}{P_i}
      \times
      \mathbf{1}\left\{
        Z_{\tilde{l},y} \le Z_{l,y}
        \right\}
    % \right)
    \right\}
\right)
$$

Note that the $Q_{i,y}(\tau)$ differs for each $i$ because of the $i$-specific population distribution across location differs ($\frac{P_{i,\tilde{l}}}{P_i}$). But all quantile functions share the same rank index $R_{l,y}$ and location-specific pollution measures $\text{Z}_{l,y}$.

Given $Q_{i,y}(\tau)$, we can construct the ratio of two quantiles. Of particular interest is the ratio of the 80th and the 20th quantile for each population group $i$:

$$
\mathcal{Q}^{80/20}_i = \frac{Q_{i,y}(\tau=0.8)}{Q_{i,y}(\tau=0.2)}
$$

Where we use $\mathcal{Q}^{80/20}_i$ to denote the P80 to P20 ratio for population group $i$. We can compare this statistics across population groups to see which group has larger within group environmental exposure variation. Given that we know the quantile function for each population group, we can easily construct any alternative statistics based on each population group's within group environmental exposure distribution.


#### Simulate Population Distribution over Location and Demographics

Use the binomial distribution to generate heterogenous demographic break-down by location. There are N demographic cells, and the binomial distribution provides the probability mass in each of the N cell. Different bernoulli "win" chance for each location. There is also probability distribution over population in each location.

First, construct empty population share dataframe:

```{r}
# 7 different age groups and 12 different locationso
it_N_pop_groups <- 100
it_M_location <- 20
# Matrix of demographics by location
mt_pop_data_frac <- matrix(data=NA, nrow=it_M_location, ncol=it_N_pop_groups)
colnames(mt_pop_data_frac) <- paste0('popgrp', seq(1,it_N_pop_groups))
rownames(mt_pop_data_frac) <- paste0('location', seq(1,it_M_location))

# For succinct visualization select subset of population groups to display
it_popgrp_disp <- 7
ar_it_popgrp_disp <- seq(1, it_N_pop_groups, length.out=it_popgrp_disp)
ar_it_popgrp_disp <- round(ar_it_popgrp_disp)
st_popgrp_disp <- paste0('(', it_popgrp_disp, ' of ', it_N_pop_groups, ' pop-groups shown)')
it_loc_disp <- 10
ar_it_loc_disp <- seq(1, it_M_location, length.out=it_loc_disp)
ar_it_loc_disp <- round(ar_it_loc_disp)
st_loc_disp <- paste0('(', it_loc_disp, ' of ', it_M_location, ' locations shown)')

# Display
st_caption = paste('Location and demographic cell',
  st_popgrp_disp, st_loc_disp, sep=" ")
mt_pop_data_frac[ar_it_loc_disp, ar_it_popgrp_disp] %>%
  kable(caption = st_caption) %>%
  kable_styling_fc()
```

Second, generate conditional population distribution for each location, and then multiply by the share of population in each locality:

```{r}
# Share of population per location
set.seed(123)
ar_p_loc <- dbinom(0:(3*it_M_location-1), 3*it_M_location-1, 0.5)
it_start <- length(ar_p_loc)/2-it_M_location/2
ar_p_loc <- ar_p_loc[it_start:(it_start+it_M_location-1)]
ar_p_loc <- ar_p_loc/sum(ar_p_loc)

# Different bernoulli "win" probability for each location
set.seed(234)
# ar_fl_unif_prob <- sort(runif(it_M_location)*(0.25)+0.4)
ar_fl_unif_prob <- sort(runif(it_M_location))

# Generate population proportion by locality
for (it_loc in 1:it_M_location ) {
  ar_p_pop_condi_loc <- dbinom(0:(it_N_pop_groups-1), it_N_pop_groups-1, ar_fl_unif_prob[it_loc])
  mt_pop_data_frac[it_loc,] <- ar_p_pop_condi_loc*ar_p_loc[it_loc]
}

# Sum of cells, should equal to 1
print(paste0('pop frac sum = ', sum(mt_pop_data_frac)))

# Display
st_caption = paste('Share of population in each location and demographic cell',
  st_popgrp_disp, st_loc_disp, sep=" ")
round((mt_pop_data_frac[ar_it_loc_disp, ar_it_popgrp_disp])*100, 3) %>%
  kable(caption=st_caption) %>%
  kable_styling_fc()
```

#### Simulate Environmental Exposure

Use log-normal distribution to describe average daily PM10 exposures distribution by locality:

```{r}
fl_meanlog <- 3.4
fl_sdlog <- 0.35
hist(rlnorm(1000, meanlog = fl_meanlog, sdlog = fl_sdlog))
```

First, draw pollution measure for each locality:

```{r}
# draw
set.seed(123)
ar_pollution_loc <- rlnorm(it_M_location, meanlog = fl_meanlog, sdlog = fl_sdlog)
# pollution dataframe
# 5 by 3 matrix

# Column Names
ar_st_varnames <- c('location','avgdailypm10')

# Combine to tibble, add name col1, col2, etc.
tb_loc_pollution <- as_tibble(ar_pollution_loc) %>%
  rowid_to_column(var = "id") %>%
  rename_all(~c(ar_st_varnames)) %>%
  mutate(location = paste0('location', location))

# Display
st_caption = paste('PM10 Exposure across locations', st_loc_disp, sep=" ")
tb_loc_pollution[ar_it_loc_disp,] %>% 
  kable(caption = st_caption) %>% kable_styling_fc()
```

Second, reshape population data:

```{r}
# Reshape population data, so each observation is location/demo
df_pop_data_frac_long <- as_tibble(mt_pop_data_frac, rownames='location') %>%
  pivot_longer(cols = starts_with('popgrp'),
               names_to = c('popgrp'),
               names_pattern = paste0("popgrp(.*)"),
               values_to = "pop_frac")
```

Third, join with pollution data:

```{r}
# Reshape population data, so each observation is location/demo
df_pop_pollution_long <- df_pop_data_frac_long %>%
  left_join(tb_loc_pollution, by='location')

# display
st_caption = paste('Population x Location Long Frame (15 rows shown)', sep=" ")
df_pop_pollution_long[
  round(seq(1, dim(df_pop_pollution_long)[1], length.out=15)),] %>% 
  kable(caption = st_caption) %>% kable_styling_fc()
```

#### Compute Demographic Group Specific Exposure Distributions

What is the p10, median, p90 and mean pollution exposure for each demographic group?

1. group by population group
2. sort by pollution exposure within group
3. generate population group specific conditional population weights
4. generate population CDF for each population group (sorted by pollution)

```{r}
# Follow four steps above
df_pop_pollution_by_popgrp_cdf <- df_pop_pollution_long %>%
  arrange(popgrp, avgdailypm10) %>%
  group_by(popgrp) %>%
  mutate(cdf_pop_condi_popgrp_sortpm10 = cumsum(pop_frac/sum(pop_frac)),
         pmf_pop_condi_popgrp_sortpm10 = (pop_frac/sum(pop_frac)))

# Display
st_caption = paste('Distribution within groups, sorted CDFs (15 rows shown)', sep=" ")
df_pop_pollution_by_popgrp_cdf[
  round(seq(1, dim(df_pop_pollution_by_popgrp_cdf)[1], length.out=15)),] %>%
  kable(caption = st_caption) %>% kable_styling_fc_wide()
```

#### Various Relative Burden Statistics

What to compute?

1. Excess pollution burden: Share of pollution burden by population group and overall population share, this is simply the ratio of population group mean and the overall weighted mean.
2. What is the fraction of the people in each population group with below and above overall average?
3. Merge results for different quantiles together.

##### Group-specific Means and Excess pollution burden

We compute within group means, $\mathcal{Z}_{i,y}$. We compute excess population burden, $\mathcal{E}_{i,y}$, *pm10_grp_exc_burden*. 0.10 means 10 percent in excess, this means the pollution burden share is 10 percent in excess of the population share. -0.10 means 10 percent less than what population share is.

Additionally, we compute the share of people within group above the overall mean: *pm10_grp_shr_exc*. This shows the share of people having excess burden. This complements the first number. Because the 10 percent excess could be due to very high exposure to a very small number of people within a population group, or it could be that most people in the group are in "excess".

```{r}
# Stats 1: excess pollution burden
df_excess_pollution_burden <- df_pop_pollution_by_popgrp_cdf %>%
  ungroup() %>%
  mutate(pm10_overall_mean = weighted.mean(avgdailypm10, pop_frac)) %>%
  group_by(popgrp) %>%
  mutate(
    popgrp_mass = sum(pop_frac), # The share of population for this group
    pm10_grp_mean = weighted.mean(avgdailypm10, pop_frac) # Pop-group mean
  ) %>%
  slice(1) %>%
  mutate(pm10_grp_exc_burden = pm10_grp_mean/pm10_overall_mean - 1) %>%
  select(popgrp, popgrp_mass,
         pm10_grp_mean, pm10_overall_mean, pm10_grp_exc_burden)
fl_pm10_overall_mean <- mean(df_excess_pollution_burden %>% pull(pm10_overall_mean))

# Stats 2: share of people within group below or above overall mean
df_share_below_or_excess <- df_pop_pollution_by_popgrp_cdf %>%
  arrange(popgrp, avgdailypm10) %>%
  filter(avgdailypm10 < fl_pm10_overall_mean) %>%
  slice_tail() %>%
  mutate(pm10_grp_shr_exc = 1 - cdf_pop_condi_popgrp_sortpm10) %>%
  select(popgrp, pm10_grp_shr_exc)
# merge stats 2 with stats 1
df_excess_pollution_burden <- df_excess_pollution_burden %>%
  left_join(df_share_below_or_excess, by="popgrp")

# display
st_caption = paste('Mean and Excess Burden by Population Groups',
  st_popgrp_disp, sep=" ")
df_excess_pollution_burden[ar_it_popgrp_disp,] %>%
  kable(caption = st_caption) %>%
  kable_styling_fc_wide()
```

##### Within Group Percentiles

Compute within group percentiles for each population groups. Use the list of percentiles below to specify which percentiles should be computed.

```{r}
# Stats 3: percentiles and ratios
ar_fl_percentiles <- c(0.1, 0.2, 0.8, 0.9)
# Stats 3a: generate key within group percentiles
# 1. 20th and 80th percentiles
# 2. 10th and 90th percentiles
# 3. 50th percentile
# Generate pollution quantiles by population groups
for (it_percentile_ctr in seq(1, length(ar_fl_percentiles))) {

  # Current within group percentile to compute
  fl_percentile <- ar_fl_percentiles[it_percentile_ctr]
  svr_percentile <- paste0('pm10_p', round(fl_percentile*100))

  # Frame with specific percentile
  df_within_percentiles_cur <- df_pop_pollution_by_popgrp_cdf %>%
    group_by(popgrp) %>%
    filter(cdf_pop_condi_popgrp_sortpm10 >= fl_percentile) %>%
    slice(1) %>%
    mutate(!!sym(svr_percentile) := avgdailypm10) %>%
    select(popgrp, one_of(svr_percentile))

  # Merge percentile frames together
  if (it_percentile_ctr > 1) {
    df_within_percentiles <- df_within_percentiles %>%
      left_join(df_within_percentiles_cur, by='popgrp')
  } else {
    df_within_percentiles <- df_within_percentiles_cur
  }
}

# display
st_caption = paste('PM10 Exposure Distribution by Population Groups',
  st_popgrp_disp, sep=" ")
df_within_percentiles[ar_it_popgrp_disp,] %>%
  kable(caption = st_caption) %>%
  kable_styling_fc()
```

##### Percentiles As Excess Burden

The 80th percentile for a population group, how is this exposed relative to the mean? We simply divide the within group pollution percentiles by the overall mean across all groups. All are properly weighted.

This is relating within group percentiles to the overall mean. These can be interpreted as excess burdens at specific percentiles. Individuals at the 80th percentile of a particular population group, how does their pollution burden compare to their population share?

```{r}
# merge stats 3 with stats 1 and 2
df_excess_pollution_burden <- df_excess_pollution_burden %>%
  left_join(df_within_percentiles, by="popgrp")

# Stats 3b: Percentiles to Relative Burdens
# Convert percentiles to be relative of overall means
for (it_percentile_ctr in seq(1, length(ar_fl_percentiles))) {

  # Current within group percentile to compute
  fl_percentile <- ar_fl_percentiles[it_percentile_ctr]
  svr_percentile <- paste0('pm10_p', round(fl_percentile*100))
  svr_perc_exc_burden <- paste0('pm10_grp_excbrd_p', round(fl_percentile*100))

  # Percentiles to excess percentiles
  df_excess_pollution_burden  <- df_excess_pollution_burden  %>%
    mutate(!!sym(svr_perc_exc_burden) := (!!sym(svr_percentile)/pm10_overall_mean) - 1)
}

# display
st_caption = paste('PM10 Within Population Group Percentiles and Excess Burden',
  st_popgrp_disp, sep=" ")
df_excess_pollution_burden[ar_it_popgrp_disp,] %>%
  select(-pm10_overall_mean,
         -starts_with('pm10_p')) %>%
  kable(caption = st_caption) %>%
  kable_styling_fc_wide()
```

##### Within Group Relative Exposure Ratios Across Percentiles

We now compute within group relative ratios of interest. This is purely within group inequality.

```{r}
# lower and upper bound or relative within group ratios
# can only use values appearing in the percentiles list prior
ar_fl_ratio_upper <- c(0.8, 0.9)
ar_fl_ratio_lower <- c(0.2, 0.1)
# Stats 4c: Ratios
# Generate P80 to P20 ratio, and P90 to P10 standard inequality ratios
for (it_ratio_ctr in seq(1, length(ar_fl_ratio_upper))) {

  # Upper and lower percentile bounds
  fl_ratio_upper <- ar_fl_ratio_upper[it_ratio_ctr]
  fl_ratio_lower <- ar_fl_ratio_lower[it_ratio_ctr]
  svr_ratio_upper_perc <- paste0('pm10_p', round(fl_ratio_upper*100))
  svr_ratio_lower_perc <- paste0('pm10_p', round(fl_ratio_lower*100))

  # New relative within group ratio variable name
  svr_ratio <- paste0('pm10_rat_p', round(fl_ratio_upper*100), '_dvd_p', round(fl_ratio_lower*100))

  # Generate P80 to P20 ratio, etc.
  df_excess_pollution_burden  <- df_excess_pollution_burden  %>%
    mutate(!!sym(svr_ratio) := !!sym(svr_ratio_upper_perc)/!!sym(svr_ratio_lower_perc))
}

# display
st_caption = paste('PM10 Exposure within Population Group P80-P20 Inequality',
  st_popgrp_disp, sep=" ")
df_excess_pollution_burden[ar_it_popgrp_disp,] %>%
  select(-pm10_overall_mean,
         -starts_with('pm10_grp_excbrd_p'),
         -starts_with('pm10_p')) %>%
  kable(caption = st_caption) %>%
  kable_styling_fc_wide()
```

#### Summary Statistics for Inequality across Groups

Whether we are looking at inequality within group, or inequality across groups, we can compute various aggregate inequality related statistics, including:

1. Gini
2. Atkinson
3. Standard deviation
4. P80 to P20 ratio

The literature has focused on Atkinson statistics computed over mean pollution exposures across groups. When we focus on variations across groups, we compute these based on group means. When we focus on variations within groups, we compute group-specific GINI, Atkinson, etc based on within group data.

Earlier, we have already computed the P80 to P20 ratio for each population group.

##### Gini, Atkinson, and S.D. Functions

For variations across groups, note that the GINI and ATKINSON require positive values for inputs, so they can be used with $\mathcal{Z}_{i,y}$, but they can not be applied to the $\mathcal{E}_{i,y}$, which is positive or negative. For variations across groups, we compute these statistics given the distribution of $P_i$.

First, we generate an internal function for GINI:
$$
\text{GINI Index} =
1 - \frac{2}{N+1}
\cdot
\left(\sum_{i=1}^N \sum_{j=1}^{i} x_j\right)
\cdot
\left(
\sum_{i=1}^N x_i
\right)^{-1}
$$
We use the Gini index which is documented in on the [fs_gini_disc](https://fanwangecon.github.io/R4Econ/math/func_ineq/htmlpdfr/fs_gini_disc.html) page.
```{r}
ffi_dist_gini_random_var_pos_test <- function(ar_data, ar_prob_data) {
  #' @param ar_data array sorted array values low to high
  #' @param ar_prob_data array probability mass for each element along `ar_data`, sums to 1

  fl_mean <- sum(ar_data*ar_prob_data);
  ar_mean_cumsum <- cumsum(ar_data*ar_prob_data);
  ar_height <- ar_mean_cumsum/fl_mean;
  fl_area_drm <- sum(ar_prob_data*ar_height);
  fl_area_below45 <- sum(ar_prob_data*(cumsum(ar_prob_data)/sum(ar_prob_data)))
  fl_gini_index <- (fl_area_below45-fl_area_drm)/fl_area_below45
  return(fl_gini_index)
}
```

Second, we create an internal function for Atkinson statistics, from [Atkinson (JET, 1970)](https://linkinghub.elsevier.com/retrieve/pii/0022053170900396). The formula is:
$$
\text{Atkinson Index} =
A\left(
\left\{Y_i\right\}_{i=1}^N,
\lambda
\right)
= 1 -
\left(
\sum_{i=1}^N \frac{1}{N}
\left(
\frac{Y_i}{\sum_{j=1}^N \left( \frac{Y_j}{N} \right) }
\right)^{\lambda}
\right)^{\frac{1}{\lambda}}
\in \left[0,1\right]
$$
Note that the Atkinson statistics differ by planner preference, and is arbitrary. We use the Atkinson index formula which is documented on the [fs_atkinson_ces](https://fanwangecon.github.io/R4Econ/math/func_ineq/htmlpdfr/fs_atkinson_ces.html) page.
```{r}
# Formula
ffi_atkinson_random_var_ineq <- function(ar_data, ar_prob_data, fl_rho) {
  #' @param ar_data array sorted array values
  #' @param ar_prob_data array probability mass for each element along `ar_data`, sums to 1
  #' @param fl_rho float inequality aversion parameter fl_rho = 1 for planner
  #' without inequality aversion. fl_rho = -infinity for fully inequality averse.

  fl_mean <- sum(ar_data*ar_prob_data);
  fl_atkinson <- 1 - (sum(ar_prob_data*(ar_data^{fl_rho}))^(1/fl_rho))/fl_mean
  return(fl_atkinson)
}
```

Third, we also create an internal function for standard deviation for discrete random variable.
```{r}
# Formula
ffi_std_drv <- function(ar_data, ar_prob_data) {
  #' @param ar_data array array values
  #' @param ar_prob_data array probability mass for each element along `ar_data`, sums to 1

  fl_mean <- sum(ar_data*ar_prob_data)
  fl_std <- sqrt(sum(ar_prob_data*(ar_data - fl_mean)^2))
  return(fl_std)
}
```

##### Inequality in Group Means

We compute a number of aggregate statistics over the means of the subgroups. We can only compute 1 gini and 1 standard deviation, but we have a spectrum of Atkinson index values, depending in the inequality aversion value picked.

First, we get the data inputs we need.
```{r}
# 1. SORT FIRST!
df_excess_pollution_burden_sorted <- df_excess_pollution_burden %>% arrange(pm10_grp_mean)
# 2. Obtain the means across groups, and also excess burden across groups
ar_data_grp_means <- df_excess_pollution_burden_sorted %>% pull(pm10_grp_mean)
ar_data_grp_exc_burden <- df_excess_pollution_burden_sorted %>% pull(pm10_grp_exc_burden)
# 3. Obtain the probability mass for each group
ar_data_grp_shares <- df_excess_pollution_burden_sorted %>% pull(popgrp_mass)
```

Second, we compute the GINI and STD over group-specific means, with group-specific population shares.
```{r}
# compute gini over group means, and standard deviations
fl_grp_means_gini <- ffi_dist_gini_random_var_pos_test(ar_data_grp_means, ar_data_grp_shares)
# STD
fl_grp_means_std <- ffi_std_drv(ar_data_grp_means, ar_data_grp_shares)
```

Third, compute an array of Atkinson index with differing inequality aversions.
```{r}
# Log10 scaled Inequality Measures
ar_rho <- 1 - (10^(c(seq(-2,2, length.out=30))))
tb_rho <- as_tibble(unique(ar_rho))
# Array of atkinson values
ar_grp_means_atkinson <- apply(tb_rho, 1, function(row){
  ffi_atkinson_random_var_ineq(ar_data_grp_means, ar_data_grp_shares, row[1])})
# atkinson results table
ar_st_varnames <- c('id','rho','atkinson_index')
tb_atkinson <- as_tibble(cbind(ar_rho, ar_grp_means_atkinson)) %>%
  rowid_to_column(var = "id") %>%
  rename_all(~c(ar_st_varnames)) %>%
  mutate(one_minus_rho = 1 - rho)
# Max Atkinson
fl_atkinson_max <- max(tb_atkinson %>% pull(atkinson_index))
# display
it_rows_shown <- 10
st_caption <- paste0('Atkinson Inequality Index',
  '(', it_rows_shown, ' of ', length(ar_rho), ' inequality preferences shown)')
tb_atkinson[round(seq(1, length(ar_rho), length.out = it_rows_shown)),] %>%
  kable(caption = st_caption) %>%
  kable_styling_fc()
```

Fourth, compute the standard deviation of excess burden. Note that excess burden is both positive and negative, so we can not compute GINI or Atkinson statistics from it.
```{r}
fl_grp_exc_burden_std <- ffi_std_drv(ar_data_grp_exc_burden, ar_data_grp_shares)
```

##### Visualize Inequality in Group Means (GINI and ATKINSON)

Now we can present what the existing literature focuses on, which is to compute Atkinson index over group means. As can be seen in the figure below, this index value is arbitrary depending on inequality aversion. Additionally, this value shows in this testing example, very low inequality in pollution exposure. Additionally, we also have the GINI index, which can be thought of as mapping to a particular level of inequality aversion.

Potentially, an issue with using these types of measures is that they are hard to interpret, potentially arbitrary, and potentially generates misleading impressions. In the test examples here, there is almost equality using GINI and Atkinson measures.

Our excess pollution burden statistics, however, provides a more clearly interpretable lense for looking at across group inequality. Additionally, we also look at within group inequality.

```{r}
# x-labels
x.labels <- c('lambda=0.99', 'lambda=0.90', 'lambda=0', 'lambda=-10', 'lambda=-100')
x.breaks <- c(0.01, 0.10, 1, 10, 100)

# title line 2
fl_grp_means_gini_fmt <- round(fl_grp_means_gini, 3)
st_title <- paste0("Inequality Over Group Means (GINI=", fl_grp_means_gini_fmt," and ATKINSON)")
title_line1 <- paste0("The literature computes group means and computes Atkinson Index over group means")
title_line2 <- paste0("BLACK = Atkinson index's values differ depending on inequality aversion (x-axis)")
title_line3 <- paste0("RED = GINI (", fl_grp_means_gini_fmt,") index has a fixed value")

# Graph Results--Draw
pl_gini_atkinson <- tb_atkinson %>%
  ggplot(aes(x=one_minus_rho, y=atkinson_index)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept=fl_grp_means_gini, linetype='dashed', color='red', size=2) +
  # geom_vline(xintercept=c(1), linetype="dotted") +
  labs(title = st_title,
       subtitle = paste0(title_line1, '\n', title_line2, '\n', title_line3),
       x = 'log10 Rescale of lambda, Log10(1-lambda)\nlambda=1 Utilitarian (Cares about mean pollution only), lambda=-infty Rawlsian (Cares about inequality only)',
       y = paste0('GINI and Atkinson Index, 0 to 1, 0 = perfect equality'),
       caption = 'Based on simulated random data for testing.') +
  scale_x_continuous(trans='log10', labels = x.labels, breaks = x.breaks) +
  ylim(0, max(min(fl_atkinson_max*1.1, 1), fl_grp_means_gini)) +
  theme(text = element_text(size = 10),
        legend.position="right")

# Print
print(pl_gini_atkinson)
```

##### Visualize Inequality in Group Means (Excess Burden)

Now we visualize excess pollution burden as a distribution, this is our preferred statistics to look at variations in means across groups.

```{r}
# Rounding excess burden s.d.
fl_grp_exc_burden_std_fmt <- round(fl_grp_exc_burden_std, 3)
st_title <- paste0("Distribution of Excess Burden (Across Group Variation), s.d.=", fl_grp_exc_burden_std_fmt)
title_line1 <- paste0("Histogram shows the distribution of excess burden by population groups")
title_line2 <- paste0("Excess Burden = (Pollution Share)/(Pop Share) - 1")

# Generate a Data Sample by Drawing from the Distribution
it_sample_draws <- 1e6
ar_it_draws <- sample(1:it_N_pop_groups, it_sample_draws, replace=TRUE, prob=ar_data_grp_shares)
ar_sample_draws <- ar_data_grp_exc_burden[ar_it_draws]
# Draw histogram
pl_excess_burden <- as_tibble(ar_sample_draws) %>%
  ggplot(aes(x=value)) +
  # geom_histogram(aes(y=..density..), 
  #                colour="darkblue", fill="lightblue")+
  geom_density(alpha=.2, fill="#FF6666") +
  geom_vline(aes(xintercept=0),
             color="blue", linetype="dashed", size=2) +
  labs(title = st_title,
       subtitle = paste0(title_line1, '\n', title_line2),
       x = 'Excess Pollution Burden  = (Pollution Share)/(Pop Share) - 1',
       y = 'Density',
       caption = 'Based on simulated random data for testing.')
# Print
print(pl_excess_burden)
```
