---
title: "R4Econ Refreshes Package, Generates RMD based PDF, HTML and R Files"
titleshort: "Knit PDF, HTML and Curl R based on RMD"
description: |
  Knit PDF, HTML and Curl R based on RMD.
core:
  - package: REconTools
    code: |
      ff_sup_clean_rmd()
date: 2020-04-18
date_start: 2020-04-18
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Process RMD Files

```{r global_options, include = FALSE}
rm(list = ls(all.names = TRUE))
try(source("C:/Users/fan/R4Econ/.Rprofile"))
library(reticulate)
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

R4Econ Rmd to PDF and HTML.

```{r}
# file_path_root <- 'C:/Users/fan/R4Econ/support/development/fs_pkg_r4econ'
# file_path_r <- paste0(file_path_root, '.R')
# file_path_rmd <- paste0(file_path_root, '.Rmd')
# knitr::purl(file_path_rmd, output=file_path_r, documentation = 1)
```

#### Local Outputs

```
file:///C:/Users/fan/R4econ/bookdown/index.html
file:///C:/Users/fan/R4econ/amto/array/
file:///C:/Users/fan/R4econ/panel/basic/
```


#### Testing Which Files Have Changed

Check through *R4Econ* and see which files have been added, and which files have been modified. bl_test set to TRUE. Not generating any new files, just checking.

Note: new files that should not get pdf and html should have *.rmd* rather than *.Rmd* as suffix.

```{r}
bl_update_book = TRUE

# run examples
spt_root <- 'C:/Users/fan/R4Econ'
spt_log <- 'C:/Users/fan/R4Econ/_log/rmdclean/'
sink(paste0(spt_log, 'rmd_clean_test.log'))
start_time <- Sys.time()
setwd(spt_root)

bl_test = TRUE
bl_gen_if_git_old = FALSE

ar_spt_root = 'C:/Users/fan/R4Econ/'
# ar_spt_root = 'C:/Users/fan/R4Econ/development/inout/'

ff_sup_clean_rmd(ar_spt_root = ar_spt_root,
                 ar_spn_skip = c('support', 'index.Rmd', 'main.Rmd', 
                                 '_mod.Rmd', 'title.Rmd', 'drm'),
                 st_folder_pdf = '/htmlpdfr/',
                 st_folder_html = '/htmlpdfr/',
                 st_folder_R = '/htmlpdfr/',
                 bl_gen_if_git_old = bl_gen_if_git_old,
                 bl_recursive = TRUE,
                 bl_verbose = FALSE,
                 bl_test = bl_test, 
                 it_hierachy_shift=2, it_toc_depth=3)

end_time <- Sys.time()
print(paste0('ff_sup_clean_rmd(), bl_test = TRUE, took:', end_time - start_time))
sink()
```

#### Run And Generate Pdfs and HTMLs Updates Only

Only generate pdf and html for files that have git changes or are new.


```{r}
spt_root <- 'C:/Users/fan/R4Econ'
spt_log <- 'C:/Users/fan/R4Econ/_log/rmdclean/'
sink(paste0(spt_log, 'rmd_clean_actual.log'))
start_time <- Sys.time()
setwd(spt_root)

bl_test = FALSE
bl_gen_if_git_old = FALSE

# ar_spt_root = c('C:/Users/fan/R4Econ/amto',
#                 'C:/Users/fan/R4Econ/development',
#                 'C:/Users/fan/R4Econ/function',
#                 'C:/Users/fan/R4Econ/linreg',
#                 'C:/Users/fan/R4Econ/math',
#                 'C:/Users/fan/R4Econ/optimization',
#                 'C:/Users/fan/R4Econ/panel',
#                 'C:/Users/fan/R4Econ/regnonlin',
#                 'C:/Users/fan/R4Econ/tabgraph',
#                 'C:/Users/fan/R4Econ/summarize')

ar_spt_root = 'C:/Users/fan/R4Econ/'
# ar_spt_root = 'C:/Users/fan/R4Econ/development/inout/'

ff_sup_clean_rmd(ar_spt_root = ar_spt_root,
                 ar_spn_skip = c('support', 'index.Rmd', 'README', 
                                 'main.Rmd', '_mod.Rmd', 'title.Rmd','drm',
                                 'Panel Data and Optimization with R'),
                 st_folder_pdf = '/htmlpdfr/',
                 st_folder_html = '/htmlpdfr/',
                 st_folder_R = '/htmlpdfr/',
                 bl_gen_if_git_old = bl_gen_if_git_old,
                 bl_recursive = TRUE,
                 bl_verbose = TRUE,
                 bl_test = bl_test,
                 it_hierachy_shift=2, it_toc_depth=3)

end_time <- Sys.time()
print(paste0('ff_sup_clean_rmd(), bl_test = FALSE, took:', end_time - start_time))
sink()
```

#### Update README.md File and Bookdown Appendix

There is python file that parses rmd front matters and goes through rmd files in the bookdown yaml file to automatically generate README contents.

- 'C:/Users/fan/Py4Econ/support/inout/bookdownparse.py'

```{python}
import pyfan.util.rmd.bookdownparse as ft_bookdownparse

# Path Roots
sfc_prj = 'R4Econ'
sph_prj = 'C:/Users/fan/R4Econ/'
spn_prj_bookdown_yml = '_bookdown.yml'

# Read me preamble, and endnotes
spn_prj_readme_pre = 'README_pre.md'
spn_prj_readme_end = 'README_end.md'
spn_prj_readme_toc = 'README_toc.md'
# combine preamble, contents, suffix together
spn_prj_readme_ful = 'README.md'

# FIles to Ignore
ls_st_ignore=['index.Rmd', 'README_appendix.md', 'title.Rmd', 'main.Rmd']

# Generate README TOC
ft_bookdownparse.fs_yml2readme(sfc_prj=sfc_prj, sph_prj=sph_prj,
                               spn_prj_bookdown_yml=spn_prj_bookdown_yml,
                               spn_prj_readme_toc=spn_prj_readme_toc,
                               ls_st_ignore=ls_st_ignore)

# Combine README Files
filenames = [spn_prj_readme_pre, spn_prj_readme_toc, spn_prj_readme_end]
with open(sph_prj + spn_prj_readme_ful, 'w') as outfile:
    for fname in filenames:
        sph_full_fnmae = sph_prj + fname
        with open(sph_full_fnmae) as infile:
            outfile.write(infile.read())
            outfile.write('\n')
            
# Also Generate Appendix for bookdown file 
```

Generate contenxt for bookdown appendix.

```{python}
sph_prj = 'C:/Users/fan/R4Econ/'
spn_prj_readme_toc = 'README_toc.md'
spn_prj_appendix = 'README_appendix.md'
spn_in = open(sph_prj + spn_prj_readme_toc, "r")
spn_out = open(sph_prj + spn_prj_appendix, "w")

spn_out.write('# (APPENDIX) Appendix {-}\n\n')
spn_out.write('# Index and Code Links\n\n')

for line in spn_in:
    if '# ' in line:
        st_text = line.replace('### ', '')
        st_text = st_text.replace('## ', '')
        st_text = st_text.replace('# ', '')
        st_text = st_text.lstrip('0123456789.- ')

        st_line_prefix = ''
        st_line_suffix = ''
        st_head_level = ''
        it_level = 0
        if '### ' in line:
            line = line.replace('### ', '')
            st_head_level = 'Subsection'
            line = st_head_level + ' ' + line
            it_level = 3
        elif '## ' in line:
            line = line.replace('## ', '')
            st_head_level = 'Section'
            line = st_head_level + ' ' + line
            st_line_prefix = '### '
            st_line_suffix = ' links'
            it_level = 2
        elif '# ' in line:
            line = line.replace('# ', '')
            st_head_level = ''
            line = st_head_level + '' + st_text
            st_line_prefix = '## '
            st_line_suffix = ' links'
            it_level = 1

        line = line.replace('  ', ' ')
        line = line.replace('\n', '')
        st_text = st_text.replace('\n', '')

        if it_level == 1:
            line = st_line_prefix + line + st_line_suffix + '\n'
        else:
            line = st_line_prefix + '[' + line + '][' + st_text + ']' + st_line_suffix + '\n'

    spn_out.write(line)

# (APPENDIX) Appendix {-}

# Code Summary and Links

spn_in.close()
spn_out.close()
```


#### Update Bookdown

Render HTML:

```{r}
if (bl_update_book) {
  spt_root <- 'C:/Users/fan/R4Econ'
  spt_log <- 'C:/Users/fan/R4Econ/_log/rmdclean/'
  
  sink(paste0(spt_log, 'render_bookdown_html.log'))
  start_time <- Sys.time()
  setwd(spt_root)
  
  # render html
  bookdown::render_book('index.Rmd', 'bookdown::gitbook')
  
  end_time <- Sys.time()
  print(paste0('bookdown::render_book, html, took:', end_time - start_time))
  sink()
}
```

Render PDF: 

```{r}
if (bl_update_book) {
  spt_root <- 'C:/Users/fan/R4Econ'
  spt_log <- 'C:/Users/fan/R4Econ/_log/rmdclean/'
  
  sink(paste0(spt_log, 'render_bookdown_pdf.log'))
  start_time <- Sys.time()
  setwd(spt_root)
  
  bookdown::render_book('index.Rmd', 'bookdown::pdf_book')
  
  end_time <- Sys.time()
  print(paste0('bookdown::render_book, pdf, took:', end_time - start_time))
  sink()
}
```

