---
title: "R Expand Each Dataframe Row into More Rows"
titleshort: "Expand Each Dataframe Row into More Rows"
description: |
  Generate row value specific arrays of varying Length, and stack expanded dataframe.
core:
  - package: dplyr
    code: |
      do()
      unnest()
      left_join()
      df %>% group_by(ID) %>% do(inc = rnorm(.$Q, mean=.$mean, sd=.$sd)) %>% unnest(c(inc))
date: 2020-04-19
date_start: 2020-04-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Mx1 to MxQ Rows

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

**Case One**: There is a dataframe with $M$ rows, based on these $m$ specific information, generate dataframes for each $m$. Stack these indivdiual dataframes together and merge original $m$ specific information in as well. The number of rows for each $m$ is $Q_m$, each $m$ could have different number of expansion rows.

Generate a panel with $M$ individuals, each individual is observed for different spans of times (*uncount*). Before expanding, generate individual specific normal distribution standard deviation. All individuals share the same mean, but have increasing standard deviations.

#### Generate Dataframe with M Rows.

This is the first step, generate $M$ rows of data, to be expanded. Each row contains the number of normal draws to make and the mean and the standard deviation for normal daraws that are $m$ specific.

```{r setup_data}
# Parameter Setups
it_M <- 3
it_Q_max <- 5
fl_rnorm_mu <- 1000
ar_rnorm_sd <- seq(0.01, 200, length.out=it_M)
ar_it_q <- sample.int(it_Q_max, it_M, replace=TRUE)

# N by Q varying parameters
mt_data = cbind(ar_it_q, ar_rnorm_sd)
tb_M <- as_tibble(mt_data) %>% rowid_to_column(var = "ID") %>%
                rename(sd = ar_rnorm_sd, Q = ar_it_q) %>%
                mutate(mean = fl_rnorm_mu)

# display
kable(tb_M) %>%
  kable_styling_fc()
```

#### Random Normal Draw Expansion

The steps are:

1. [do anything](https://dplyr.tidyverse.org/reference/do.html)
2. use ".$" sign to refer to variable names, or [['name']]
3. unnest
4. left_join expanded and original

Note these all give the same results

Use dot dollar to get variables

```{r normal draw expansion dot dollar}
# Generate $Q_m$ individual specific incomes, expanded different number of times for each m
tb_income <- tb_M %>% group_by(ID) %>%
  do(income = rnorm(.$Q, mean=.$mean, sd=.$sd)) %>%
  unnest(c(income))

# Merge back with tb_M
tb_income_full_dd <- tb_income %>%
  left_join(tb_M)

# display
kable(tb_income) %>%
  kable_styling_fc()
kable(tb_income_full_dd) %>%
  kable_styling_fc()
```
