---
title: "R Do Anything Function over Dataframe Subset and Stack Output Dataframes, (MxP by N) to (MxQ by N+Z-1)"
titleshort: "Dataframe Subset to Dataframe (MxP by N) to (MxQ by N+Z-1)"
description: |
  Group by mini dataframes as inputs for function. Stack output dataframes with group id.
core:
  - package: dplyr
    code: |
      group_by()
      do()
      unnest()
date: 2020-05-27
date_start: 2020-05-27
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### (MxP by N) to (MxQ by N+Z)

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

There is a dataframe composed of *M* mini-dataframes. Group by a variable that identifies each unique sub-dataframe, and use the sub-dataframes  with *P* rows as inputs to a function.

The function outputs *Q by Z* rows and columns of results, stack the results. The output file has *MxQ* rows and the *Z* columns of additional results should be appended.

#### Generate the MxP by N Dataframe

*M* Grouping characteristics, *P* rows for each group, and *N* Variables. 

1. *M* are individuals
2. *P* are dates
3. A wage variable for individual wage at each date. And a savings varaible as well.

```{r}
# Define
it_M <- 3
it_P <- 5
svr_m <- 'group_m'
svr_mp <- 'info_mp'

# dataframe
set.seed(123)
df_panel_skeleton <- as_tibble(matrix(it_P, nrow=it_M, ncol=1)) %>%
  rowid_to_column(var = svr_m) %>%
  uncount(V1) %>%
  group_by(!!sym(svr_m)) %>% mutate(!!sym(svr_mp) := row_number()) %>%
  ungroup() %>%
  rowwise() %>% mutate(wage = rnorm(1, 100, 10), 
                       savings = rnorm(1, 200, 30)) %>%
  ungroup() %>% 
  rowid_to_column(var = "id_ji")

# Print
kable(df_panel_skeleton) %>% kable_styling_fc()
```

#### Subgroup Compute and Expand

Use the *M* sub-dataframes, generate *Q* by *Z* result for each of the *M* groups. Stack all results together.

Base on all the wages for each individual, generate individual specific mean and standard deviations. Do this for three things, the wage variable, the savings variable, and the sum of wage and savings:

1. *Z=2*: 2 columns, mean and standard deviation
2. *Q=3*: 3 rows, statistics based on wage, savings, and the sum of both

First, here is the processing function that takes the dataframe as input, with a parameter for rounding:

```{r}
# define function
ffi_subset_mean_sd <- function(df_sub, it_round=1) {
  #' A function that generates mean and sd for several variables
  #'
  #' @description
  #' Assume there are two variables in df_sub wage and savings
  #'
  #' @param df_sub dataframe where each individual row is a different 
  #' data point, over which we compute mean and sd, Assum there are two 
  #' variables, savings and wage
  #' @param it_round integer rounding for resulting dataframe
  #' @return a dataframe where each row is aggregate for a different type
  #' of variablea and each column is a different statistics
  
  fl_wage_mn = mean(df_sub$wage)
  fl_wage_sd = sd(df_sub$wage)
  
  fl_save_mn = mean(df_sub$savings)
  fl_save_sd = sd(df_sub$savings)
  
  fl_wgsv_mn = mean(df_sub$wage + df_sub$savings)
  fl_wgsv_sd = sd(df_sub$wage + df_sub$savings)  
  
  ar_mn <- c(fl_wage_mn, fl_save_mn, fl_wgsv_mn)
  ar_sd <- c(fl_wage_sd, fl_save_sd, fl_wgsv_sd)
  ar_st_row_lab <- c('wage', 'savings', 'wage_and_savings')
  
  mt_stats <- cbind(ar_mn, ar_sd)
  mt_stats <- round(mt_stats, it_round)
  
  ar_st_varnames <- c('mean', 'sd', 'variables')
  df_combine <- as_tibble(mt_stats) %>% 
    add_column(ar_st_row_lab) %>%
    rename_all(~c(ar_st_varnames)) %>%
    select(variables, 'mean', 'sd') %>%
    rowid_to_column(var = "id_q")
    
  return(df_combine)
}
# testing function
ffi_subset_mean_sd(df_panel_skeleton %>% filter(!!sym(svr_m)==1))
```


Second, call *ffi_subset_mean_sd* function for each of the groups indexed by *j* and stack results together with *j* index:

1. group by 
2. call function 
3. unnest

```{r}
# run group stats and stack dataframes
df_outputs <- df_panel_skeleton %>% group_by(!!sym(svr_m)) %>%
  do(df_stats = ffi_subset_mean_sd(., it_round=2)) %>%
  unnest() %>%
  rowid_to_column(var = "id_mq")
# print
kable(df_outputs) %>% kable_styling_fc()
```

In the resulting file, we went from a matrix with *MxP* rows to a matrix with *MxQ* Rows. 
