---
title: "DPLYR Evaluate Function over N Individuals with Individual Specific and Shared Arrays of Parameters"
titleshort: "Nonlinear Function of Scalars and Arrays over Rows"
description: |
  Five methods to evaluate scalar nonlinear function over matrix.
  Evaluate non-linear function with scalar from rows and arrays as constants.
core:
  - package: r
    code: |
      .$fl_A
      fl_A=$`(., 'fl_A')
      .[[svr_fl_A]]
  - package: dplyr
    code: |
      rowwise()
      mutate(out = funct(inputs))
date: 2020-04-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Row Input Functions

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

We want evaluate nonlinear function f(Q_i, y_i, ar_x, ar_y, c, d), where c and d are constants, and ar_x and ar_y are arrays, both fixed. x_i and y_i vary over each row of matrix. We would like to evaluate this nonlinear function concurrently across $N$ individuals. The eventual goal is to find the $i$ specific $Q$ that solves the nonlinear equations.

This is a continuation of [R use Apply, Sapply and dplyr Mutate to Evaluate one Function Across Rows of a Matrix](https://fanwangecon.github.io/R4Econ/function/noloop/fs_applysapplymutate.html)

#### Set up Input Arrays

There is a function that takes $M=Q+P$ inputs, we want to evaluate this function $N$ times. Each time, there are $M$ inputs, where all but $Q$ of the $M$ inputs, meaning $P$ of the $M$ inputs, are the same. In particular, $P=Q*N$.

$$M = Q+P = Q + Q*N$$

```{r setup_data}
# it_child_count = N, the number of children
it_N_child_cnt = 5
# it_heter_param = Q, number of parameters that are heterogeneous across children
it_Q_hetpa_cnt = 2

# P fixed parameters, nN is N dimensional, nP is P dimensional
ar_nN_A = seq(-2, 2, length.out = it_N_child_cnt)
ar_nN_alpha = seq(0.1, 0.9, length.out = it_N_child_cnt)
ar_nP_A_alpha = c(ar_nN_A, ar_nN_alpha)
ar_nN_N_choice = seq(1,it_N_child_cnt)/sum(seq(1,it_N_child_cnt))

# N by Q varying parameters
mt_nN_by_nQ_A_alpha = cbind(ar_nN_A, ar_nN_alpha, ar_nN_N_choice)

# Convert Matrix to Tibble
ar_st_col_names = c('fl_A', 'fl_alpha', 'fl_N')
tb_nN_by_nQ_A_alpha <- as_tibble(mt_nN_by_nQ_A_alpha) %>% rename_all(~c(ar_st_col_names))

# Show
kable(tb_nN_by_nQ_A_alpha) %>%
  kable_styling_fc()
```

#### Mutate over Simple Function

For this example, use a very simple function with only one type of input, all inputs are scalars.

```{r}
# Define Implicit Function
ffi_nonlinear <- function(fl_A, fl_alpha){

  fl_out <- (fl_A + fl_alpha*fl_A)/(fl_A)^2

  return(fl_out)
}
```

Apply the function over the dataframe, note five different ways below, the third way allows for parameters to be strings.

```{r}
# variable names
svr_fl_A <- 'fl_A'
svr_fl_alpha <- 'fl_alpha'

# Evaluate
tb_nN_by_nQ_A_alpha_mutate_rows <- tb_nN_by_nQ_A_alpha %>%
  mutate(fl_out_m1 = ffi_nonlinear(fl_A=.$fl_A, fl_alpha=.$fl_alpha),
         fl_out_m2 = ffi_nonlinear(fl_A=`$`(., 'fl_A'), fl_alpha=`$`(., 'fl_alpha')),
         fl_out_m3 = ffi_nonlinear(fl_A=.[[svr_fl_A]], fl_alpha=.[[svr_fl_alpha]]),
         fl_out_m4 = ffi_nonlinear(fl_A=fl_A, fl_alpha=fl_alpha),
         fl_out_m5 = ffi_nonlinear(fl_A, fl_alpha))

# print
kable(tb_nN_by_nQ_A_alpha_mutate_rows) %>% kable_styling_fc()
```

#### Testing Function with Scalar and Arrays

Test non-linear Equation.

```{r nonlinear_test}
# Test Parameters
fl_N_agg = 100
fl_rho = -1
fl_N_q = ar_nN_N_choice[4]*fl_N_agg
ar_A_alpha = mt_nN_by_nQ_A_alpha[4,]
# Apply Function
ar_p1_s1 = exp((ar_A_alpha[1] - ar_nN_A)*fl_rho)
ar_p1_s2 = (ar_A_alpha[2]/ar_nN_alpha)
ar_p1_s3 = (1/(ar_nN_alpha*fl_rho - 1))
ar_p1 = (ar_p1_s1*ar_p1_s2)^ar_p1_s3
ar_p2 = fl_N_q^((ar_A_alpha[2]*fl_rho-1)/(ar_nN_alpha*fl_rho-1))
ar_overall = ar_p1*ar_p2
fl_overall = fl_N_agg - sum(ar_overall)
print(fl_overall)
```

Implement the non-linear problem's evaluation using apply over all $N$ individuals.

```{r nonlinear_func}
# Define Implicit Function
ffi_nonlin_dplyrdo <- function(fl_A, fl_alpha, fl_N, ar_A, ar_alpha, fl_N_agg, fl_rho){
  # ar_A_alpha[1] is A
  # ar_A_alpha[2] is alpha

  # # Test Parameters
  # fl_N = 100
  # fl_rho = -1
  # fl_N_q = 10

  # Apply Function
  ar_p1_s1 = exp((fl_A - ar_A)*fl_rho)
  ar_p1_s2 = (fl_alpha/ar_alpha)
  ar_p1_s3 = (1/(ar_alpha*fl_rho - 1))
  ar_p1 = (ar_p1_s1*ar_p1_s2)^ar_p1_s3
  ar_p2 = fl_N^((fl_alpha*fl_rho-1)/(ar_alpha*fl_rho-1))
  ar_overall = ar_p1*ar_p2
  fl_overall = fl_N_agg - sum(ar_overall)

  return(fl_overall)
}

# Parameters
fl_rho = -1

# Evaluate Function
print(ffi_nonlin_dplyrdo(mt_nN_by_nQ_A_alpha[1,1],
                         mt_nN_by_nQ_A_alpha[1,2],
                         mt_nN_by_nQ_A_alpha[1,3]*fl_N_agg,
                         ar_nN_A, ar_nN_alpha, fl_N_agg, fl_rho))
for (i in seq(1,dim(mt_nN_by_nQ_A_alpha)[1])){
  fl_eval = ffi_nonlin_dplyrdo(mt_nN_by_nQ_A_alpha[i,1],
                               mt_nN_by_nQ_A_alpha[i,2],
                               mt_nN_by_nQ_A_alpha[i,3]*fl_N_agg,
                               ar_nN_A, ar_nN_alpha, fl_N_agg, fl_rho)
  print(fl_eval)
}
```

#### Evaluate Nonlinear Function using dplyr mutate

```{r nonlinear_dplyr_mutate}

# Define Implicit Function
ffi_nonlin_dplyrdo <- function(fl_A, fl_alpha, fl_N, ar_A, ar_alpha, fl_N_agg, fl_rho){

  # Test Parameters
  # ar_A = ar_nN_A
  # ar_alpha = ar_nN_alpha
  # fl_N = 100
  # fl_rho = -1
  # fl_N_q = 10

  # Apply Function
  ar_p1_s1 = exp((fl_A - ar_A)*fl_rho)
  ar_p1_s2 = (fl_alpha/ar_alpha)
  ar_p1_s3 = (1/(ar_alpha*fl_rho - 1))
  ar_p1 = (ar_p1_s1*ar_p1_s2)^ar_p1_s3
  ar_p2 = (fl_N*fl_N_agg)^((fl_alpha*fl_rho-1)/(ar_alpha*fl_rho-1))
  ar_overall = ar_p1*ar_p2
  fl_overall = fl_N_agg - sum(ar_overall)

  return(fl_overall)
}

# fl_A, fl_alpha are from columns of tb_nN_by_nQ_A_alpha
tb_nN_by_nQ_A_alpha = tb_nN_by_nQ_A_alpha %>% rowwise() %>%
                        mutate(dplyr_eval = ffi_nonlin_dplyrdo(fl_A, fl_alpha, fl_N,
                                                               ar_nN_A, ar_nN_alpha,
                                                               fl_N_agg, fl_rho))
# Show
kable(tb_nN_by_nQ_A_alpha) %>%
  kable_styling_fc()
```
