---
title: "Examples of Lists and Named One and Two Dimensional Lists in R"
titleshort: "Multi-dimensional Named Lists"
description: |
  Initiate Empty List. Named one and two dimensional lists.
  Collapse named and unamed list to string and print input code.
core:
  - package: r
    code: |
      deparse(substitute())
      vector(mode = "list", length = it_N)
      names(list) <- paste0('e',seq())
      dimnames(ls2d)[[1]] <- paste0('r',seq())
      dimnames(ls2d)[[2]] <- paste0('c',seq())
  - package: tidyr
    code: |
      unnest()
date: 2020-05-14
date_start: 2020-04-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Multiple Dimensional List

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

- r list tutorial
- r vector vs list
- r initialize empty multiple element list
- r name rows and columns of 2 dimensional list
- r row and colum names of list
- list dimnames
- r named list to string


#### One Dimensional Named List

1. define list
2. slice list
3. print r named list as a single line string

- [R Unlist named list into one string with preserving list names](https://stackoverflow.com/a/55622024/8280804)

```{r}
# Define Lists
ls_num <- list(1,2,3)
ls_str <- list('1','2','3')
ls_num_str <- list(1,2,'3')

# Named Lists
ar_st_names <- c('e1','e2','e3')
ls_num_str_named <- ls_num_str
names(ls_num_str_named) <- ar_st_names

# Add Element to Named List
ls_num_str_named$e4 <- 'this is added'
```

- r print input as string
- r print parameter code as string
- [How to convert variable (object) name into String](https://stackoverflow.com/a/14577878/8280804)

```{r}
# list to String printing function
ffi_lst2str <- function(ls_list, st_desc, bl_print=TRUE) {
  
  # string desc
  if(missing(st_desc)){
    st_desc <- deparse(substitute(ls_list))
  }
    
  # create string
  st_string_from_list = paste0(paste0(st_desc, ':'), 
                               paste(names(ls_list), ls_list, sep="=", collapse=";" ))
  
  if (bl_print){
    print(st_string_from_list)
  }
}

# print full
ffi_lst2str(ls_num)
ffi_lst2str(ls_str)
ffi_lst2str(ls_num_str)
ffi_lst2str(ls_num_str_named)

# print subset
ffi_lst2str(ls_num[2:3])
ffi_lst2str(ls_str[2:3])
ffi_lst2str(ls_num_str[2:4])
ffi_lst2str(ls_num_str_named[c('e2','e3','e4')])
```

#### Two Dimensional Unnamed List

Generate a multiple dimensional list:

1. Initiate with an N element empty list
2. Reshape list to M by Q
3. Fill list elements
4. Get list element by row and column number

List allows for different data types to be stored together.

Note that element specific names in named list are not preserved when the list is reshaped to be two dimensional. Two dimensional list, however, could have row and column names.

```{r}
# Dimensions
it_M <- 2
it_Q <- 3
it_N <- it_M*it_Q

# Initiate an Empty MxQ=N element list
ls_2d_flat <- vector(mode = "list", length = it_N)
ls_2d <- ls_2d_flat

# Named flat
ls_2d_flat_named <- ls_2d_flat
names(ls_2d_flat_named) <- paste0('e',seq(1,it_N))
ls_2d_named <- ls_2d_flat_named

# Reshape
dim(ls_2d) <- c(it_M, it_Q)
# named 2d list can not carry 1d name after reshape
dim(ls_2d_named) <- c(it_M, it_Q)
```

Print Various objects generated above, print list flattened.

```{r}
# display
ffi_lst2str(ls_2d_flat_named)
# print(ls_2d_flat_named)
ffi_lst2str(ls_2d_named)
print(ls_2d_named)
```

Select element from list:

```{r}
# Select Values, double bracket to select from 2dim list
print('ls_2d[[1,2]]')
print(ls_2d[[1,2]])
```

#### Define Two Dimensional Named LIst

For naming two dimensional lists, *rowname* and *colname* does not work. Rather, we need to use *dimnames*. Note that in addition to dimnames, we can continue to have element specific names. Both can co-exist. But note that the element specific names are not preserved after dimension transform, so need to be redefined afterwards.

How to select an element of a two dimensional list:

1. row and column names: *dimnames*, *ls_2d_flat_named[['row2','col2']]*
2. named elements: *names*, *ls_2d_flat_named[['e5']]*
3. select by index: *index*, *ls_2d_flat_named[[5]]*
4. converted two dimensional named list to tibble/matrix

Neither *dimnames* nor *names* are required, but both can be used to select elements.

```{r}
# Dimensions
it_M <- 3
it_Q <- 4
it_N <- it_M*it_Q

# Initiate an Empty MxQ=N element list
ls_2d_flat_named <- vector(mode = "list", length = it_N)
dim(ls_2d_flat_named) <- c(it_M, it_Q)

# Fill with values
for (it_Q_ctr in seq(1,it_Q)) {
  for (it_M_ctr in seq(1,it_M)) {
    # linear index
    ls_2d_flat_named[[it_M_ctr, it_Q_ctr]] <- (it_Q_ctr-1)*it_M+it_M_ctr
  }
}

# Replace row names, note rownames does not work
dimnames(ls_2d_flat_named)[[1]] <- paste0('row',seq(1,it_M))
dimnames(ls_2d_flat_named)[[2]] <- paste0('col',seq(1,it_Q))

# Element Specific Names
names(ls_2d_flat_named) <- paste0('e',seq(1,it_N))

# Convert to Matrix
tb_2d_flat_named <- as_tibble(ls_2d_flat_named) %>% unnest()
mt_2d_flat_named <- as.matrix(tb_2d_flat_named)
```

Print various objects generated above:

```{r}
# These are not element names, can still name each element
# display
print('ls_2d_flat_named')
print(ls_2d_flat_named)
print('tb_2d_flat_named')
print(tb_2d_flat_named)
print('mt_2d_flat_named')
print(mt_2d_flat_named)
```

Select elements from list: 

```{r}
# Select elements with with dimnames
ffi_lst2str(ls_2d_flat_named[['row2','col2']])
# Select elements with element names
ffi_lst2str(ls_2d_flat_named[['e5']])
# Select elements with index
ffi_lst2str(ls_2d_flat_named[[5]])
```

<!-- # Fill values
ls_2d_fill <- ls_2d
ls_2d_fill[1,1] -->
