---
title: "R Example Binary Outcome Nonlinear Estimation"
output:
  pdf_document: default
  word_document: default
  html_document: default
  html_notebook: default
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/CodeDynaAsset/)'s [R4Econ](https://fanwangecon.github.io/R4Econ/) Repository or [Intro Stats with R](https://fanwangecon.github.io/Stat4Econ/) Repository.

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
rm(list = ls(all.names = TRUE))
options(knitr.duplicate.label = 'allow')
```

```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(tidyverse)
library(knitr)
library(kableExtra)
library(REconTools)
```

# Logit Regression

## Logit Regresion

There is a variable, sort by this variable, then generate index from 1 to N representing sorted values of this index. If there are repeating values, still assign index, different index each value.

- r generate index sort
- dplyr mutate equals index
- [LOGIT REGRESSION R DATA ANALYSIS EXAMPLES](https://stats.idre.ucla.edu/r/dae/logit-regression/)
- [Generalized Linear Models](https://www.statmethods.net/advstats/glm.html)

```{r}
df_mtcars <- mtcars
# Regress
rs_logit <- glm(am ~ mpg, data = df_mtcars, family = "binomial")
summary(rs_logit)
confint(rs_logit)
# Predcit Using Regresion Data
df_mtcars$pred_prob <- predict(rs_logit, newdata = df_mtcars, type = "response")
```
