---
title: "R Generate and Combine Fixed and Random Matrix"
titleshort: "Matrix Basics"
description: |
  Generate and combine NA, fixed and random matrixes. Name columns and rows.
  Sort all rows and all columns of a matrix.
  Replace values outside min and max in matrix by NA values. 
core:
  - package: R
    code: |
      rep()
      rbind()
      matrix(NA)
      matrix(NA_real_)
      matrix(NA_integer_)
      colnames()
      rownames()
      t(apply(mt, 1, sort))
      apply(mt, 2, sort)
      colMeans
      rowMeans
      which()
date: 2022-07-14
date_start: 2020-04-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Generate Matrixes

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Create a N by 2 Matrix from 3 arrays

Names of each array become row names automatically.

```{r fixed matrix}
ar_row_one <- c(-1,+1)
ar_row_two <- c(-3,-2)
ar_row_three <- c(0.35,0.75)

mt_n_by_2 <- rbind(ar_row_one, ar_row_two, ar_row_three)
kable(mt_n_by_2) %>%
  kable_styling_fc()
```

#### Name Matrix Columns and Rows

```{r}
# An empty matrix with Logical NA
mt_named <- matrix(data=NA, nrow=2, ncol=2)
colnames(mt_named) <- paste0('c', seq(1,2))
rownames(mt_named) <- paste0('r', seq(1,2))
mt_named
```

#### Generate NA Matrix

- [Best way to allocate matrix in R, NULL vs NA?](https://stackoverflow.com/a/26724451/8280804)

Allocate with NA or NA_real_ or NA_int_. Clarity in type definition is preferred.

```{r}
# An empty matrix with Logical NA
mt_na <- matrix(data=NA, nrow=2, ncol=2)
str(mt_na)

# An empty matrix with numerica NA
mt_fl_na <- matrix(data=NA_real_, nrow=2, ncol=2)
mt_it_na <- matrix(data=NA_integer_, nrow=2, ncol=2)

str(mt_fl_na)
str(mt_fl_na)
```

#### Generate Random Matrixes

Random draw from the normal distribution, random draw from the uniform distribution, and combine resulting matrixes.

```{r random matrix}
# Generate 15 random normal, put in 5 rows, and 3 columns
mt_rnorm <- matrix(rnorm(15,mean=0,sd=1), nrow=5, ncol=3)

# Generate 15 random normal, put in 5 rows, and 3 columns
mt_runif <- matrix(runif(15,min=0,max=1), nrow=5, ncol=5)

# Combine
mt_rnorm_runif <- cbind(mt_rnorm, mt_runif)

# Display
kable(round(mt_rnorm_runif, 3)) %>% kable_styling_fc()
```

#### Replace a Subset of Matrix Values by NA_real_

For values in matrix that fall below or above some thresholds, we will replace these values by NA_real_.

```{r}
fl_max_val <- 0.8
fl_min_val <- 0.2
mt_rnorm_runif_bd <- mt_rnorm_runif
mt_rnorm_runif_bd[which(mt_rnorm_runif < fl_min_val)] <- NA_real_
mt_rnorm_runif_bd[which(mt_rnorm_runif > fl_max_val)] <- NA_real_
# Print
print(mt_rnorm_runif_bd)
```

#### Sort Each Matrix Row or Column

Now we sort within each row or within each column of the random matrix.

```{r}
# Within row sort
mt_rnorm_runif_row_sort <- t(apply(
  mt_rnorm_runif, 1, sort
))
# Within column sort, note no transpose
mt_rnorm_runif_col_sort <- apply(
  mt_rnorm_runif, 2, sort
)
# Display
kable(round(mt_rnorm_runif_row_sort, 3),
      caption="Each row sort low to high") %>%
  kable_styling_fc()
kable(round(mt_rnorm_runif_col_sort, 3),
      caption="Each column sort low to high") %>%
  kable_styling_fc()
```

#### Compute Column and Row Statistics

Compute column and row means, and also column and row sums

```{r}
print(paste0('colSums=',
             paste(round(
               colSums(mt_rnorm_runif),3), collapse=',')
             ))
print(paste0('colMeans=',
             paste(round(
               colMeans(mt_rnorm_runif),3), collapse=',')
             ))
print(paste0('rowSums=',
             paste(round(
               rowSums(mt_rnorm_runif),3), collapse=',')
             ))
print(paste0('rowMeans=',
             paste(round(
               rowMeans(mt_rnorm_runif),3), collapse=',')
             ))
```


#### Add Column to Matrix with Common Scalar Value

Given some matrix of information, add a column, where all rows of the column have the same numerical value. Use the matrix created prior.
- R add column to matrix
- r append column to matrix constant value

```{r}
fl_new_first_col_val <- 111
fl_new_last_col_val <- 999
mt_with_more_columns <- cbind(rep(fl_new_first_col_val, dim(mt_rnorm_runif)[1]),
                              mt_rnorm_runif,
                              rep(fl_new_last_col_val, dim(mt_rnorm_runif)[1]))
# Display
kable(mt_with_more_columns) %>% kable_styling_fc_wide()
```
