---
title: "Find the Closest Point Along a Line to Another Point"
titleshort: "Find the Closest Point Along a Line to Another Point"
description: |
  A line crosses through the origin, what is the closest point along this line to another point.
  Graph several functions jointly with points and axis.
core:
  - package: graphics
    code: |
      par(mfrow = c(1, 1))
      curve(fc)
      points(x, y)
      abline(v=0, h=0)
date: 2021-06-25
date_start: 2021-06-25
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Find Nearest

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Neart Point Along a Line Through Orgin to Another Point

We first have $X_1$ an $Y_1$, given these, we are able to generate $R =\frac{Y_2}{X_2}$, a ratio. We want to iteratively update $X_1$ and $Y_1$, where $1$ subscript indicates the first iteration, but we only know the ratio. Think of $R$ as a line through the origin with $R$ as the slope.

We generate $X_2$ and $Y_2$ by finding the point along the $R$ as slope origin line that is the closest to the $X_1$ and $Y_1$. At the resulting point, $R$ will be respected, and it will differ least in distance to the earlier iteration's $X_1$ and $Y_1$ points.

1. The slope of the diagonal line is $-\frac{X_2}{Y_2}=-\frac{1}{R}$
2. The diagonal line must cross $X_1$ and $Y_1$, solve for this line's y-intercept
3. Solve for the intersection of the diagonal line and the origin line with $R$ as slope

Implementing step (2):

$$
\begin{aligned}
Y_1 &= I - \frac{1}{R}\cdot X_1\\
I &= Y_1 + \frac{X_1}{R} \\
I &=  \frac{Y_1\cdot R + X_1}{R}\\
\end{aligned}
$$

Implementing step (3):
$$
\begin{aligned}
Y &= \frac{Y_1\cdot R + X_1}{R} - \frac{1}{R}\cdot X\\
Y &= R \cdot X \\
R \cdot X &= \frac{Y_1\cdot R + X_1}{R} - \frac{1}{R}\cdot X\\
\left(R + \frac{1}{R}\right) \cdot X &= \frac{Y_1\cdot R + X_1}{R}\\
\frac{R^2 + 1}{R} \cdot X &= \frac{Y_1\cdot R + X_1}{R}\\
X &= \frac{Y_1\cdot R + X_1}{R} \cdot \frac{R}{R^2 + 1}\\
\end{aligned}
$$

And we have:
$$
\begin{aligned}
X &= \frac{Y_1\cdot R + X_1}{R^2 + 1}\\
Y &= \frac{Y_1\cdot R^2 + X_1\cdot R}{R^2 + 1}\\
\end{aligned}
$$

Visualize the results:

```{r, fig.width=5.25, fig.height=5.625}
# Set random parameter Values for X1, Y1, and X2/Y2 ratio
set.seed(3)
fl_x1 <- runif(1) * 10
fl_y1 <- runif(1) * 10
fl_r <- runif(1) * 5

# Diaganol
fl_diag_slope <- -1 / fl_r
fl_diag_yintercept <- (fl_y1 * fl_r + fl_x1) / fl_r

# Closest point
fl_x2 <- (fl_y1 * fl_r + fl_x1) / (fl_r^2 + 1)
fl_y2 <- (fl_y1 * fl_r^2 + fl_x1 * fl_r) / (fl_r^2 + 1)

# Print state
print(paste("x1=", fl_x1, "x2=", fl_x2, "R=", fl_r, sep = " "))
print(paste("x2=", fl_x2, "y2=", fl_y2, sep = " "))

# X and y lims
ar_xylim <- c(-1, max(fl_x1, fl_y2) * 1.5)

# Visualize
par(mfrow = c(1, 1))
# Line through origin
curve(0 + fl_r * x, ar_xylim[1], ar_xylim[2],
  col = "black", lwd = 2, lty = 1,
  ylim = ar_xylim,
  ylab = "", xlab = ""
)
# Diaganol line
curve(fl_diag_yintercept + fl_diag_slope * x,
  add = TRUE,
  col = "blue", lwd = 2, lty = 2,
  ylim = ar_xylim,
  ylab = "", xlab = ""
)
# Point
points(fl_x1, fl_y1,
  add = TRUE,
  pch = 8, col = "red", cex = 3, lwd = 2,
  ylab = "", xlab = ""
)
points(fl_x2, fl_y2,
  add = TRUE,
  pch = 9, col = "darkgreen", cex = 3, lwd = 2,
  ylab = "", xlab = ""
)
# Origin lines
abline(
  v = 0, h = 0,
  col = "gray", lwd = 3, lty = 2
)

# Titles
title(
  main = paste0(
    "Line through origin and orthogonal line\n",
    "Find closest point along black line to red star"
  ),
  sub = paste0(
    "Black line goes through origin,",
    " blue line goes through (x1,y1) and (x2, y2)"
  ),
  xlab = "x values", ylab = "y values"
)
```
