---
title: "Quadratic and other Rescaling of Parameters with Fixed Min and Max"
titleshort: "Quadratic and other Rescaling of Parameters with Fixed Min and Max"
description: |
  Given a < x < b, use f(x) to rescale x, such that f(a)=a, f(b)=b, but f(z)=0.5*z for some z between a and b. Solve using the quadratic function with three equations and three unknowns uniquely. 
date: 2021-03-08
date_start: 2021-03-08
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Rescale a Parameter with Fixed Min and Max

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Using A Quadratic Function to Fit Three Points Uniquely

Given e < x < f, use f(x) to rescale x, such that f(e)=e, f(f)=f, but f(z)=$\alpha \cdot z$ for one particular z between e and f, where 
$\alpha$ > 1. And in this case, assume that $\alpha \cdot z < f$. We can fit these three points using the Quadratic function uniquely. In another word, there is a unique quadratic function that crosses these three points. Note the quadratic function is either concave or convex through the entire domain. 

Suppose that $e=0$, $f=10$, $z=2$, and $\alpha=1.5$. Using a quadratic to fit:

$$
y(x) = a \cdot x^2 + b \cdot x + c
$$
We have three equations:
$$
0 = a \cdot 0 + b \cdot 0 + c\\
2 \cdot 1.5 = a \cdot 2^2 + b \cdot 2 + c\\
10 = a \cdot 10^2 + b \cdot 10 + c
$$
Given these, we have, $c = 0$, and subsequently, 2 equations and 2 unknowns:
$$
3 = a \cdot 4 + b \cdot 2\\
10 = a \cdot 100 + b \cdot 10
$$

Hence:
$$
a = \frac{3-2b}{4}\\
10 = \frac{3-2b}{4}\cdot 100 + b \cdot 10\\
10 = 75 - 50b + 10b\\
$$
And finally:
$$
a = \frac{3-2*1.625}{4}=-0.0625\\
b = \frac{65}{40}=1.625\\
c = 0
$$

Generate the $a$, $b$ and $c$ points above for the quadratic function:
```{r}
# set values
e <- 0
f <- 10
z <- 2
alpha <- 1.5
# apply formulas from above
a <- -0.0625
b <- 1.625
c <- 0
# grid of values beween a and b, 11 points covering z = 2
ar_x <- seq(e, f, length.out = 11)
# rescale
ar_grid_quad <- a*ar_x^2 + b*ar_x + c
# show values
kable(print(as_tibble(cbind(ar_x, ar_grid_quad))), 
      caption = paste0("Quadratic Fit of Three Equations and Three Unknowns\n",
                       "Satisfies: f(0)=0, f(10)=10, f(2)=3")) %>% 
  kable_styling_fc()
```





We have three equations:
$$
0 = a \cdot 0 + b \cdot 0 + c\\
0.75 = a \cdot 0.5^2 + b \cdot 0.5 + c\\
3.5 = a \cdot 3.5^2 + b \cdot 3.5 + c
$$
Given these, we have, $c = 0$, and subsequently, 2 equations and 2 unknowns:
$$
0.75 = a \cdot 0.25 + b \cdot 0.5\\
3.5 = a \cdot 12.25 + b \cdot 3.5
$$

Hence:
$$
a = \frac{0.75-0.5b}{0.25}\\
3.5 = \frac{0.75-0.5b}{0.25}\cdot 12.25 + b \cdot 3.5\\
3.5 = 36.75 - 24.5b + 3.5b\\
$$
And finally:
$$
a = \frac{0.75-0.5*1.58333}{0.25}=-0.1666\\
b = \frac{36.75-3.5}{24.5-3.5}=1.58333\\
c = 0
$$
Generate the $a$, $b$ and $c$ points above for the quadratic function:

```{r}
# set values
e <- 0
f <- 3.5
z <- 0.5
alpha <- 1.5
# apply formulas from above
a <- -0.16666666
b <- 1.583333333
c <- 0
# grid of values beween a and b, 11 points covering z = 2
ar_x <- seq(e, f, length.out = 100000)
# rescale
ar_grid_quad <- a*ar_x^2 + b*ar_x + c
# show values
# cbind(ar_x, ar_grid_quad)
ar_x[which.min(abs(ar_grid_quad - 0.75))]
```
