---
title: "Factor Basics with R"
description: |
  Basics of unobserved factor models.
core:
  - package: r
    code: |
      rep()
      head()
      tail()
      na_if()
date: 2020-11-12
date_start: 2020-11-12
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---


### Factor Basics

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Unobserved Factor and Measurements

##### Ability and English and Math Scores

There are two tests, $j=1$ the math test, and $j=2$ the word test. Student ID=1 takes both tests. What determines his test scores? (1) how the test is designed, max and min points, and grading difficulty and points per question; (2) his own ability; (3) the extend to which the test reflects his ability versus random other factors. 

We have:

$$Z_{i,j=\text{math}} = \mu_{j=\text{math}} + \alpha_{j=\text{math}}\cdot\theta_{i} + \epsilon_{i,j=\text{math}}$$

Note that there are three possibilities for index:

1. An element is $i$ specific: $\theta$
2. An element is $j$ specific: $\mu$ and $\alpha$
3. An element is $i,j$ specific: $Z$ and $\epsilon$

We can ignore the $i$ subscripts, and think of $Z$ and $\theta$ and $\epsilon$ as vectors:

$$Z_{j} = \mu_{j} + \alpha_{j}\cdot{\theta} + \epsilon_{j}$$

##### Simulate Scores and Ability

We are interested in the distribution of $\theta$, when we have multiple test scores, suppose:

1. $\theta$ is normal, and suppose its mean is 0.
2. Suppose $\alpha_{\text{math}} = 1$.
3. $\epsilon$ are zero mean and normal.

Set some parameters:

```{r}
# define parameters fixed
alpha_math <- 1
theta_cognitive_mean <- 0

# Parameters to estiamtes
mu_math <- 60
mu_engl <- 80
sd_eps_math <- 6
sd_eps_engl <- 3
# alpha_math fixed at 1
alpha_engl <- 5
# theta_cognitive_mean fixed at 0
theta_cognitive_stdv <- 5
```

Simulate Math and English scores:

```{r}
# set seed
set.seed(1)

# Number of students
N <- 100

# Draw vectors of the epsilon terms
ar_eps_math <- rnorm(N, mean=0, sd=sd_eps_math)
ar_eps_engl <- rnorm(N, mean=0, sd=sd_eps_engl)

# Draw the Theta Ability vector
ar_theta_cognitive <- rnorm(N, mean=theta_cognitive_mean, sd=theta_cognitive_stdv)

# Simulate the English and Math Test Scores
ar_math <- mu_math + alpha_math*ar_theta_cognitive + ar_eps_math
ar_engl <- mu_engl + alpha_engl*ar_theta_cognitive + ar_eps_engl
```

Generate a Dataframe with all the information:

```{r}
# Gather all data together
mt_data <- cbind(ar_math, ar_engl, ar_theta_cognitive, ar_eps_math, ar_eps_engl)

# Column Names
ar_st_varnames <- c('id', 'scores_math','scores_english','cog_ability','math_eps', 'engl_eps')

# Combine to tibble, add name col1, col2, etc.
tb_combine <- as_tibble(mt_data) %>%
  rowid_to_column(var = "id") %>%
  rename_all(~c(ar_st_varnames))

# Display
kable(tb_combine[1:10,]) %>% kable_styling_fc_wide()
```

Draw scatter plot of test score and ability:

```{r}
# Pivot Longer
tb_combine_long <- tb_combine %>%
  pivot_longer(cols = starts_with('scores'),
               names_to = c('test'),
               names_pattern = paste0("scores_(.*)"),
               values_to = "scores") %>%
  mutate(scores = as.numeric(scores))

# Draw Scatter Plot
scatter <- ggplot(tb_combine_long, aes(x=cog_ability, y=scores,
                                       colour=test)) +
      geom_jitter(size=1, width = 0.15) +
      labs(title = 'Coginitive Ability and Test Scores',
           x = 'Unobserved Cognitive Ability',
           y = 'Test Scores in Math and English',
           caption = 'Simulated Data https://github.com/FanWangEcon/R4Econ') +
      theme_bw()
print(scatter)
```

##### Mean, Variance and Covariance Formula

Note that we assume that $\mathrm{E}[\theta]=0$, and $\alpha_m=1$, and we also assume that the measurement errors for Math and English scores are uncorrelated. 

First, we can either set one of the $\mu$ to be 0 or $\mathrm{E}[\theta]$ to 0. From the mean scores, given our assumption, we can identify $\mu_m$ and $\mu_e$:

$$\mathrm{E}[Z_{m}] = \mu_m + \alpha_m\cdot\mathrm{E}[\theta] = \mu_m$$
$$\mathrm{E}[Z_{e}] = \mu_e + \alpha_e\cdot\mathrm{E}[\theta] = \mu_e $$

Second, we have two observed variance and one observed covariance, since $\alpha_m=1$, we have three equations with three unknowns, where the three unkonwns are $\mathrm{Var}[\theta]$, $\mathrm{Var}[\epsilon_m]$ and $\alpha_e$:

$$\mathrm{Var}[Z_{m}] = \alpha_m^2 \cdot\mathrm{Var}[\theta] + \mathrm{Var}[\epsilon_m] = \mathrm{Var}[\theta] + \mathrm{Var}[\epsilon_m]$$
$$\mathrm{Var}[Z_{e}] = \alpha_e^2 \cdot\mathrm{Var}[\theta] + \mathrm{Var}[\epsilon_e]$$

$$\mathrm{Cov}[Z_{m}, Z_{e}] = \alpha_m \cdot \alpha_e \cdot\mathrm{Var}[\theta] = \alpha_e \cdot\mathrm{Var}[\theta] $$

Third, replace the $\alpha_e$ in the variance of English equation as a function of covariance:

$$\mathrm{Var}[Z_{e}] = \left(\frac{\mathrm{Cov}[Z_{m}, Z_{e}]}{\mathrm{Var}[\theta]}\right)^2 \cdot\mathrm{Var}[\theta] + \mathrm{Var}[\epsilon_e]$$

##### Estimate/Solve for the Measurement Equation Parameters


We can compute a number of objects from the data:

1. the mean of Math scores
2. the mean of English scores
3. the variance of Math scores
4. the variance of English scores
5. the covariance of Math and English scores

```{r}
# 1. the mean of Math scores
fl_mean_math = mean(tb_combine[['scores_math']])
# 2. the mean of English scores
fl_mean_english = mean(tb_combine[['scores_english']])
# 3. the variance of Math scores
fl_var_math = cov(tb_combine[['scores_math']], tb_combine[['scores_math']])
# 4. the variance of English scores
fl_var_english = cov(tb_combine[['scores_english']], tb_combine[['scores_english']])
# 5. the covariance of Math and English scores
fl_cov = cov(tb_combine[['scores_math']], tb_combine[['scores_english']])
# Print means
print(paste0('fl_mean_math=', fl_mean_math, ', fl_mean_english=', fl_mean_english))
# Print var and cov
print(paste0('fl_cov=', fl_cov, ', fl_var_math=', fl_var_math, ', fl_var_english=', fl_var_english))
```
