---
title: "Approximate Normal Random Variable with Binomial Discrete Random Variable"
titleshort: "Binomial Approximation of Normal"
description: |
  Approximate a continuous normal random variable with a discrete binomial random variable.
core:
  - package: r
    code: |
      hist()
      plot()
  - package: stats
    code: |
      dbinom()
      rnorm()
date: 2021-06-13
date_start: 2021-06-13
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Discrete Approximation of Continuous Random Variables

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Use Binomial Discrete Random Variable to Approximate Continuous Normal

First, draw from a Continuous Random Variable. Sample $N$ draws from a normal random variable.

```{r}
# Random normal Data Vector (not equal outcomes)
set.seed(123)
it_sample_N <- 10000
fl_rnorm_mean <- 100
fl_rnorm_sd <- 6
ar_data_rnorm <- rnorm(it_sample_N, mean=fl_rnorm_mean, sd=fl_rnorm_sd)
# Visualize
par(new=FALSE)
hist(ar_data_rnorm, xlab = 'x Values', ylab = 'Frequencies', main = '')
title(main = 'Continuous Normal Random Variable Draws')
grid()
```
We use the [binomial to approximate the normal distribution](https://en.wikipedia.org/wiki/Binomial_distribution), we know that these relationships are approximately true, where $\mu$ and $\sigma$ are the mean and standard deviations of the normal random variable, and $n$ and $p$ and the number of "trials" and the "probability-of-success" for the binomial distribution:

$$
\begin{aligned}
\mu &= n\cdot p\\
n &= \frac{\mu}{p}\\
\sigma^2 &= n\cdot p\cdot(1-p) = \mu\cdot(1-p)\\
\end{aligned}
$$

Given these, we have can translate between the normal random variable's parameters and the binomial discrete random variable's parameters:

$$
\begin{aligned}
p &= 1 - \frac{\sigma^2}{\mu}\\
n &= \frac{\mu}{1 - \frac{\sigma^2}{\mu}} = \frac{\mu}{\frac{\mu - \sigma^2}{\mu}} = \frac{\mu^2}{\mu - \sigma^2}
\end{aligned}
$$

There are two important aspects to note here:

1. Since $p$ must be positive, this means $\frac{\sigma^2}{\mu}<1$ and $\sigma^2<\mu$, which is the condition for the above transformation to work.
2. The binomial discrete random variable will have non-zero mass for very small probability events at the left-tail. These very low outcome events are highly unlikely to be observed or drawn from sampling the continuous random variable. The presence of these left-tail values might impact the computation of certain statistics, for example the [Atkinson Index for highly inequality averse planners](https://fanwangecon.github.io/R4Econ/math/func_ineq/htmlpdfr/fs_atkinson_ces.html). 

```{r}
# Use binomial to approximate normal
fl_p_binom <- 1 - fl_rnorm_sd^2/fl_rnorm_mean
fl_n_binom <- round(fl_rnorm_mean^2/(fl_rnorm_mean - fl_rnorm_sd^2))
fl_binom_mean <- fl_n_binom*fl_p_binom
fl_binom_sd <- sqrt(fl_n_binom*fl_p_binom*(1-fl_p_binom))
# Mean and sd
print(paste0('BINOMI mean=', fl_binom_mean))
print(paste0('BINOMI mean=', fl_binom_sd))
# drv = discrete random variable
ar_drv_rbinom_xval <- seq(1, fl_n_binom)
ar_drv_rbinom_prob <- dbinom(ar_drv_rbinom_xval, size=fl_n_binom, prob=fl_p_binom)
# ignore weight at x=0
ar_drv_rbinom_prob <- ar_drv_rbinom_prob/sum(ar_drv_rbinom_prob)
```

Visualize the binomial discrete random variable:
```{r}
# graph
par(new=FALSE)
ar_ylim = c(0,1)
ffi_atkinson_ineq(ar_data_rnorm, -1)
ar_atkinson <- apply(mt_rho, 1, function(row){ffi_atkinson_ineq(ar_data_rnorm, row[1])})
plot(ar_drv_rbinom_xval, ar_drv_rbinom_prob, xlab = 'Binomial x Values', ylab = 'Discrete Probabilitys, P(x)')
title(main = paste0('Binomial Approximate of Normal Random Variable'),
      sub = paste0('binop=', round(fl_p_binom, 2), ';binon=', round(fl_n_binom, 2), ';binomean=', round(fl_binom_mean, 2), ';binomsd=', round(fl_binom_sd, 2), ';normmean=', round(fl_rnorm_mean, 2), ';normsd=', round(fl_rnorm_sd, 2)))
grid()
```
