---
title: "Convert R Markdown File to R, PDF and HTML"
description: |
  Find all files in a folder with a particula suffix, with exclusion.
  Convert R Markdow File to R, PDF and HTML.
  Modify markdown pounds hierarchy.
core:
  - package: r
    code: |
      file()
      writeLines()
      readLines()
      close()
      gsub()
date: 2020-05-16
date_start: 2020-05-16
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Rmd to HTML

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

#### Search and Find all Files in Repository

Search inside directories, for all files in a repository that have a particular suffix and that don't contain skip pattern list string items.

```{r}
# Serch Folder and skip list
spt_roots <- c('C:/Users/fan/R4Econ/amto', 'C:/Users/fan/R4Econ/summarize')
spn_skip <- c('summarize', 'panel', 'support')

# Search and get all Path
ls_sfls <- list.files(path=spt_roots, recursive=T, pattern=".Rmd", full.names=T)

# Skip path if contains words in skip list
if(!missing(spn_skip)) {
  ls_sfls <- ls_sfls[!grepl(paste(spn_skip, collapse = "|"), ls_sfls)]
}

# Loop and print
for (spt_file in ls_sfls) {
    st_fullpath_nosufx <- tail(strsplit(spt_file, "/")[[1]],n=1)
    print(paste0(spt_file, '---', st_fullpath_nosufx))
}
```

#### Search and Find all Git Modified or New Rmd

Search inside directories, for all files in a git repo folder that are new or have been modified. Ignore possible subset of file based on string search.

```{r}
# Serch Folder and skip list
spt_roots <- c('C:/Users/fan/R4Econ/amto', 'C:/Users/fan/R4Econ/development')
spn_skip <- c('summarize', 'panel', 'support')
ls_sfls <- list.files(path=spt_roots, recursive=T, pattern=".Rmd", full.names=T)
if(!missing(spn_skip)) {
  ls_sfls <- ls_sfls[!grepl(paste(spn_skip, collapse = "|"), ls_sfls)]
}

# Loop and print
for (spt_file in ls_sfls) {
  spg_check_git_status <- paste0('git status -s ', spt_file)
  st_git_status <- toString(system(spg_check_git_status, intern=TRUE))
  bl_modified <- grepl(' M ', st_git_status, fixed=TRUE)
  bl_anewfile <- grepl('?? ', st_git_status, fixed=TRUE)
  bl_nochange <- (st_git_status == "")

  if (bl_modified == 1) {
    print(paste0('MODIFIED: ', spt_file))
  } else if (bl_anewfile == 1) {
    print(paste0('A NEW FL: ', spt_file))
  } else {
    print(paste0('NO CHNGE: ', spt_file))
  }
}
```

#### Resave an Existing File with Different Name Different Folder

Given an existing Rmd File, Resave it with a different name (add to name suffix), and then save in a different folder:

- old file: */R4Econ/development/fs_rmd_pdf_html.Rmd*
- new file: *R4Econ/development/inout/_file/rmd/fs_rmd_pdf_html_mod.Rmd*

```{r}
# Serch Folder and skip list
spt_roots <- c('C:/Users/fan/R4Econ/development/inout/')
spn_skip <- c('_main', '_file')
ls_sfls <- list.files(path=spt_roots, recursive=T, pattern=".Rmd", full.names=T)
if(!missing(spn_skip)) {
  ls_sfls <- ls_sfls[!grepl(paste(spn_skip, collapse = "|"), ls_sfls)]
}

# Loop and print
for (spt_file in ls_sfls) {
  spt_new <- paste0(spt_roots, '_file/rmd/')
  spn_new <- paste0(spt_new, sub('\\.Rmd$', '', basename(spt_file)), '_mod.Rmd')
  print(spt_new)
  print(spn_new)

  fileConn_rd <- file(spt_file, "r")
  st_file_read <- readLines(fileConn_rd)

  fileConn_sr <- file(spn_new)
  writeLines(st_file_read, fileConn_sr)

  close(fileConn_rd)
  close(fileConn_sr)
}
```

##### Replacment Function Change Markdown Hierarchy and Add to YAML

Given an existing Rmd File, Resave it with a different name, and replace (add in) additional yaml contents.

```{r}
spn_file = '_file/rmd/fs_rmd_pdf_html_mod.Rmd'
fileConn_sr <- file(spn_file)
st_file <- readLines(fileConn_sr)
# print(st_file)

st_search <- "html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      toc_depth: 3"
st_replace <- paste0("html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      toc_depth: 3\n",
                     "    toc: true\n",
                     "    number_sections: true\n",
                     "    toc_float:\n",
                     "      collapsed: false\n",
                     "      smooth_scroll: false\n",
                     "      toc_depth: 3")
st_file_updated <- gsub(x = st_file,
                        pattern = st_search,
                        replacement = st_replace)

st_search <- "../../"
st_replace <- paste0("../../../../")
st_file_updated <- gsub(x = st_file_updated,
                        pattern = st_search,
                        replacement = st_replace)

st_file_updated <- gsub(x = st_file_updated, pattern = '# ', replacement = '# ')
st_file_updated <- gsub(x = st_file_updated, pattern = '## ', replacement = '## ')
st_file_updated <- gsub(x = st_file_updated, pattern = '### ', replacement = '# ')

spn_file = '_file/rmd/fs_rmd_pdf_html_mod.Rmd'
fileConn_sr <- file(spn_file)
st_file <- writeLines(st_file_updated, fileConn_sr)

```
