---
title: "Meshgrid Arrays in R"
output:
  pdf_document: default
  word_document: default
  html_document: default
  html_notebook: default
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/CodeDynaAsset/)'s [R4Econ](https://fanwangecon.github.io/R4Econ/) Repository or [Intro Stats with R](https://fanwangecon.github.io/Stat4Econ/) Repository.

- r expand.grid meshed array to matrix
- r meshgrid
- r array to matrix
- r reshape array to matrix

```{r GlobalOptions, echo = T, results = 'hide', message=F, warning=F}
options(knitr.duplicate.label = 'allow')
```
```{r loadlib, echo = T, results = 'hide', message=F, warning=F}
library(knitr)
library(kableExtra)
# file name
st_file_name = 'fs_meshr'
# Generate R File
purl(paste0(st_file_name, ".Rmd"), output=paste0(st_file_name, ".R"), documentation = 2)
# Generate PDF and HTML
# rmarkdown::render("C:/Users/fan/R4Econ/support/array/fs_meshr.Rmd", "pdf_document")
# rmarkdown::render("C:/Users/fan/R4Econ/support/array/fs_meshr.Rmd", "html_document")
```

# Meshgrid for R

## Define Two Arrays and Mesh Them

Given two arrays, mesh the two arrays together. 

```{r}

# use expand.grid to generate all combinations of two arrays

it_ar_A = 5
it_ar_alpha = 10

ar_A = seq(-2, 2, length.out=it_ar_A)
ar_alpha = seq(0.1, 0.9, length.out=it_ar_alpha)

mt_A_alpha = expand.grid(A = ar_A, alpha = ar_alpha)

mt_A_meshed = mt_A_alpha[,1]
dim(mt_A_meshed) = c(it_ar_A, it_ar_alpha)
  
mt_alpha_meshed = mt_A_alpha[,2]
dim(mt_alpha_meshed) = c(it_ar_A, it_ar_alpha)

# display
kable(mt_A_meshed) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
kable(mt_alpha_meshed) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

## Two Identical Arrays, Mesh to Generate Square

Two Identical Arrays, individual attributes, each column is an individual for a matrix, and each row is also an individual

```{r}
# use expand.grid to generate all combinations of two arrays

it_ar_A = 5

ar_A = seq(-2, 2, length.out=it_ar_A)
mt_A_A = expand.grid(Arow = ar_A, Arow = ar_A)
mt_Arow = mt_A_A[,1]
dim(mt_Arow) = c(it_ar_A, it_ar_A)
mt_Acol = mt_A_A[,2]
dim(mt_Acol) = c(it_ar_A, it_ar_A)

# display
kable(mt_Arow) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
kable(mt_Acol) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

