---
title: "Decompose Right Hand Side Variables from Linear Regression"
titleshort: "Regression Decomposition"
description: |
  Post multiple regressions, fraction of outcome variables' variances explained by multiple subsets of right hand side variables.
core:
  - package: dplyr
    code: |
      gather()
      group_by(var)
      mutate_at(vars, funs(mean = mean(.)))
      rowSums(mat*mat)
      mutate_if(is.numeric, funs(frac = (./value_var)))
date: 2020-04-01
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Decompose RHS

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

One runs a number of regressions. With different outcomes, and various right hand side variables.

What is the remaining variation in the left hand side variable if right hand side variable one by one is set to the average of the observed values.

- Dependency: *R4Econ/linreg/ivreg/ivregdfrow.R*

The code below does not work with categorical variables (except for dummies). Dummy variable inputs need to be converted to zero/one first.

#### Decomposition Program

```{r}
ff_lr_decompose <- function(df, vars.y, vars.x, vars.c, vars.z, vars.other.keep,
                            list.vars.tomean, list.vars.tomean.name.suffix,
                            df.reg.out = NULL,
                            graph=FALSE, graph.nrow=2) {

    vars.xc <- c(vars.x, vars.c)

    # Regressions
    # regf.iv from C:\Users\fan\R4Econ\linreg\ivreg\ivregdfrow.R
    if(is.null(df.reg.out)) {
      df.reg.out <- as_tibble(
        bind_rows(lapply(vars.y, regf.iv,
                         vars.x=vars.x, vars.c=vars.c, vars.z=vars.z, df=df)))
    }

    # Select Variables
    str.esti.suffix <- '_Estimate'
    arr.esti.name <- paste0(vars.xc, str.esti.suffix)
    str.outcome.name <- 'vars_var.y'
    arr.columns2select <- c(arr.esti.name, str.outcome.name)
    # arr.columns2select

    # Generate dataframe for coefficients
    df.coef <- df.reg.out[,c(arr.columns2select)] %>% 
      mutate_at(vars(arr.esti.name), as.numeric) %>% column_to_rownames(str.outcome.name)
    # df.coef
    # str(df.coef)

    # Decomposition Step 1: gather
    df.decompose <- df %>% 
      filter(svymthRound %in% c(12, 18, 24)) %>%
      select(one_of(c(vars.other.keep, vars.xc, vars.y))) %>%
      drop_na() %>%
      gather(variable, value, -one_of(c(vars.other.keep, vars.xc)))

    # Decomposition Step 2: mutate_at(vars, funs(mean = mean(.)))
    # the xc averaging could have taken place earlier, no difference in mean across variables
    df.decompose <- df.decompose %>%
      group_by(variable) %>%
      mutate_at(vars(c(vars.xc, 'value')), funs(mean = mean(.))) %>%
      ungroup()

    # Decomposition Step 3 With Loop
    for (i in 1:length(list.vars.tomean)) {
        var.decomp.cur <- (paste0('value', list.vars.tomean.name.suffix[[i]]))
        vars.tomean <- list.vars.tomean[[i]]
        var.decomp.cur
        df.decompose <- df.decompose %>% 
          mutate((!!var.decomp.cur) := 
                   ff_lr_decompose_valadj(., df.coef, vars.tomean, str.esti.suffix))
    }

    # Additional Statistics
    df.decompose.var.frac <- df.decompose %>%
            select(variable, contains('value')) %>%
            group_by(variable) %>%
            summarize_all(funs(mean = mean, var = var)) %>%
            select(variable, matches('value')) %>% select(variable, ends_with("_var")) %>%
            mutate_if(is.numeric, funs( frac = (./value_var))) %>%
            mutate_if(is.numeric, round, 3)

    # Graph
    g.graph.dist <- NULL
    if (graph) {
      g.graph.dist <- df.decompose %>%
          select(variable, contains('value'), -value_mean) %>%
          rename(outcome = variable) %>%
          gather(variable, value, -outcome) %>%
          ggplot(aes(x=value, color = variable, fill = variable)) +
              geom_line(stat = "density") +
              facet_wrap(~ outcome, scales='free', nrow=graph.nrow)
    }

    # Return
    return(list(dfmain = df.decompose,
                dfsumm = df.decompose.var.frac,
                graph = g.graph.dist))

}

# Support Function
ff_lr_decompose_valadj <- function(df, df.coef, vars.tomean, str.esti.suffix) {
    new_value <- (df$value +
                  rowSums((df[paste0(vars.tomean, '_mean')] - df[vars.tomean])
                          *df.coef[df$variable, paste0(vars.tomean, str.esti.suffix)]))
    return(new_value)
}
```

#### Prepare Decomposition Data

```{r}
# Library
library(tidyverse)
library(AER)

# Load Sample Data
setwd('C:/Users/fan/R4Econ/_data/')
df <- read_csv('height_weight.csv')

# Source Dependency
source('C:/Users/fan/R4Econ/linreg/ivreg/ivregdfrow.R')

# Setting
options(repr.matrix.max.rows=50, repr.matrix.max.cols=50)
```

Data Cleaning.

```{r}
# Convert Variable for Sex which is categorical to Numeric
df <- df
df$male <- (as.numeric(factor(df$sex)) - 1)
summary(factor(df$sex))
summary(df$male)
```

Parameters.

```{r}
var.y1 <- c('hgt')
var.y2 <- c('wgt')
vars.y <- c(var.y1, var.y2)
vars.x <- c('prot')
vars.c <- c('male', 'wgt0', 'hgt0', 'svymthRound')
vars.other.keep <- c('S.country', 'vil.id', 'indi.id', 'svymthRound')

# Decompose sequence
vars.tomean.first <- c('male', 'hgt0')
var.tomean.first.name.suffix <- '_A'
vars.tomean.third <- c(vars.tomean.first, 'prot')
var.tomean.third.name.suffix <- '_B'
vars.tomean.fourth <- c(vars.tomean.third, 'svymthRound')
var.tomean.fourth.name.suffix <- '_C'
list.vars.tomean = list(vars.tomean.first,
                        vars.tomean.third,
                        vars.tomean.fourth)
list.vars.tomean.name.suffix <- list(var.tomean.first.name.suffix,
                                     var.tomean.third.name.suffix,
                                     var.tomean.fourth.name.suffix)
```

#### Example Guatemala OLS

```{r}
df.use <- df %>% filter(S.country == 'Guatemala') %>% 
  filter(svymthRound %in% c(12, 18, 24))
vars.z <- NULL
list.out <- 
  ff_lr_decompose(df=df.use, vars.y, vars.x, vars.c, vars.z, vars.other.keep,
                  list.vars.tomean, list.vars.tomean.name.suffix,
                  graph=TRUE, graph.nrow=1)
options(repr.matrix.max.rows=10, repr.matrix.max.cols=50)
head(list.out$dfmain,10) %>%
  kable() %>%
  kable_styling_fc_wide()
options(repr.plot.width = 10, repr.plot.height = 4)
list.out$dfsumm %>%
  kable() %>%
  kable_styling_fc_wide()
```

#### Example Guatemala IV = vil.id

```{r}
df.use <- df %>% filter(S.country == 'Guatemala') %>% 
  filter(svymthRound %in% c(12, 18, 24))
vars.z <- c('vil.id')
list.out <- ff_lr_decompose(
  df=df.use, vars.y, vars.x, vars.c, vars.z, vars.other.keep,
  list.vars.tomean, list.vars.tomean.name.suffix,
  graph=TRUE, graph.nrow=1)
list.out$dfsumm %>%
  kable() %>%
  kable_styling_fc_wide()
options(repr.plot.width = 10, repr.plot.height = 2)
list.out$graph
```

#### Example Cebu OLS

```{r}
df.use <- df %>% filter(S.country == 'Cebu') %>% 
  filter(svymthRound %in% c(12, 18, 24))
vars.z <- NULL
list.out <- ff_lr_decompose(
  df=df.use, vars.y, vars.x, vars.c, vars.z, vars.other.keep,
  list.vars.tomean, list.vars.tomean.name.suffix,
  graph=TRUE, graph.nrow=1)
options(repr.matrix.max.rows=10, repr.matrix.max.cols=50)
head(list.out$dfmain, 10) %>%
  kable() %>%
  kable_styling_fc_wide()
options(repr.plot.width = 10, repr.plot.height = 4)
list.out$dfsumm %>%
  kable() %>%
  kable_styling_fc_wide()
```

#### Example Cebu IV

```{r}
df.use <- df %>% filter(S.country == 'Cebu') %>% 
  filter(svymthRound %in% c(12, 18, 24))
vars.z <- c('wealthIdx')
list.out <- ff_lr_decompose(
  df=df.use, vars.y, vars.x, vars.c, vars.z, vars.other.keep,
  list.vars.tomean, list.vars.tomean.name.suffix,
  graph=TRUE, graph.nrow=1)
list.out$dfsumm %>%
  kable() %>%
  kable_styling_fc_wide()
options(repr.plot.width = 10, repr.plot.height = 2)
list.out$graph
```

#### Examples Line by Line

The examples are just to test the code with different types of variables.

```{r}
df.use <- df %>% filter(S.country == 'Guatemala') %>% 
  filter(svymthRound %in% c(12, 18, 24))
dim(df.use)
```

Setting Up Parameters.

```{r}
# Define Left Hand Side Variables
var.y1 <- c('hgt')
var.y2 <- c('wgt')
vars.y <- c(var.y1, var.y2)
# Define Right Hand Side Variables
vars.x <- c('prot')
vars.c <- c('male', 'wgt0', 'hgt0', 'svymthRound')
# vars.z <- c('p.A.prot')
vars.z <- c('vil.id')
# vars.z <- NULL
vars.xc <- c(vars.x, vars.c)

# Other variables to keep
vars.other.keep <- c('S.country', 'vil.id', 'indi.id', 'svymthRound')

# Decompose sequence
vars.tomean.first <- c('male', 'hgt0')
var.tomean.first.name.suffix <- '_mh02m'
vars.tomean.second <- c(vars.tomean.first, 'hgt0', 'wgt0')
var.tomean.second.name.suffix <- '_mh0me2m'
vars.tomean.third <- c(vars.tomean.second, 'prot')
var.tomean.third.name.suffix <- '_mh0mep2m'
vars.tomean.fourth <- c(vars.tomean.third, 'svymthRound')
var.tomean.fourth.name.suffix <- '_mh0mepm2m'
list.vars.tomean = list(
#                         vars.tomean.first,
                        vars.tomean.second,
                        vars.tomean.third,
                        vars.tomean.fourth
                        )
list.vars.tomean.name.suffix <- list(
#                                     var.tomean.first.name.suffix,
                                     var.tomean.second.name.suffix,
                                     var.tomean.third.name.suffix,
                                     var.tomean.fourth.name.suffix
                                    )
```

##### Obtain Regression Coefficients from somewhere

```{r}
# Regressions
# regf.iv from C:\Users\fan\R4Econ\linreg\ivreg\ivregdfrow.R
df.reg.out <- as_tibble(
  bind_rows(lapply(vars.y, regf.iv,
                   vars.x=vars.x, vars.c=vars.c, vars.z=vars.z, df=df)))
# Regressions
# reg1 <- regf.iv(var.y = var.y1, vars.x, vars.c, vars.z, df.use)
# reg2 <- regf.iv(var.y = var.y2, vars.x, vars.c, vars.z, df.use)
# df.reg.out <- as_tibble(bind_rows(reg1, reg2))
```

```{r}
# df.reg.out 
```

```{r}
# Select Variables
str.esti.suffix <- '_Estimate'
arr.esti.name <- paste0(vars.xc, str.esti.suffix)
str.outcome.name <- 'vars_var.y'
arr.columns2select <- c(arr.esti.name, str.outcome.name)
arr.columns2select
```

```{r}
# Generate dataframe for coefficients
df.coef <- df.reg.out[,c(arr.columns2select)] %>% mutate_at(vars(arr.esti.name), as.numeric) %>% column_to_rownames(str.outcome.name)
df.coef %>%
  kable() %>%
  kable_styling_fc()
str(df.coef)
```

##### Decomposition Step 1

```{r}
# Decomposition Step 1: gather
df.decompose_step1 <- df.use %>%
                        filter(svymthRound %in% c(12, 18, 24)) %>%
                        select(one_of(c(vars.other.keep, vars.xc, vars.y))) %>%
                        drop_na() %>%
                        gather(variable, value, -one_of(c(vars.other.keep, vars.xc)))
options(repr.matrix.max.rows=20, repr.matrix.max.cols=20)
dim(df.decompose_step1)
head(df.decompose_step1, 10) %>%
  kable() %>%
  kable_styling_fc()
```

##### Decomposition Step 2

```{r}
# Decomposition Step 2: mutate_at(vars, funs(mean = mean(.)))
# the xc averaging could have taken place earlier, no difference in mean across variables
df.decompose_step2 <- df.decompose_step1 %>%
                        group_by(variable) %>%
                        mutate_at(vars(c(vars.xc, 'value')), funs(mean = mean(.))) %>%
                        ungroup()

options(repr.matrix.max.rows=20, repr.matrix.max.cols=20)
dim(df.decompose_step2)
head(df.decompose_step2,10) %>%
  kable() %>%
  kable_styling_fc_wide()
```

##### Decomposition Step 3 Non-Loop

```{r}
ff_lr_decompose_valadj <- function(df, df.coef, vars.tomean, str.esti.suffix) {
    new_value <- (df$value +
                  rowSums((df[paste0(vars.tomean, '_mean')] - df[vars.tomean])
                          *df.coef[df$variable, paste0(vars.tomean, str.esti.suffix)]))
    return(new_value)
}

# # Decomposition Step 3: mutate_at(vars, funs(mean = mean(.)))
# var.decomp.one <- (paste0('value', list.vars.tomean.name.suffix[[1]]))
# var.decomp.two <- (paste0('value', list.vars.tomean.name.suffix[[2]]))
# var.decomp.thr <- (paste0('value', list.vars.tomean.name.suffix[[3]]))
# df.decompose_step3 <- df.decompose_step2 %>%
#                         mutate((!!var.decomp.one) := f_decompose_here(., df.coef, list.vars.tomean[[1]], str.esti.suffix),
#                                (!!var.decomp.two) := f_decompose_here(., df.coef, list.vars.tomean[[2]], str.esti.suffix),
#                                (!!var.decomp.thr) := f_decompose_here(., df.coef, list.vars.tomean[[3]], str.esti.suffix))

# options(repr.matrix.max.rows=10, repr.matrix.max.cols=20)
# dim(df.decompose_step3)
# df.decompose_step3
```

##### Decomposition Step 3 With Loop

```{r}
df.decompose_step3 <- df.decompose_step2
for (i in 1:length(list.vars.tomean)) {
    var.decomp.cur <- (paste0('value', list.vars.tomean.name.suffix[[i]]))
    vars.tomean <- list.vars.tomean[[i]]
    var.decomp.cur
    df.decompose_step3 <- df.decompose_step3 %>% 
      mutate((!!var.decomp.cur) := 
               ff_lr_decompose_valadj(., df.coef, vars.tomean, str.esti.suffix))

}
options(repr.matrix.max.rows=10, repr.matrix.max.cols=20)
dim(df.decompose_step3)
head(df.decompose_step3, 10) %>%
  kable() %>%
  kable_styling_fc_wide()
```

##### Decomposition Step 4 Variance

```{r}
df.decompose_step3 %>%
        select(variable, contains('value')) %>%
        group_by(variable) %>%
        summarize_all(funs(mean = mean, var = var)) %>%
        select(matches('value')) %>% select(ends_with("_var")) %>%
        mutate_if(is.numeric, funs( frac = (./value_var))) %>%
        mutate_if(is.numeric, round, 3) %>%
  kable() %>%
  kable_styling_fc()
```

##### Graphical Results

Graphically, difficult to pick up exact differences in variance, a 50 percent reduction in variance visually does not look like 50 percent. Intuitively, we are kind of seeing standard deviation, not variance on the graph if we think abou the x-scale.

```{r}
head(df.decompose_step3 %>%
    select(variable, contains('value'), -value_mean), 10) %>%
  kable() %>%
  kable_styling_fc_wide()
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 4)
df.decompose_step3 %>%
    select(variable, contains('value'), -value_mean) %>%
    rename(outcome = variable) %>%
    gather(variable, value, -outcome) %>%
    ggplot(aes(x=value, color = variable, fill = variable)) +
        geom_line(stat = "density") +
        facet_wrap(~ outcome, scales='free', nrow=2)
```

#### Additional Decomposition Testings

```{r}
head(df.decompose_step2[vars.tomean.first],3)
head(df.decompose_step2[paste0(vars.tomean.first, '_mean')], 3)
head(df.coef[df.decompose_step2$variable, 
             paste0(vars.tomean.first, str.esti.suffix)], 3)
df.decompose.tomean.first <- df.decompose_step2 %>%
    mutate(pred_new = df.decompose_step2$value +
        rowSums((df.decompose_step2[paste0(vars.tomean.first, '_mean')] 
                 - df.decompose_step2[vars.tomean.first])
            *df.coef[df.decompose_step2$variable, 
                     paste0(vars.tomean.first, str.esti.suffix)])) %>%
        select(variable, value, pred_new)
head(df.decompose.tomean.first, 10)
df.decompose.tomean.first %>%
        group_by(variable) %>%
        summarize_all(funs(mean = mean, sd = sd))  %>%
  kable() %>%
  kable_styling_fc()
```

Note the r-square from regression above matches up with the 1 - ratio below. This is the proper decomposition method that is equivalent to r2.

```{r}
df.decompose_step2 %>%
    mutate(pred_new = df.decompose_step2$value +
        rowSums((df.decompose_step2[paste0(vars.tomean.second, '_mean')] 
                 - df.decompose_step2[vars.tomean.second])
            *df.coef[df.decompose_step2$variable, 
                     paste0(vars.tomean.second, str.esti.suffix)])) %>%
        select(variable, value, pred_new) %>%
        group_by(variable) %>%
        summarize_all(funs(mean = mean, var = var)) %>%
        mutate(ratio = (pred_new_var/value_var))  %>%
  kable() %>%
  kable_styling_fc()
```
